### 2024-11-28 joss/paper.preprint.tex
update heuristic parameter convergence plot, add tex preprint format.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={PyBOP: A Python package for battery model optimisation and
parameterisation},
  pdfauthor={Brady Planden, Nicola E. Courtier, Martin Robinson, Ferran
Brosa Planella, David A. Howey},
  pdflang={en-US},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{PyBOP: A Python package for battery model optimisation and
parameterisation}

\definecolor{c53baa1}{RGB}{83,186,161}
\definecolor{c202826}{RGB}{32,40,38}
\def \rorglobalscale {0.1}
\newcommand{\rorlogo}{%
\begin{tikzpicture}[y=1cm, x=1cm, yscale=\rorglobalscale,xscale=\rorglobalscale, every node/.append style={scale=\rorglobalscale}, inner sep=0pt, outer sep=0pt]
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(0,3.0889892500000005)},xscale=0.9943,yscale=0.9894]
    \path[fill=white,even odd rule,line join=round,miter limit=2.0] (0.0, 3.1221) rectangle (4.364, -0.0336);
  \end{scope}
  \begin{scope}[even odd rule,line join=round,miter limit=2.0,shift={(-0.025, 0.0216)}]
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.8164, 3.012) -- (1.4954, 2.5204) -- (1.1742, 3.012) -- (1.8164, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (3.1594, 3.012) -- (2.8385, 2.5204) -- (2.5172, 3.012) -- (3.1594, 3.012) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (1.1742, 0.0669) -- (1.4954, 0.5588) -- (1.8164, 0.0669) -- (1.1742, 0.0669) -- cycle;
    \path[fill=c53baa1,nonzero rule,line join=round,miter limit=2.0] (2.5172, 0.0669) -- (2.8385, 0.5588) -- (3.1594, 0.0669) -- (2.5172, 0.0669) -- cycle;
    \path[fill=c202826,nonzero rule,line join=round,miter limit=2.0] (3.8505, 1.4364).. controls (3.9643, 1.4576) and (4.0508, 1.5081) .. (4.1098, 1.5878).. controls (4.169, 1.6674) and (4.1984, 1.7642) .. (4.1984, 1.8777).. controls (4.1984, 1.9719) and (4.182, 2.0503) .. (4.1495, 2.1132).. controls (4.1169, 2.1762) and (4.0727, 2.2262) .. (4.0174, 2.2635).. controls (3.9621, 2.3006) and (3.8976, 2.3273) .. (3.824, 2.3432).. controls (3.7505, 2.359) and (3.6727, 2.367) .. (3.5909, 2.367) -- (2.9676, 2.367) -- (2.9676, 1.8688).. controls (2.9625, 1.8833) and (2.9572, 1.8976) .. (2.9514, 1.9119).. controls (2.9083, 2.0164) and (2.848, 2.1056) .. (2.7705, 2.1791).. controls (2.6929, 2.2527) and (2.6014, 2.3093) .. (2.495, 2.3487).. controls (2.3889, 2.3881) and (2.2728, 2.408) .. (2.1468, 2.408).. controls (2.0209, 2.408) and (1.905, 2.3881) .. (1.7986, 2.3487).. controls (1.6925, 2.3093) and (1.6007, 2.2527) .. (1.5232, 2.1791).. controls (1.4539, 2.1132) and (1.3983, 2.0346) .. (1.3565, 1.9436).. controls (1.3504, 2.009) and (1.3351, 2.0656) .. (1.3105, 2.1132).. controls (1.2779, 2.1762) and (1.2338, 2.2262) .. (1.1785, 2.2635).. controls (1.1232, 2.3006) and (1.0586, 2.3273) .. (0.985, 2.3432).. controls (0.9115, 2.359) and (0.8337, 2.367) .. (0.7519, 2.367) -- (0.1289, 2.367) -- (0.1289, 0.7562) -- (0.4837, 0.7562) -- (0.4837, 1.4002) -- (0.6588, 1.4002) -- (0.9956, 0.7562) -- (1.4211, 0.7562) -- (1.0118, 1.4364).. controls (1.1255, 1.4576) and (1.2121, 1.5081) .. (1.2711, 1.5878).. controls (1.2737, 1.5915) and (1.2761, 1.5954) .. (1.2787, 1.5991).. controls (1.2782, 1.5867) and (1.2779, 1.5743) .. (1.2779, 1.5616).. controls (1.2779, 1.4327) and (1.2996, 1.3158) .. (1.3428, 1.2113).. controls (1.3859, 1.1068) and (1.4462, 1.0176) .. (1.5237, 0.944).. controls (1.601, 0.8705) and (1.6928, 0.8139) .. (1.7992, 0.7744).. controls (1.9053, 0.735) and (2.0214, 0.7152) .. (2.1474, 0.7152).. controls (2.2733, 0.7152) and (2.3892, 0.735) .. (2.4956, 0.7744).. controls (2.6016, 0.8139) and (2.6935, 0.8705) .. (2.771, 0.944).. controls (2.8482, 1.0176) and (2.9086, 1.1068) .. (2.952, 1.2113).. controls (2.9578, 1.2253) and (2.9631, 1.2398) .. (2.9681, 1.2544) -- (2.9681, 0.7562) -- (3.3229, 0.7562) -- (3.3229, 1.4002) -- (3.4981, 1.4002) -- (3.8349, 0.7562) -- (4.2603, 0.7562) -- (3.8505, 1.4364) -- cycle(0.9628, 1.7777).. controls (0.9438, 1.7534) and (0.92, 1.7357) .. (0.8911, 1.7243).. controls (0.8623, 1.7129) and (0.83, 1.706) .. (0.7945, 1.7039).. controls (0.7588, 1.7015) and (0.7252, 1.7005) .. (0.6932, 1.7005) -- (0.4839, 1.7005) -- (0.4839, 2.0667) -- (0.716, 2.0667).. controls (0.7477, 2.0667) and (0.7805, 2.0643) .. (0.8139, 2.0598).. controls (0.8472, 2.0553) and (0.8768, 2.0466) .. (0.9025, 2.0336).. controls (0.9282, 2.0206) and (0.9496, 2.0021) .. (0.9663, 1.9778).. controls (0.9829, 1.9534) and (0.9914, 1.9209) .. (0.9914, 1.8799).. controls (0.9914, 1.8362) and (0.9819, 1.8021) .. (0.9628, 1.7777) -- cycle(2.6125, 1.3533).. controls (2.5889, 1.2904) and (2.5553, 1.2359) .. (2.5112, 1.1896).. controls (2.4672, 1.1433) and (2.4146, 1.1073) .. (2.3529, 1.0814).. controls (2.2916, 1.0554) and (2.2228, 1.0427) .. (2.1471, 1.0427).. controls (2.0712, 1.0427) and (2.0026, 1.0557) .. (1.9412, 1.0814).. controls (1.8799, 1.107) and (1.8272, 1.1433) .. (1.783, 1.1896).. controls (1.7391, 1.2359) and (1.7052, 1.2904) .. (1.6817, 1.3533).. controls (1.6581, 1.4163) and (1.6465, 1.4856) .. (1.6465, 1.5616).. controls (1.6465, 1.6359) and (1.6581, 1.705) .. (1.6817, 1.7687).. controls (1.7052, 1.8325) and (1.7388, 1.8873) .. (1.783, 1.9336).. controls (1.8269, 1.9799) and (1.8796, 2.0159) .. (1.9412, 2.0418).. controls (2.0026, 2.0675) and (2.0712, 2.0804) .. (2.1471, 2.0804).. controls (2.223, 2.0804) and (2.2916, 2.0675) .. (2.3529, 2.0418).. controls (2.4143, 2.0161) and (2.467, 1.9799) .. (2.5112, 1.9336).. controls (2.5551, 1.8873) and (2.5889, 1.8322) .. (2.6125, 1.7687).. controls (2.636, 1.705) and (2.6477, 1.6359) .. (2.6477, 1.5616).. controls (2.6477, 1.4856) and (2.636, 1.4163) .. (2.6125, 1.3533) -- cycle(3.8015, 1.7777).. controls (3.7825, 1.7534) and (3.7587, 1.7357) .. (3.7298, 1.7243).. controls (3.701, 1.7129) and (3.6687, 1.706) .. (3.6333, 1.7039).. controls (3.5975, 1.7015) and (3.5639, 1.7005) .. (3.5319, 1.7005) -- (3.3226, 1.7005) -- (3.3226, 2.0667) -- (3.5547, 2.0667).. controls (3.5864, 2.0667) and (3.6192, 2.0643) .. (3.6526, 2.0598).. controls (3.6859, 2.0553) and (3.7155, 2.0466) .. (3.7412, 2.0336).. controls (3.7669, 2.0206) and (3.7883, 2.0021) .. (3.805, 1.9778).. controls (3.8216, 1.9534) and (3.8301, 1.9209) .. (3.8301, 1.8799).. controls (3.8301, 1.8362) and (3.8206, 1.8021) .. (3.8015, 1.7777) -- cycle;
  \end{scope}
\end{tikzpicture}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Authors and Affiliations

\usepackage[affil-it]{authblk}
\usepackage{orcidlink}
%% \renewcommand\Authsep{, }
\setlength{\affilsep}{1em}
\author[1%
  %
  \ensuremath\mathparagraph]{Brady Planden%
    \,\orcidlink{0000-0002-1082-9125}\,%
    }
\author[1,2%
  %
  ]{Nicola E. Courtier%
    \,\orcidlink{0000-0002-5714-1096}\,%
    }
\author[3%
  %
  ]{Martin Robinson%
    \,\orcidlink{0000-0002-1572-6782}\,%
    }
\author[2,4%
  %
  ]{Ferran Brosa Planella%
    \,\orcidlink{0000-0001-6363-2812}\,%
    }
\author[1,2%
  %
  ]{David A. Howey%
    \,\orcidlink{0000-0002-0620-3955}\,%
    }

\affil[1]{Department of Engineering Science, University of Oxford,
Oxford, UK%
  }
\affil[2]{The Faraday Institution, Harwell Campus, Didcot, UK%
  }
\affil[3]{Research Software Engineering Group, University of Oxford,
Oxford, UK%
  }
\affil[4]{Mathematics Institute, University of Warwick, Coventry, UK%
  }
\affil[$\mathparagraph$]{Corresponding author: %
   %
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\date{28 November 2024}

\begin{document}
\maketitle

\section{Summary}\label{summary}

The Python Battery Optimisation and Parameterisation (\texttt{PyBOP})
package provides a set of methods for the parameterisation and
optimisation of battery models, offering both Bayesian and frequentist
approaches with example workflows to assist the user. \texttt{PyBOP} has
been developed to enable parameter identification of various battery
models, including the electrochemical and equivalent circuit models
provided by the popular open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}).

Similarly, \texttt{PyBOP} can be used for parameter design optimisation
under user-defined operating conditions across a variety of model
structures. \texttt{PyBOP} allows battery model parameterisation using a
range of methods with diagnostics on the performance and convergence of
the identified or optimised parameters. The identified parameters can be
used for prediction, on-line control and design optimisation, all of
which support improved battery utilisation and development.

\section{Statement of need}\label{statement-of-need}

\texttt{PyBOP} is a Python package designed to provide a user-friendly,
object-oriented interface for optimising battery models. \texttt{PyBOP}
leverages the open-source \texttt{PyBaMM} package
(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
of these battery models. \texttt{PyBOP} is intended to serve a broad
audience of students, engineers, and researchers in both academia and
the battery industry by enabling the use of predictive battery models
where previously this was not possible. \texttt{PyBOP} emphasises clear
and informative diagnostics and workflows for users of varying
expertise, by providing advanced optimisation and sampling algorithms.
These methods are provided through interfaces to \texttt{PINTS}
(\citeproc{ref-Clerx:2019}{Clerx et al., 2019}), \texttt{SciPy}
(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), in addition to the
PyBOP's own algorithms such as Adaptive Moment Estimation with Weight
Decay (AdamW), and Cuckoo search.

\texttt{PyBOP} supports the Battery Parameter eXchange (BPX) standard
(\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing battery
parameter sets. As these parameter sets are costly to obtain due to: the
equipment and time required for characterisation experiments, the need
for battery domain knowledge, and the computational cost of parameter
estimation. \texttt{PyBOP} reduces these costs by providing fast
computational estimation with parameter set interoperability.

This package complements other lithium-ion battery modelling packages
built around \texttt{PyBaMM}, such as \texttt{liionpack} for battery
pack simulation (\citeproc{ref-Tranter2022}{Tranter et al., 2022}) and
\texttt{pybamm-eis} for numerical impedance spectroscopy, as the
identified parameters from \texttt{PyBOP} are easily exportable.

\section{Architecture}\label{architecture}

\texttt{PyBOP} has a layered data structure designed to compute and
process the forward model predictions and to package the necessary
information for the optimisation and sampling algorithms. The forward
model is solved using the popular battery modelling package,
\texttt{PyBaMM}, with construction, parameterisation, and discretisation
managed by PyBOP's model interface to PyBaMM. This approach provides a
robust object construction process with a consistent interface between
the models and optimisers. The statistical methods and optimisation
algorithms are then constructed to interface cleanly with the forward
model predictions. Furthermore, identifiability metrics are provided
alongside the estimated parameters through Hessian approximation of the
cost functions in the frequentist workflows and posterior moments in the
Bayesian workflows.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP-high-level.pdf}
\caption{PyBOP's interface to supporting funding agencies, alongside a
visualisation of the general workflow for parameterisation and
optimisation \label{fig:high-level}}
\end{figure}

\texttt{PyBOP} formulates the inference process into four key classes,
namely the model, the problem, the cost, and the optimiser/sampler, as
shown in \autoref{fig:classes}. Each of these objects represents a base
class with child classes constructing specialised functionality for
inference or optimisation workflows. The model class constructs a
\texttt{PyBaMM} forward model for a given set of model equations
provided by \texttt{PyBaMM}, initial conditions, spatial discretisation,
and numerical solver. By composing \texttt{PyBaMM} directly into
\texttt{PyBOP}, specialised models can be constructed alongside the
standard models, which can be modified, and optimally constructed for
the inference tasks. One such example is spatial rediscretisation, which
is performed when geometric parameters are optimised. In this situation,
\texttt{PyBOP} minimally rediscretises the \texttt{PyBaMM} model while
maintaining the problem, cost, and optimiser objects, providing improved
performance benefits to users. Alongside construction of the forward
model, \texttt{PyBOP}`s model class provides methods for obtaining
sensitivities from the prediction, enabling gradient-based optimisation
algorithms. This prediction, along with it's corresponding
sensitivities, is provided to the problem class for processing and
exception control. A standardised data structure is then provided to the
cost classes, providing a distance, design, or likelihood-based metric
for optimisation. For deterministic optimisation, the optimisers
minimise the corresponding cost function or the negative log-likelihood
if a likelihood class is provided. Bayesian inference is provided by
Monte Carlo sampling classes, which accept the LogPosterior class and
sample from it using Pints' based Monte Carlo algorithms at the time of
submission. In the typical workflow, the classes in
\autoref{fig:classes} are constructed in sequence.

In addition to the core architecture, \texttt{PyBOP} provides several
specialised inference and optimisation processes. One such instance is
numerical electrochemical impedance spectroscopy predictions by
discretising the forward model into sparse mass matrix form with
accompanying auto-differentiated Jacobian. These objects are then
translated into the frequency domain with a linear solution used to
compute the battery model impedance. In this situation, the forward
models are constructed within the spatial rediscretisation workflow,
allowing for geometric parameter inference from EIS forward model
predictions. Furthermore, \texttt{PyBOP} builds on the JAX
(\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
provided by \texttt{PyBaMM} by providing JAX-based cost functions for
automatic forward model differentiation with respect to the parameters.
This functionality provides a performance improvement alongside an
interface to JAX-based inference packages, such as Numpyro
(\citeproc{ref-numpyro:2019}{Phan et al., 2019}), BlackJAX
(\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}), and Optax
(\citeproc{ref-optax:2020}{DeepMind et al., 2020}).

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP_components.drawio.png}
\caption{The core \texttt{PyBOP} architecture, showcasing the base class
interfaces. Each class provide direct mapping to a classical step in the
optimisation workflow. \label{fig:classes}}
\end{figure}

The currently implemented subclasses for the model, problem, and cost
classes are listed in \autoref{tab:subclasses}. The cost functions are
grouped by problem type, while the model and optimiser classes can be
selected in combination with any problem-cost pair.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4458}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2048}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3494}}@{}}
\caption{List of available model, problem and cost (or likelihood)
classes. \label{tab:subclasses}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Battery Models
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Problem Types
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cost / Likelihood Functions
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Battery Models
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Problem Types
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Cost / Likelihood Functions
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Single particle model (SPM) & Fitting problem & Sum squared error \\
SPM with electrolyte (SPMe) & & Root mean squared error \\
Doyle-Fuller-Newman (DFN) & & Minkowski \\
Many particle model (MPM) & & Sum of power \\
Multi-species multi-reaction (MSMR) & & Gaussian log likelihood \\
Weppner Huggins & & Maximum a posteriori \\
Equivalent circuit model (ECM) & Observer & Unscented Kalman filter \\
& Design problem & Gravimetric energy density \\
& & Volumetric energy density \\
\end{longtable}

Similarly, the current algorithms available for optimisation tasks are
presented in \autoref{tab:optimisers}. From now on, the point-based
parameterisation and design optimisation tasks will simply be referred
to as optimisation tasks. This simplification can be justified by
examining \autoref{eqn:parameterisation} and \autoref{eqn:design} and
confirming that deterministic parameterisation can be viewed as an
optimisation task to minimise a distance-based cost function.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4062}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2604}}@{}}
\caption{The currently supported optimisation algorithms classified by
candidate solution type, including gradient information. (*) Scipy
minimize has gradient and non-gradient methods.
\label{tab:optimisers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
(Meta)heuristic
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Weight decayed adaptive moment estimation (AdamW) & Covariance matrix
adaptation (CMA-ES) & Particle swarm (PSO) \\
Improved resilient backpropagation (iRProp-) & Exponential natural
(xNES) & Nelder-Mead \\
Gradient descent & Separable natural (sNES) & Cuckoo search \\
SciPy minimize (*) & SciPy differential evolution & \\
& & \\
\end{longtable}

As discussed above, \texttt{PyBOP} provides Bayesian inference methods
such as maximum a posteriori (MAP) alongside the point-based methods in
\autoref{tab:subclasses}; however, for a full Bayesian framework, Monte
Carlo sampling is implemented within \texttt{PyBOP}. These methods
construct a posterior distribution on the inference parameters, which
can be used for uncertainty and practical identifiability. The
individual sampler classes are currently composed within \texttt{PyBOP}
from the \texttt{PINTS} library, with a base sampling class implemented
for interoperability and direct integration with the \texttt{PyBOP}
model, problem, and likelihood classes. The currently supported samplers
are listed in \autoref{tab:samplers}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1810}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2155}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1379}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2069}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2586}}@{}}
\caption{Sampling methods supported by PyBOP, classified according to
the proposed method. \label{tab:samplers}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Gradient-based
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptive
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Slicing
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evolutionary
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Monomial Gamma & Delayed Rejection Adaptive & Rank Shrinking &
Differential Evolution & Metropolis Random Walk \\
No-U-Turn & Haario Bardenet & Doubling & & Emcee Hammer \\
Hamiltonian & Haario & Stepout & & Metropolis Adjusted Langevin \\
Relativistic & Rao Blackwell & & & \\
\end{longtable}

\section{Background}\label{background}

\subsection{Battery models}\label{battery-models}

In general, battery models can be written in the form of a
differential-algebraic system of equations: \begin{equation}
\frac{\mathrm{d} \mathbf{x}}{\mathrm{d} t} = f(t,\mathbf{x},\mathbf{y},\mathbf{u}(t),\mathbf{\theta}),
\label{dynamics}
\end{equation} \begin{equation}
\mathbf{y}(t) = g(t,\mathbf{x},\mathbf{y},\mathbf{u}(t),\mathbf{\theta}),
\label{output}
\end{equation} with initial conditions \begin{equation}
\mathbf{x}(0) = \mathbf{x}_0(\mathbf{\theta}).
\label{initial_conditions}
\end{equation}

Here, \(t\) is time, \(\mathbf{x}(t)\) are the (spatially discretised)
states, \(\mathbf{y}(t)\) are the outputs (e.g.~the terminal voltage),
\(\mathbf{u}(t)\) are the inputs (e.g.~the applied current) and
\(\mathbf{\theta}\) are the unknown parameters.

Common battery models include various types of equivalent circuit models
(e.g.~the Thévenin model), the Doyle--Fuller--Newman (DFN) model
(\citeproc{ref-Doyle:1993}{Doyle et al., 1993};
\citeproc{ref-Fuller:1994}{Fuller et al., 1994}) based on porous
electrode theory and its reduced-order variants including the single
particle model (SPM) (\citeproc{ref-Planella:2022}{Planella et al.,
2022}), and the multi-species, multi-reaction (MSMR) model
(\citeproc{ref-Verbrugge:2017}{Verbrugge et al., 2017}). Simplified
models that retain acceptable predictive capabilities at a lower
computational cost are widely used, for example in battery management
systems, while physics-based models are required to understand the
impact of physical parameters on battery performance. This separation of
complexity traditionally results in multiple parameterisations for a
single battery type, depending on the model structure.

\section{Examples}\label{examples}

\subsection{Parameterisation}\label{parameterisation}

The parameterisation of battery models is challenging due to the large
number of parameters that need to be identified compared to the
measurable outputs (\citeproc{ref-Andersson:2022}{Andersson et al.,
2022}; \citeproc{ref-Miguel:2021}{Miguel et al., 2021};
\citeproc{ref-Wang:2022}{Wang et al., 2022}). A complete
parameterisation often requires a stepwise identification of smaller
sets of parameters from a variety of different data sets
(\citeproc{ref-Chen:2020}{Chen et al., 2020};
\citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
et al., 2023}) and excitations.

A generic data fitting optimisation problem may be formulated as:
\begin{equation}
\min_{\mathbf{\theta}} ~ \mathcal{L}_{(\mathbf{y}_i)}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:parameterisation}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost (or
likelihood) function that quantifies the agreement between the model and
a sequence of observations \((\mathbf{y}_i)\) measured at times \(t_i\).
For gradient-based optimisers, the Jacobian of the cost function with
respect to the unknown parameters,
\((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
size and directional information.

Next, we demonstrate the fitting of synthetic data where the system
parameters are known. In this example problem, we use \texttt{PyBaMM}'s
implementation of the single particle model (SPM) with an added contact
resistance submodel. We assume that the battery model is already
parameterised except for two dynamic parameters, namely the lithium
diffusivity of the negative electrode active material particle (denoted
``negative particle diffusivity'') and the contact resistance. We
generate synthetic data from a one-hour discharge from 100\% state of
charge, to 0\% (denoted as 1C rate), followed by 30 minutes of
relaxation. This data is then corrupted with zero mean Gaussian noise of
amplitude 2mV, shown by the dots in
\autoref{fig:inference-time-landscape} (left). The initial states are
assumed known, although such an assumption is not generally necessary.
The underlying cost landscape explored by the optimiser is shown in
\autoref{fig:inference-time-landscape} (right).

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
\caption{The cost landscape for the parameterisation problem with a root
mean squared error cost function. \label{fig:inference-time-landscape}}
\end{figure}

As gradient information is available for this problem, the choice of
distance-based cost function and optimiser is not constrained. Due to
the different magnitudes of the two parameters, we apply the logarithmic
parameter transformation offered by \texttt{PyBOP}. This transforms the
optimisers search space of the optimiser to allow for a common step size
between the parameters, which is generally is not required, but improves
convergence in this problem. As a demonstration of the parameterisation
capabilities of \texttt{PyBOP}, \autoref{fig:convergence-min-max} (left)
shows the rate of convergence for each of the distance-minimising cost
functions, while \autoref{fig:convergence-min-max} (right) shows
analogous results for maximising a likelihood. The optimisation is
performed with SciPy Minimize using the gradient-based L-BFGS-B method.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/converge.png}
\caption{Convergence in the likelihood functions obtained using various
likelihood functions and the L-BFGS-B algorithm.
\label{fig:convergence-min-max}}
\end{figure}

Next, the performance of the various optimisation algorithms is
presented by category: gradient-based in
\autoref{fig:optimiser-inference} (left), evolutionary strategies in
\autoref{fig:optimiser-inference} (middle) and (meta)heuristics in
\autoref{fig:optimiser-inference} (right) for a mean squared error cost
function. Note that the performance of the optimiser depends on the cost
environment, prior information and corresponding hyperparameters for
each specific problem.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/optimisers_parameters.png}
\caption{Convergence in the parameter values obtained for the various
optimisation algorithms provided by \texttt{PyBOP}.
\label{fig:optimiser-inference}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_total.png}
\caption{Cost landscape contour plot with corresponding optimisation
traces. The top row represents the gradient-based optimisers, the middle
row is the evolution-based, and the bottom row is the (meta)heuristics.
The order from left to right corresponds to the entries in
\autoref{tab:optimisers}. \label{fig:optimiser-inference}}
\end{figure}

This parameterisation task can also be approached from a Bayesian
perspective, which we will present below using PyBOP's sampler methods.
The optimisation equation presented in equation
\autoref{eqn:parameterisation} does not represent the Bayesian parameter
identification task, and as such we introduce the Bayes theorem as,

\begin{equation}
P(\theta|D) = \frac{P(D|\theta)P(\theta)}{P(D)}
\label{eqn:bayes_theorem}
\end{equation}

where, \(P(\theta|D)\) is the posterior and represents the probability
density function of the parameter. \(P(D|\theta)\) is the likelihood
function and assesses the parameter values alongside a noise model.
\(P(\theta)\) encapsulates the prior knowledge about the parameters, and
finally \(P(D)\) is the model evidence and acts as a normalising
constant so that the final posterior is a correctly scaled density
function. Our goal in parameter inference is to identify the parameter
values with the highest probability, which can be represented as a
point-based metric or as the posterior distribution, which provides
additional information about the uncertainty of the identified
parameters. Monte Carlo sampling methods are available to obtain this
posterior distribution. These methods sample from the posterior using a
variety of methods, including gradient-based methods such as No-U-Turn
(\citeproc{ref-NUTS:2011}{Hoffman \& Gelman, 2011}) and Hamiltonian
(\citeproc{ref-Hamiltonian:2011}{Brooks et al., 2011}), as well as
heuristic methods such as Differential Evolution
(\citeproc{ref-DiffEvolution:2006}{Braak, 2006}), and finally
conventional methods based on random sampling with rejection criteria
(\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}). PyBOP offers
a sampling class that provides an interface to these samplers, which are
supported by the Probabilistic Inference of Noise Time-Series (PINTS)
package. \autoref{fig:posteriors} below shows the sampled posterior for
the synthetic workflow described above, using an adaptive
covariance-based sampler, Haario Bardenet
(\citeproc{ref-Haario:2001}{Haario et al., 2001}).

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/posteriors.png}
\caption{Posterior distributions for model parameters alongside
identified noise on the observations. Shaded area denotes confidence
bounds for each parameter. \label{fig:posteriors}}
\end{figure}

\subsection{Design optimisation}\label{design-optimisation}

Design optimisation is supported within \texttt{PyBOP} to guide future
battery design development by identifying parameter sensitivities that
can unlock improvements in battery performance. This problem can be
viewed in a similar way to the parameterisation workflows described
above, but with the aim of maximising a distance metric rather than
minimising it. In the case of design optimisation to maximise
gravimetric energy density, \texttt{PyBOP} minimises the negative of the
cost function, where the cost metric is no longer a distance between two
time series vectors, but the integrated energy of the vector normalised
by with the corresponding cell mass. This is typically quantified for
operating conditions such as a 1C discharge, at a given temperature.

Design optimisation can be written in the form of a constrained
optimisation problem as: \begin{equation}
\min_{\mathbf{\theta} \in \Omega} ~ \mathcal{L}(\mathbf{\theta}) ~~~
\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
\label{eqn:design}
\end{equation}

where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
function that quantifies the desirability of the design and \(\Omega\)
is the set of allowable parameter values.

As an example, we consider the problem of maximising the gravimetric
energy density subject to constraints on two of the geometric electrode
parameters (\citeproc{ref-Couto:2023}{Couto et al., 2023}). For this
example, we use the\texttt{PyBaMM} implementation of the single particle
model with electrolyte (SPMe) to investigate the effect of the positive
electrode thickness and the active material volume fraction on the
target cost. Since the active material volume fraction is related to the
electrode porosity, the porosity is defined with a driven constraint
from the volume fraction. In this problem, we estimate the 1C rate from
the theoretical capacity for each iteration of the design. For this
example, we employ the Particle Swarm Optimisation (PSO) algorithm.

\begin{figure}
\centering
\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/design.png}
\caption{The gravimetric landscape alongside the corresponding initial
and optimised voltage profiles for a 1C discharge.
\label{fig:design_gravimetric}}
\end{figure}

\autoref{fig:design_gravimetric} (left) shows the optimiser's search
over the gravimetric energy density parameter space. The predicted
improvement in the discharge profile between the initial and optimised
parameter values (right) for their respective applied 1C current.

\section{Acknowledgements}\label{acknowledgements}

We gratefully acknowledge all
\href{https://github.com/pybop-team/PyBOP?tab=readme-ov-file\#contributors-}{contributors}
to this package. This work was supported by the Faraday Institution
Multiscale Modelling (MSM) project (ref. FIRG059), UKRI's Horizon Europe
Guarantee (ref. 10038031), and EU IntelLiGent project (ref. 101069765).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Andersson:2022}
Andersson, M., Streb, M., Ko, J. Y., Löfqvist Klass, V., Klett, M.,
Ekström, H., Johansson, M., \& Lindbergh, G. (2022). Parametrization of
physics-based battery models from input-output data: {A} review of
methodology and current research. \emph{Journal of Power Sources},
\emph{521}(November 2021), 230859.
\url{https://doi.org/10.1016/j.jpowsour.2021.230859}

\bibitem[\citeproctext]{ref-DiffEvolution:2006}
Braak, C. J. F. T. (2006). A {Markov} {Chain} {Monte} {Carlo} version of
the genetic algorithm {Differential} {Evolution}: Easy {Bayesian}
computing for real parameter spaces. \emph{Statistics and Computing},
\emph{16}(3), 239--249. \url{https://doi.org/10.1007/s11222-006-8769-1}

\bibitem[\citeproctext]{ref-jax:2018}
Bradbury, J., Frostig, R., Hawkins, P., Johnson, M. J., Leary, C.,
Maclaurin, D., Necula, G., Paszke, A., VanderPlas, J., Wanderman-Milne,
S., \& Zhang, Q. (2018). \emph{{JAX}: Composable transformations of
{P}ython+{N}um{P}y programs} (Version 0.3.13).
\url{http://github.com/jax-ml/jax}

\bibitem[\citeproctext]{ref-Hamiltonian:2011}
Brooks, S., Gelman, A., Jones, G., \& Meng, X.-L. (2011). \emph{Handbook
of markov chain monte carlo}. Chapman; Hall/CRC.
\url{https://doi.org/10.1201/b10905}

\bibitem[\citeproctext]{ref-blackjax:2024}
Cabezas, A., Corenflos, A., Lao, J., \& Louf, R. (2024). \emph{BlackJAX:
Composable {B}ayesian inference in {JAX}}.
\url{https://arxiv.org/abs/2402.10797}

\bibitem[\citeproctext]{ref-Chen:2020}
Chen, C.-H., Brosa Planella, F., O'Regan, K., Gastol, D., Widanage, W.
D., \& Kendrick, E. (2020). Development of experimental techniques for
parameterization of multi-scale lithium-ion battery models.
\emph{Journal of The Electrochemical Society}, \emph{167}(8), 080534.
\url{https://doi.org/10.1149/1945-7111/ab9050}

\bibitem[\citeproctext]{ref-Chu:2019}
Chu, Z., Plett, G. L., Trimboli, M. S., \& Ouyang, M. (2019). A
control-oriented electrochemical model for lithium-ion battery, {P}art
{I}: {L}umped-parameter reduced-order model with constant phase element.
\emph{Journal of Energy Storage}, \emph{25}(August), 100828.
\url{https://doi.org/10.1016/j.est.2019.100828}

\bibitem[\citeproctext]{ref-Clerx:2019}
Clerx, M., Robinson, M., Lambert, B., Lei, C. L., Ghosh, S., Mirams, G.
R., \& Gavaghan, D. J. (2019). Probabilistic inference on noisy time
series ({PINTS}). \emph{Journal of Open Research Software}, \emph{7}(1),
23. \url{https://doi.org/10.5334/jors.252}

\bibitem[\citeproctext]{ref-Couto:2023}
Couto, L. D., Charkhgard, M., Karaman, B., Job, N., \& Kinnaert, M.
(2023). Lithium-ion battery design optimization based on a dimensionless
reduced-order electrochemical model. \emph{Energy}, \emph{263}(PE),
125966. \url{https://doi.org/10.1016/j.energy.2022.125966}

\bibitem[\citeproctext]{ref-optax:2020}
DeepMind, Babuschkin, I., Baumli, K., Bell, A., Bhupatiraju, S., Bruce,
J., Buchlovsky, P., Budden, D., Cai, T., Clark, A., Danihelka, I.,
Dedieu, A., Fantacci, C., Godwin, J., Jones, C., Hemsley, R., Hennigan,
T., Hessel, M., Hou, S., \ldots{} Viola, F. (2020). \emph{The
{D}eep{M}ind {JAX} {E}cosystem}. \url{http://github.com/google-deepmind}

\bibitem[\citeproctext]{ref-Doyle:1993}
Doyle, M., Fuller, T. F., \& Newman, J. (1993). {Modeling of
Galvanostatic Charge and Discharge of the Lithium/Polymer/Insertion
Cell}. \emph{Journal of The Electrochemical Society}, \emph{140}(6),
1526--1533. \url{https://doi.org/10.1149/1.2221597}

\bibitem[\citeproctext]{ref-Fuller:1994}
Fuller, T. F., Doyle, M., \& Newman, J. (1994). Simulation and
optimization of the dual lithium ion insertion cell. \emph{Journal of
The Electrochemical Society}, \emph{141}(1), 1.
\url{https://doi.org/10.1149/1.2054684}

\bibitem[\citeproctext]{ref-Haario:2001}
Haario, H., Saksman, E., \& Tamminen, J. (2001). An {Adaptive}
{Metropolis} {Algorithm}. \emph{Bernoulli}, \emph{7}(2), 223.
\url{https://doi.org/10.2307/3318737}

\bibitem[\citeproctext]{ref-NUTS:2011}
Hoffman, M. D., \& Gelman, A. (2011). \emph{The no-u-turn sampler:
Adaptively setting path lengths in hamiltonian monte carlo}.
\url{https://arxiv.org/abs/1111.4246}

\bibitem[\citeproctext]{ref-Kirk:2022}
Kirk, T. L., Lewis-Douglas, A., Howey, D., Please, C. P., \& Jon
Chapman, S. (2023). Nonlinear electrochemical impedance spectroscopy for
lithium-ion battery model parameterization. \emph{Journal of The
Electrochemical Society}, \emph{170}(1), 010514.
\url{https://doi.org/10.1149/1945-7111/acada7}

\bibitem[\citeproctext]{ref-BPX:2023}
Korotkin, I., Timms, R., Foster, J. F., Dickinson, E., \& Robinson, M.
(2023). Battery parameter eXchange. In \emph{GitHub repository}. The
Faraday Institution. \url{https://github.com/FaradayInstitution/BPX}

\bibitem[\citeproctext]{ref-metropolis:1953}
Metropolis, N., Rosenbluth, A. W., Rosenbluth, M. N., Teller, A. H., \&
Teller, E. (1953). Equation of {State} {Calculations} by {Fast}
{Computing} {Machines}. \emph{The Journal of Chemical Physics},
\emph{21}(6), 1087--1092. \url{https://doi.org/10.1063/1.1699114}

\bibitem[\citeproctext]{ref-Miguel:2021}
Miguel, E., Plett, G. L., Trimboli, M. S., Oca, L., Iraola, U., \&
Bekaert, E. (2021). Review of computational parameter estimation methods
for electrochemical models. \emph{Journal of Energy Storage},
\emph{44}(PB), 103388. \url{https://doi.org/10.1016/j.est.2021.103388}

\bibitem[\citeproctext]{ref-numpyro:2019}
Phan, D., Pradhan, N., \& Jankowiak, M. (2019). Composable effects for
flexible and accelerated probabilistic programming in NumPyro.
\emph{arXiv Preprint arXiv:1912.11554}.

\bibitem[\citeproctext]{ref-Planella:2022}
Planella, F. B., Ai, W., Boyce, A. M., Ghosh, A., Korotkin, I., Sahu,
S., Sulzer, V., Timms, R., Tranter, T. G., Zyskin, M., Cooper, S. J.,
Edge, J. S., Foster, J. M., Marinescu, M., Wu, B., \& Richardson, G.
(2022). {A Continuum of Physics-Based Lithium-Ion Battery Models
Reviewed}. \emph{Progress in Energy}, \emph{4}(4), 042003.
\url{https://doi.org/10.1088/2516-1083/ac7d31}

\bibitem[\citeproctext]{ref-Sulzer:2021}
Sulzer, V., Marquis, S. G., Timms, R., Robinson, M., \& Chapman, S. J.
(2021). {Python Battery Mathematical Modelling (PyBaMM)}. \emph{Journal
of Open Research Software}, \emph{9}(1), 14.
\url{https://doi.org/10.5334/jors.309}

\bibitem[\citeproctext]{ref-Tranter2022}
Tranter, T. G., Timms, R., Sulzer, V., Planella, F. B., Wiggins, G. M.,
Karra, S. V., Agarwal, P., Chopra, S., Allu, S., Shearing, P. R., \&
Brett, D. J. l. (2022). Liionpack: A python package for simulating packs
of batteries with PyBaMM. \emph{Journal of Open Source Software},
\emph{7}(70), 4051. \url{https://doi.org/10.21105/joss.04051}

\bibitem[\citeproctext]{ref-Verbrugge:2017}
Verbrugge, M., Baker, D., Koch, B., Xiao, X., \& Gu, W. (2017).
Thermodynamic model for substitutional materials: Application to
lithiated graphite, spinel manganese oxide, iron phosphate, and layered
nickel-manganese-cobalt oxide. \emph{Journal of The Electrochemical
Society}, \emph{164}(11), E3243.
\url{https://doi.org/10.1149/2.0341708jes}

\bibitem[\citeproctext]{ref-SciPy:2020}
Virtanen, P., Gommers, R., Oliphant, T. E., Haberland, M., Reddy, T.,
Cournapeau, D., Burovski, E., Peterson, P., Weckesser, W., Bright, J.,
van der Walt, S. J., Brett, M., Wilson, J., Millman, K. J., Mayorov, N.,
Nelson, A. R. J., Jones, E., Kern, R., Larson, E., \ldots{} SciPy 1.0
Contributors. (2020). {{SciPy} 1.0: Fundamental Algorithms for
Scientific Computing in Python}. \emph{Nature Methods}, \emph{17},
261--272. \url{https://doi.org/10.1038/s41592-019-0686-2}

\bibitem[\citeproctext]{ref-Wang:2022}
Wang, A. A., O'Kane, S. E. J., Brosa Planella, F., Houx, J. L., O'Regan,
K., Zyskin, M., Edge, J., Monroe, C. W., Cooper, S. J., Howey, D. A.,
Kendrick, E., \& Foster, J. M. (2022). Review of parameterisation and a
novel database {(LiionDB)} for continuum {Li-ion} battery models.
\emph{Progress in Energy}, \emph{4}(3), 032004.
\url{https://doi.org/10.1088/2516-1083/ac692c}

\end{CSLReferences}

\end{document}

### 2024-11-28 joss/param_plots.py

@@ -18,9 +18,9 @@ create_plot["simulation"] = False
 create_plot["landscape"] = False
 create_plot["minimising"] = False
 create_plot["maximising"] = False
-create_plot["gradient"] = True
+create_plot["gradient"] = False
 create_plot["evolution"] = False
-create_plot["heuristic"] = False
+create_plot["heuristic"] = True
 create_plot["posteriors"] = False
 
 
@@ -441,10 +441,10 @@ if create_plot["heuristic"]:
         optim = optimiser(
             cost,
             verbose=True,
-            sigma0=0.02,
-            max_iterations=500,
-            max_unchanged_iterations=25,
-            max_evaluations=300,
+            sigma0=0.05,
+            max_iterations=150,
+            max_unchanged_iterations=150,
+            max_evaluations=150,
             **kwargs,
         )
 
@@ -474,6 +474,7 @@ if create_plot["heuristic"]:
     parameter_fig.data = []
     parameter_fig.add_traces(parameter_traces)
     parameter_fig = plotly.subplots.make_subplots(figure=parameter_fig, rows=2, cols=1)
+    parameter_fig.update_layout(yaxis1=dict(range=[1e-14, 10e-14]))
     parameter_fig.show()
     parameter_fig.write_image("joss/figures/heuristic_parameters.png")
 

### 2024-11-29 pybop/__init__.py
Unit Hypercube Transformation (#566)
* feat: adds unit hypercube transformation w/ tests

* fix: incorrect application of gradient transformation

* adds changlog entry

* fix: update prior bounds for mcmc integration tests

* tests: adds check for incorrect UnitHyperCube bounds

* fix: correctly apply t_interp, restore spm_descent.py

* tests: adds addition transformations to integration tests

* Update tests/integration/test_transformation.py

@@ -64,6 +64,7 @@ from .transformation.transformations import (
     ScaledTransformation,
     LogTransformation,
     ComposedTransformation,
+    UnitHyperCube,
 )
 
 #

### 2024-11-29 joss/image_combine.sh
adds numerical impedance predictions, figs, citation. updts nyquist grids,

@@ -6,6 +6,8 @@ convert +append figures/convergence_minimising.png figures/convergence_maximisin
 
 convert +append figures/simulation.png figures/landscape.png figures/joss/sim-landscape.png
 
+convert +append figures/impedance_spectrum.png figures/impedance_contour.png figures/joss/impedance.png
+
 convert +append figures/contour_gradient_0.png figures/contour_gradient_1.png figures/contour_gradient_2.png figures/contour_gradient_3.png figures/joss/contour_gradient.png
 
 convert +append figures/contour_evolution_0.png figures/contour_evolution_1.png figures/contour_evolution_2.png figures/contour_evolution_3.png figures/joss/contour_evolution.png

### 2024-11-29 joss/param_plots.py

@@ -15,13 +15,14 @@ np.random.seed(8)
 # Choose which plots to show and save
 create_plot = {}
 create_plot["simulation"] = False
-create_plot["landscape"] = False
+create_plot["landscape"] = True
 create_plot["minimising"] = False
 create_plot["maximising"] = False
 create_plot["gradient"] = False
 create_plot["evolution"] = False
-create_plot["heuristic"] = True
+create_plot["heuristic"] = False
 create_plot["posteriors"] = False
+create_plot["eis"] = True
 
 
 # Parameter set and model definition
@@ -127,7 +128,7 @@ if create_plot["landscape"]:
                 color="white",
                 line_color="black",
                 line_width=1,
-                size=14,
+                size=16,
                 showscale=False,
             ),
             name="Initial values",
@@ -143,7 +144,7 @@ if create_plot["landscape"]:
                 color="black",
                 line_color="white",
                 line_width=1,
-                size=14,
+                size=16,
                 showscale=False,
             ),
             name="True values",
@@ -607,3 +608,101 @@ if create_plot["posteriors"]:
     plt.tight_layout()
     plt.savefig("joss/figures/joss/posteriors.png")
     plt.show()
+
+
+if create_plot["eis"]:
+
+    def noise(sigma, values):
+        # Generate real part noise
+        real_noise = np.random.normal(0, sigma, values)
+
+        # Generate imaginary part noise
+        imag_noise = np.random.normal(0, sigma, values)
+
+        # Combine them into a complex noise
+        return real_noise + 1j * imag_noise
+
+    var = pybamm.standard_spatial_vars
+    var_pts = {var.x_n: 10, var.x_s: 10, var.x_p: 10, var.r_n: 30, var.r_p: 30}
+
+    # Construct model
+    model = pybop.lithium_ion.SPMe(
+        parameter_set=parameter_set,
+        eis=True,
+        options={"surface form": "differential", "contact resistance": "true"},
+        var_pts=var_pts,
+    )
+
+    initial_state = {"Initial SoC": 0.05}
+    n_frequency = 35
+    sigma0 = 5e-4
+    f_eval = np.logspace(-3.75, 4, n_frequency)
+
+    # Create synthetic data for parameter inference
+    sim = model.simulateEIS(
+        inputs={
+            "Negative particle diffusivity [m2.s-1]": parameter_set[
+                "Negative particle diffusivity [m2.s-1]"
+            ],
+            "Contact resistance [Ohm]": 0.01,
+        },
[…]

### 2024-11-29 pybop/plot/nyquist.py

@@ -120,13 +120,13 @@ def nyquist(problem, problem_inputs: Inputs = None, show=True, **layout_kwargs):
             showgrid=True,
             gridwidth=1,
             gridcolor="lightgray",
-            minor=dict(showgrid=True, gridwidth=0.5, gridcolor="lightgray"),
+            zerolinecolor="lightgray",
         )
         fig.update_yaxes(
             showgrid=True,
             gridwidth=1,
             gridcolor="lightgray",
-            minor=dict(showgrid=True, gridwidth=0.5, gridcolor="lightgray"),
+            zerolinecolor="lightgray",
         )
 
         # Overwrite with user-kwargs

### 2024-11-29 joss/paper.preprint.tex
Adds Agriya w/ correct affiliation, updts tex pre-print

@@
 parameterisation},
-  pdfauthor={Brady Planden, Nicola E. Courtier, Martin Robinson, Ferran
-Brosa Planella, David A. Howey},
+  pdfauthor={Brady Planden, Nicola E. Courtier, Martin Robinson, Agriya
+Khetarpal, Ferran Brosa Planella, David A. Howey},
   pdflang={en-US},
@@
     }
-\author[2,4%
+\author[4%
+  %
+  ]{Agriya Khetarpal%
+    \,\orcidlink{0000-0002-1112-1786}\,%
+    }
+\author[2,5%
   %
@@
   }
-\affil[4]{Mathematics Institute, University of Warwick, Coventry, UK%
+\affil[4]{Quansight Labs%
+  }
+\affil[5]{Mathematics Institute, University of Warwick, Coventry, UK%
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\date{28 November 2024}
+\date{29 November 2024}
 
@@
 PyBOP's own algorithms such as Adaptive Moment Estimation with Weight
-Decay (AdamW), and Cuckoo search.
+Decay (AdamW), Gradient descent, and Cuckoo search.
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
-\caption{The cost landscape for the parameterisation problem with a root
-mean squared error cost function. \label{fig:inference-time-landscape}}
+\caption{The cost landscape for the time-series parameterisation problem
+with a root mean squared error cost function.
+\label{fig:inference-time-landscape}}
 \end{figure}
 
+As mentioned above, \texttt{PyBOP} also provides inference and
+optimisation capabilities through numerical impedance spectroscopy. This
+is based on the methods presented in the \texttt{pybamm-eis} package and
+allows fast impedance computation by inversion of the sparse PyBaMM mass
+matrix after scaling and subtraction of the auto-differentiated Jacobian
+matrix. More information about this procedure is available in
+(\citeproc{ref-pybamm-eis}{Dhoot et al., n.d.}). The
+\autoref{fig:impedance-landscape} below shows the numerical impedance
+prediction available in \texttt{PyBOP} alongside the cost landscape
+constructed for the inference task. At the time of publication,
+gradient-based optimisation and sampling methods are not available when
+using an impedance workflow.
+
+\begin{figure}
+\centering
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/impedance.png}
+\caption{The cost landscape for the frequency domain impedance
+parameterisation problem with a root mean squared error cost function at
+5\% SOC. \label{fig:impedance-landscape}}
+\end{figure}
+
+Moving forward, we will continue with identification in the time-domain.
 As gradient information is available for this problem, the choice of
@@
 
+\bibitem[\citeproctext]{ref-pybamm-eis}
+Dhoot, R., Timms, R., \& Please, C. (n.d.). \emph{PyBaMM EIS: Efficient
+linear algebra methods to determine li-ion battery behaviour} (Version
+0.1.4). \url{https://www.github.com/pybamm-team/pybamm-eis}
+
 \bibitem[\citeproctext]{ref-Doyle:1993}

### 2024-11-29 joss/paper.preprint.tex
Fix typos, update table format

@@
 (\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), in addition to the
-PyBOP's own algorithms such as Adaptive Moment Estimation with Weight
-Decay (AdamW), Gradient descent, and Cuckoo search.
+\texttt{PyBOP}'s own algorithms such as Adaptive Moment Estimation with
+Weight Decay (AdamW), Gradient descent, and Cuckoo search.
 
@@
 \texttt{PyBaMM}, with construction, parameterisation, and discretisation
-managed by PyBOP's model interface to PyBaMM. This approach provides a
-robust object construction process with a consistent interface between
-the models and optimisers. The statistical methods and optimisation
-algorithms are then constructed to interface cleanly with the forward
-model predictions. Furthermore, identifiability metrics are provided
-alongside the estimated parameters through Hessian approximation of the
-cost functions in the frequentist workflows and posterior moments in the
-Bayesian workflows.
+managed by \texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This
+approach provides a robust object construction process with a consistent
+interface between the models and optimisers. The statistical methods and
+optimisation algorithms are then constructed to interface cleanly with
+the forward model predictions. Furthermore, identifiability metrics are
+provided alongside the estimated parameters through Hessian
+approximation of the cost functions in the frequentist workflows and
+posterior moments in the Bayesian workflows.
 
@@
 performance benefits to users. Alongside construction of the forward
-model, \texttt{PyBOP}`s model class provides methods for obtaining
+model, \texttt{PyBOP}'s model class provides methods for obtaining
 sensitivities from the prediction, enabling gradient-based optimisation
@@
 Monte Carlo sampling classes, which accept the LogPosterior class and
-sample from it using Pints' based Monte Carlo algorithms at the time of
-submission. In the typical workflow, the classes in
+sample from it using \texttt{PINTS} based Monte Carlo algorithms at the
+time of submission. In the typical workflow, the classes in
 \autoref{fig:classes} are constructed in sequence.
@@
 allowing for geometric parameter inference from EIS forward model
-predictions. Furthermore, \texttt{PyBOP} builds on the JAX
+predictions. Furthermore, \texttt{PyBOP} builds on the \texttt{JAX}
 (\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
-provided by \texttt{PyBaMM} by providing JAX-based cost functions for
-automatic forward model differentiation with respect to the parameters.
-This functionality provides a performance improvement alongside an
-interface to JAX-based inference packages, such as Numpyro
-(\citeproc{ref-numpyro:2019}{Phan et al., 2019}), BlackJAX
-(\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}), and Optax
-(\citeproc{ref-optax:2020}{DeepMind et al., 2020}).
+provided by \texttt{PyBaMM} by providing \texttt{JAX}-based cost
+functions for automatic forward model differentiation with respect to
+the parameters. This functionality provides a performance improvement
+alongside an interface to JAX-based inference packages, such as
+\texttt{Numpyro} (\citeproc{ref-numpyro:2019}{Phan et al., 2019}),
+\texttt{BlackJAX} (\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}),
+and \texttt{Optax} (\citeproc{ref-optax:2020}{DeepMind et al., 2020}).
 
@@
 The currently implemented subclasses for the model, problem, and cost
-classes are listed in \autoref{tab:subclasses}. The cost functions are
-grouped by problem type, while the model and optimiser classes can be
-selected in combination with any problem-cost pair.
+classes are listed in \autoref{tab:subclasses}. The model and optimiser
+classes can be selected in combination with any problem-cost pair.
 
@@
 Single particle model (SPM) & Fitting problem & Sum squared error \\
-SPM with electrolyte (SPMe) & & Root mean squared error \\
-Doyle-Fuller-Newman (DFN) & & Minkowski \\
+SPM with electrolyte (SPMe) & Design problem & Root mean squared
+error \\
+Doyle-Fuller-Newman (DFN) & Observer & Minkowski \\
 Many particle model (MPM) & & Sum of power \\
@@
 Weppner Huggins & & Maximum a posteriori \\
-Equivalent circuit model (ECM) & Observer & Unscented Kalman filter \\
-& Design problem & Gravimetric energy density \\
-& & Volumetric energy density \\
+Equivalent circuit model (ECM) & & Volumetric energy density \\
+& & Gravimetric energy density \\
+& & Unscented Kalman filter \\
 \end{longtable}
@@
 Similarly, the current algorithms available for optimisation tasks are
-presented in \autoref{tab:optimisers}. From now on, the point-based
-parameterisation and design optimisation tasks will simply be referred
-to as optimisation tasks. This simplification can be justified by
-examining \autoref{eqn:parameterisation} and \autoref{eqn:design} and
-confirming that deterministic parameterisation can be viewed as an
-optimisation task to minimise a distance-based cost function.
+presented in \autoref{tab:optimisers}. It should be noted that
+\texttt{SciPy} minimize has gradient and non-gradient methods. From now
+on, the point-based parameterisation and design optimisation tasks will
+simply be referred to as optimisation tasks. This simplification can be
+justified by examining \autoref{eqn:parameterisation} and
+\autoref{eqn:design} and confirming that deterministic parameterisation
+can be viewed as an optimisation task to minimise a distance-based cost
+function.
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4062}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2604}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4722}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2963}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2315}}@{}}
 \caption{The currently supported optimisation algorithms classified by
-candidate solution type, including gradient information. (*) Scipy
-minimize has gradient and non-gradient methods.
+candidate solution type, including gradient information.
 \label{tab:optimisers}}\tabularnewline
@@
 Gradient descent & Separable natural (sNES) & Cuckoo search \\
-SciPy minimize (*) & SciPy differential evolution & \\
+SciPy minimize & SciPy differential evolution & \\
 & & \\
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2586}}@{}}
-\caption{Sampling methods supported by PyBOP, classified according to
-the proposed method. \label{tab:samplers}}\tabularnewline
+\caption{Sampling methods supported by \texttt{PyBOP}, classified
+according to the proposed method. \label{tab:samplers}}\tabularnewline
 \toprule\noalign{}
@@
 \endlastfoot
-Monomial Gamma & Delayed Rejection Adaptive & Rank Shrinking &
+Monomial Gamma & Delayed Rejection Adaptive & Rank---Shrinking &
 Differential Evolution & Metropolis Random Walk \\
@@
 This parameterisation task can also be approached from a Bayesian
-perspective, which we will present below using PyBOP's sampler methods.
-The optimisation equation presented in equation
+perspective, which we will present below using \texttt{PyBOP}'s sampler
+methods. The optimisation equation presented in equation
 \autoref{eqn:parameterisation} does not represent the Bayesian parameter
@@
 conventional methods based on random sampling with rejection criteria
-(\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}). PyBOP offers
-a sampling class that provides an interface to these samplers, which are
-supported by the Probabilistic Inference of Noise Time-Series (PINTS)
-package. \autoref{fig:posteriors} below shows the sampled posterior for
-the synthetic workflow described above, using an adaptive
-covariance-based sampler, Haario Bardenet
+(\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}).
+\texttt{PyBOP} offers a sampling class that provides an interface to
+these samplers, which are supported by the Probabilistic Inference of
+Noise Time-Series (\texttt{PINTS}) package. \autoref{fig:posteriors}
+below shows the sampled posterior for the synthetic workflow described
+above, using an adaptive covariance-based sampler, Haario Bardenet
 (\citeproc{ref-Haario:2001}{Haario et al., 2001}).

### 2024-11-29 pybop/models/lithium_ion/basic_SPMe.py
Fix electrolyte scaling

@@ -543,7 +543,7 @@ def convert_physical_to_grouped_parameters(parameter_set):
 
     # Grouped parameters
     Q_meas = (Q_meas_n + Q_meas_p) / 2
-    Q_e = F * ce0 * L * A
+    Q_e = F * epsilon_sep * ce0 * L * A
 
     zeta_p = epsilon_p / epsilon_sep
     zeta_n = epsilon_n / epsilon_sep
@@ -551,9 +551,9 @@ def convert_physical_to_grouped_parameters(parameter_set):
     tau_d_p = R_p**2 / D_p
     tau_d_n = R_n**2 / D_n
 
-    tau_e_p = L**2 / (epsilon_p**b_p * De)
-    tau_e_n = L**2 / (epsilon_n**b_n * De)
-    tau_e_sep = L**2 / (epsilon_sep**b_sep * De)
+    tau_e_p = epsilon_sep * L**2 / (epsilon_p**b_p * De)
+    tau_e_n = epsilon_sep * L**2 / (epsilon_n**b_n * De)
+    tau_e_sep = epsilon_sep * L**2 / (epsilon_sep**b_sep * De)
 
     tau_ct_p = F * R_p / (m_p * np.sqrt(ce0))
     tau_ct_n = F * R_n / (m_n * np.sqrt(ce0))

### 2024-11-30 scripts/ci/build_matrix.sh
Sort the releases before saving to a variable

@@ -12,7 +12,7 @@
 python_version=("3.9" "3.10" "3.11" "3.12")
 os=("ubuntu-latest" "windows-latest" "macos-13" "macos-14")
 # This command fetches the last PyBaMM version excluding release candidates from PyPI
-pybamm_version=($(curl -s https://pypi.org/pypi/pybamm/json | jq -r '.releases | keys[]' | grep -v rc | tail -n 1 | paste -sd " " -))
+pybamm_version=($(curl -s https://pypi.org/pypi/pybamm/json | jq -r '.releases | keys[]' | grep -v rc | sort -V | tail -n 1 | paste -sd " " -))
 
 # This command fetches the last PyBaMM versions including release candidates from PyPI
 #pybamm_version=($(curl -s https://pypi.org/pypi/pybamm/json | jq -r '.releases | keys[]' | tail -n 1 | paste -sd " " -))

### 2024-12-04 pybop/samplers/mcmc_summary.py
refactor: ESS method, adds tests

@@ -198,32 +198,43 @@ class PosteriorSummary:
 
     def effective_sample_size(self, mixed_chains=False):
         """
-        Computes the effective sample size for each parameter in each chain.
+        Computes the effective sample size (ESS) for each parameter in each chain.
 
         Parameters
-        -----------
+        ----------
         mixed_chains : bool, optional
-            If true, the effective sample size is computed for all samplers mixed
-            into a single chain.
-        """
+            If True, the ESS is computed for all samplers mixed into a single chain.
+            Defaults to False.
 
+        Returns
+        -------
+        list
+            A list of effective sample sizes for each parameter in each chain,
+            or for the mixed chain if `mixed_chains` is True.
+
+        Raises
+        ------
+        ValueError
+            If there are fewer than two samples in the data.
+        """
         if self.all_samples.shape[0] < 2:
             raise ValueError("At least two samples must be given.")
 
-        ess = []
-        if mixed_chains:
+        def compute_ess(samples):
+            """Helper function to compute the ESS for a single set of samples."""
+            ess = []
             for j in range(self.num_parameters):
-                rho = self.autocorrelation(self.all_samples[:, j])
+                rho = self.autocorrelation(samples[:, j])
                 T = self._autocorrelate_negative(rho)
-                ess.append(len(self.all_samples[:, j]) / (1 + 2 * rho[:T].sum()))
+                ess.append(len(samples[:, j]) / (1 + 2 * rho[:T].sum()))
             return ess
 
-        for _, chain in enumerate(self.chains):
-            for j in range(self.num_parameters):
-                rho = self.autocorrelation(chain[:, j])
-                T = self._autocorrelate_negative(rho)
-                ess.append(len(chain[:, j]) / (1 + 2 * rho[:T].sum()))
+        if mixed_chains:
+            return compute_ess(self.all_samples)
 
+        ess = []
+        for chain in self.chains:
+            ess.extend(compute_ess(chain))
         return ess
 
     def rhat(self):

### 2024-12-04 pybop/models/lithium_ion/basic_SPMe.py
Adding Battery voltage to the variables list for time domain simulations.

@@ -54,11 +54,13 @@ class BaseGroupedSPMe(pybamm_lithium_ion.BaseModel):
         include_double_layer = self.options["surface form"] == "differential"
 
         pybamm.citations.register("Chen2020")  # for the OCPs
-        pybamm.citations.register("""
+        pybamm.citations.register(
+            """
             @article{HallemansPreprint,
             title={{Hallemans Preprint}},
             }
-        """)
+        """
+        )
 
         ######################
         # Variables
@@ -373,6 +375,7 @@ class BaseGroupedSPMe(pybamm_lithium_ion.BaseModel):
             "Discharge capacity [A.h]": Q,
             "Throughput capacity [A.h]": Qt,
             "Voltage [V]": V,
+            "Battery voltage [V]": V,
             "Open-circuit voltage [V]": U_p - U_n,
         }
 

### 2024-12-04 examples/scripts/comparison_examples/grouped_SPMe_experiment.py
Create grouped_SPMe_experiment.py

@@ -1,0 +1,38 @@
+import pybop
+from pybop.models.lithium_ion.basic_SPMe import convert_physical_to_grouped_parameters
+
+
+# Define parameter set
+parameter_set = pybop.ParameterSet.pybamm("Chen2020")
+parameter_set["Electrolyte diffusivity [m2.s-1]"] = 1.769e-10
+parameter_set["Electrolyte conductivity [S.m-1]"] = 1e16
+parameter_set["Negative electrode conductivity [S.m-1]"] = 1e16
+parameter_set["Positive electrode conductivity [S.m-1]"] = 1e16
+grouped_parameters = convert_physical_to_grouped_parameters(parameter_set)
+grouped_parameters["Series resistance [Ohm]"] = 0.01
+
+## Create model
+var_pts = {"x_n": 20, "x_s": 20, "x_p": 20, "r_n": 20, "r_p": 20}
+model_options = {"surface form": "differential", "contact resistance": "true"}
+model = pybop.lithium_ion.GroupedSPMe(
+    parameter_set=grouped_parameters, var_pts=var_pts, options=model_options
+)
+
+## Test model in the time domain
+model.set_initial_state({"Initial SoC": 0.9})
+experiment = pybop.Experiment(
+    [
+        "Rest for 10 minutes (10 seconds period)",
+        # "Discharge at 1C until 2.5 V (10 seconds period)",
+        "Rest for 20 minutes (10 seconds period)",
+    ],
+)
+simulation = model.predict(experiment=experiment)
+dataset = pybop.Dataset(
+    {
+        "Time [s]": simulation["Time [s]"].data,
+        "Current function [A]": simulation["Current [A]"].data,
+        "Voltage [V]": simulation["Voltage [V]"].data,
+    }
+)
+pybop.plot.dataset(dataset, signal=["Voltage [V]", "Current function [A]"])

### 2024-12-04 joss/paper.preprint.tex
fixes: internal review feedback, minus figures

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\date{29 November 2024}
+\date{04 December 2024}
 
@@
 The Python Battery Optimisation and Parameterisation (\texttt{PyBOP})
-package provides a set of methods for the parameterisation and
-optimisation of battery models, offering both Bayesian and frequentist
-approaches with example workflows to assist the user. \texttt{PyBOP} has
-been developed to enable parameter identification of various battery
-models, including the electrochemical and equivalent circuit models
-provided by the popular open-source \texttt{PyBaMM} package
-(\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}).
-
-Similarly, \texttt{PyBOP} can be used for parameter design optimisation
-under user-defined operating conditions across a variety of model
-structures. \texttt{PyBOP} allows battery model parameterisation using a
-range of methods with diagnostics on the performance and convergence of
-the identified or optimised parameters. The identified parameters can be
-used for prediction, on-line control and design optimisation, all of
-which support improved battery utilisation and development.
+package provides methods for parameterising and optimising battery
+models, offering both deterministic and stochastic approaches with
+example workflows to assist users. \texttt{PyBOP} enables parameter
+identification of various battery models, including the electrochemical
+and equivalent circuit models provided by the popular open-source
+\texttt{PyBaMM} package (\citeproc{ref-Sulzer:2021}{Sulzer et al.,
+2021}).
+
+Similarly, \texttt{PyBOP} can be used for design optimisation under
+user-defined operating conditions across a variety of model structures
+and design goals. \texttt{PyBOP} facilitates model parameterisation
+using a range of methods, with diagnostics for examining the performance
+and convergence of the resulting parameters. Identified parameters can
+be used for prediction, on-line control, and design optimisation,
+supporting improved battery utilisation and development.
 
@@
 
-\texttt{PyBOP} is a Python package designed to provide a user-friendly,
-object-oriented interface for optimising battery models. \texttt{PyBOP}
-leverages the open-source \texttt{PyBaMM} package
+\texttt{PyBOP} is a Python package that provides a user-friendly,
+object-oriented interface for optimising battery model parameters.
+\texttt{PyBOP} leverages the open-source \texttt{PyBaMM} package
 (\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
-of these battery models. \texttt{PyBOP} is intended to serve a broad
-audience of students, engineers, and researchers in both academia and
-the battery industry by enabling the use of predictive battery models
-where previously this was not possible. \texttt{PyBOP} emphasises clear
-and informative diagnostics and workflows for users of varying
-expertise, by providing advanced optimisation and sampling algorithms.
-These methods are provided through interfaces to \texttt{PINTS}
+battery models. \texttt{PyBOP} is intended to serve a broad audience
+including students, engineers, and researchers in academia and industry,
+enabling the use of battery models for performance prediction where
+previously this was not possible without specialised knowledge of both
+modelling, parameter inference and software development. \texttt{PyBOP}
+emphasises clear and informative diagnostics and workflows to support
+users with varying expertise, and provides access to a wide range of
+different optimisation and sampling algorithms. These methods are
+provided through interfaces to \texttt{PINTS}
 (\citeproc{ref-Clerx:2019}{Clerx et al., 2019}), \texttt{SciPy}
-(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), in addition to the
-\texttt{PyBOP}'s own algorithms such as Adaptive Moment Estimation with
-Weight Decay (AdamW), Gradient descent, and Cuckoo search.
-
-\texttt{PyBOP} supports the Battery Parameter eXchange (BPX) standard
+(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), and \texttt{PyBOP}'s
+own algorithms such as adaptive moment estimation with weight decay
+(AdamW), gradient descent, and Cuckoo search.
+
+\texttt{PyBOP} supports the battery parameter exchange (BPX) standard
 (\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing battery
-parameter sets. As these parameter sets are costly to obtain due to: the
-equipment and time required for characterisation experiments, the need
-for battery domain knowledge, and the computational cost of parameter
-estimation. \texttt{PyBOP} reduces these costs by providing fast
-computational estimation with parameter set interoperability.
+parameter sets. These parameter sets are typically costly to obtain due
+to the specialised equipment and time required for characterisation
+experiments, the need for battery domain knowledge, and the
+computational cost of parameter estimation. \texttt{PyBOP} reduces these
+costs by providing fast parameter estimation, standardised workflows,
+and parameter set interoperability (via BPX).
 
@@
 pack simulation (\citeproc{ref-Tranter2022}{Tranter et al., 2022}) and
-\texttt{pybamm-eis} for numerical impedance spectroscopy, as the
-identified parameters from \texttt{PyBOP} are easily exportable.
+\texttt{pybamm-eis} for numerical impedance spectroscopy, since the
+identified parameters from \texttt{PyBOP} are easily exportable to these
+other packages.
 
@@
 
-\texttt{PyBOP} has a layered data structure designed to compute and
-process the forward model predictions and to package the necessary
-information for the optimisation and sampling algorithms. The forward
-model is solved using the popular battery modelling package,
-\texttt{PyBaMM}, with construction, parameterisation, and discretisation
-managed by \texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This
-approach provides a robust object construction process with a consistent
-interface between the models and optimisers. The statistical methods and
-optimisation algorithms are then constructed to interface cleanly with
-the forward model predictions. Furthermore, identifiability metrics are
-provided alongside the estimated parameters through Hessian
-approximation of the cost functions in the frequentist workflows and
-posterior moments in the Bayesian workflows.
+\texttt{PyBOP} has a layered structure designed to compute forward model
+predictions, process the results, and package the necessary information
+for the optimisation and sampling algorithms. The forward model is
+solved using the popular battery modelling package, \texttt{PyBaMM},
+with construction, parameterisation, and discretisation managed by
+\texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This provides a
+robust object construction process with a consistent interface between
+models and optimisers. The statistical methods and optimisation
+algorithms are constructed to interface cleanly with the forward model
+predictions. Furthermore, identifiability metrics are provided with the
+estimated parameters (through Hessian approximation of the cost
+functions in frequentist workflows, and posterior distributions in
+Bayesian workflows).
 
@@
 visualisation of the general workflow for parameterisation and
-optimisation \label{fig:high-level}}
+optimisation. \label{fig:high-level}}
 \end{figure}
@@
 sensitivities from the prediction, enabling gradient-based optimisation
-algorithms. This prediction, along with it's corresponding
-sensitivities, is provided to the problem class for processing and
-exception control. A standardised data structure is then provided to the
-cost classes, providing a distance, design, or likelihood-based metric
-for optimisation. For deterministic optimisation, the optimisers
-minimise the corresponding cost function or the negative log-likelihood
-if a likelihood class is provided. Bayesian inference is provided by
-Monte Carlo sampling classes, which accept the LogPosterior class and
-sample from it using \texttt{PINTS} based Monte Carlo algorithms at the
-time of submission. In the typical workflow, the classes in
+algorithms. This prediction, along with its corresponding sensitivities,
+is provided to the problem class for processing and exception control. A
+standardised data structure is then provided to the cost classes, which
+then computes a distance, design, or likelihood-based metric for
+optimisation. For deterministic optimisation, the optimisers minimise
+the corresponding cost function or the negative log-likelihood if a
+likelihood class is provided. Bayesian inference is provided by Monte
+Carlo sampling classes, which accept the LogPosterior class and sample
+from it using \texttt{PINTS} based Monte Carlo algorithms at the time of
+submission. In the typical workflow, the classes in
 \autoref{fig:classes} are constructed in sequence.
@@
 specialised inference and optimisation processes. One such instance is
-numerical electrochemical impedance spectroscopy predictions by
-discretising the forward model into sparse mass matrix form with
+parameter inference from EIS forward model predicitions. PyBOP
+discretises the EIS forward model inso sparse mass matrix form with
 accompanying auto-differentiated Jacobian. These objects are then
@@
 functions for automatic forward model differentiation with respect to
-the parameters. This functionality provides a performance improvement
-alongside an interface to JAX-based inference packages, such as
-\texttt{Numpyro} (\citeproc{ref-numpyro:2019}{Phan et al., 2019}),
-\texttt{BlackJAX} (\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}),
-and \texttt{Optax} (\citeproc{ref-optax:2020}{DeepMind et al., 2020}).
+the parameters. This functionality provides a performance improvement,
+and allows users to use many other JAX-based inference packages in order
+to optimise their cost function, such as \texttt{Numpyro}
+(\citeproc{ref-numpyro:2019}{Phan et al., 2019}), \texttt{BlackJAX}
+(\citeproc{ref-blackjax:2024}{Cabezas et al., 2024}), and \texttt{Optax}
+(\citeproc{ref-optax:2020}{DeepMind et al., 2020}).
 
@@
 
-As discussed above, \texttt{PyBOP} provides Bayesian inference methods
-such as maximum a posteriori (MAP) alongside the point-based methods in
-\autoref{tab:subclasses}; however, for a full Bayesian framework, Monte
-Carlo sampling is implemented within \texttt{PyBOP}. These methods
-construct a posterior distribution on the inference parameters, which
-can be used for uncertainty and practical identifiability. The
-individual sampler classes are currently composed within \texttt{PyBOP}
-from the \texttt{PINTS} library, with a base sampling class implemented
-for interoperability and direct integration with the \texttt{PyBOP}
-model, problem, and likelihood classes. The currently supported samplers
-are listed in \autoref{tab:samplers}.
+As discussed above, \texttt{PyBOP} provides both point-based
+optimisation estimates \autoref{tab:subclasses} in addition to Monte
+Carlo sampling routines in order to recover distributions of likely
+parameter values. The Monte Carlco methods construct a posterior
+distribution on the inference parameters, which can be used for
+uncertainty and practical identifiability. The individual sampler
+classes are currently composed within \texttt{PyBOP} from the
+\texttt{PINTS} library, with a base sampling class implemented for
+interoperability and direct integration with the \texttt{PyBOP} model,
+problem, and likelihood classes. The currently supported samplers are
+listed in \autoref{tab:samplers}.
 
@@
 
-In general, battery models can be written in the form of a
-differential-algebraic system of equations: \begin{equation}
-\frac{\mathrm{d} \mathbf{x}}{\mathrm{d} t} = f(t,\mathbf{x},\mathbf{y},\mathbf{u}(t),\mathbf{\theta}),
+In general, battery models (after spatial discretisation) can be written
+in the form of a differential-algebraic system of equations:
+\begin{equation}
+\frac{\mathrm{d} \mathbf{x}}{\mathrm{d} t} = f(t,\mathbf{x},\mathbf{\theta}),
 \label{dynamics}
 \end{equation} \begin{equation}
-\mathbf{y}(t) = g(t,\mathbf{x},\mathbf{y},\mathbf{u}(t),\mathbf{\theta}),
+0 = g(t, \mathbf{x}, \mathbf{\theta})
+\label{algebraic}
+\end{equation} \begin{equation}
+\mathbf{y}(t) = h(t,\mathbf{x}),
 \label{output}
@@
 Here, \(t\) is time, \(\mathbf{x}(t)\) are the (spatially discretised)
-states, \(\mathbf{y}(t)\) are the outputs (e.g.~the terminal voltage),
-\(\mathbf{u}(t)\) are the inputs (e.g.~the applied current) and
-\(\mathbf{\theta}\) are the unknown parameters.
+states, \(\mathbf{y}(t)\) are the outputs (e.g.~the terminal voltage)
+and \(\mathbf{\theta}\) are the unknown parameters.
 
@@
 electrode theory and its reduced-order variants including the single
-particle model (SPM) (\citeproc{ref-Planella:2022}{Planella et al.,
-2022}), and the multi-species, multi-reaction (MSMR) model
+particle model (SPM) (\citeproc{ref-Planella:2022}{Brosa Planella et
+al., 2022}), and the multi-species, multi-reaction (MSMR) model
 (\citeproc{ref-Verbrugge:2017}{Verbrugge et al., 2017}). Simplified
@@
 parameterisation often requires a stepwise identification of smaller
-sets of parameters from a variety of different data sets
+sets of parameters from a variety of excitations and different data sets
 (\citeproc{ref-Chen:2020}{Chen et al., 2020};
 \citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
-et al., 2023}) and excitations.
+et al., 2023}).
 
@@
 
-where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost (or
-likelihood) function that quantifies the agreement between the model and
-a sequence of observations \((\mathbf{y}_i)\) measured at times \(t_i\).
-For gradient-based optimisers, the Jacobian of the cost function with
-respect to the unknown parameters,
+where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
+function that quantifies the agreement between the model,
+\(\mathbf{y}(t_i)\) and a sequence of measured observations
+\(\hat{\mathbf{y}_i}\) measured at times \(t_i\). Within the
+\texttt{PyBOP} framework, the \texttt{FittingProblem} class packages the
+model output along with the measured observations, both of which are
+then passed to the \texttt{FittingCost} classes for the computation of
+the specific cost function. For gradient-based optimisers, the Jacobian
+of the cost function with respect to the unknown parameters,
 \((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
@@
 assumed known, although such an assumption is not generally necessary.
-The underlying cost landscape explored by the optimiser is shown in
-\autoref{fig:inference-time-landscape} (right).
+The \texttt{PyBOP} repository contains multiple
+\href{https://github.com/pybop-team/PyBOP/tree/develop/examples/notebooks}{example
+notebooks} that follow a similar inference process for readers that
+would like further information. The underlying cost landscape to be
+explored by the optimiser is shown in
+\autoref{fig:inference-time-landscape} (right) with the initial position
+denoted alongside the known system parameters for this synthetic
+inference task. In general, the true parameters are not known a priori.
 
@@
 
-Moving forward, we will continue with identification in the time-domain.
-As gradient information is available for this problem, the choice of
-distance-based cost function and optimiser is not constrained. Due to
-the different magnitudes of the two parameters, we apply the logarithmic
-parameter transformation offered by \texttt{PyBOP}. This transforms the
-optimisers search space of the optimiser to allow for a common step size
-between the parameters, which is generally is not required, but improves
-convergence in this problem. As a demonstration of the parameterisation
-capabilities of \texttt{PyBOP}, \autoref{fig:convergence-min-max} (left)
-shows the rate of convergence for each of the distance-minimising cost
-functions, while \autoref{fig:convergence-min-max} (right) shows
-analogous results for maximising a likelihood. The optimisation is
-performed with SciPy Minimize using the gradient-based L-BFGS-B method.
+To avoid over complicating this example, we will continue with
+identification in the time-domain; however, in general these two
+forward-model prediction methods can be combined for improved system
+excitation. As gradient information is available for this problem, the
+choice of distance-based cost function and optimiser is not constrained.
+Due to the different magnitudes of the two parameters, we apply the
+logarithmic parameter transformation offered by \texttt{PyBOP}. This
+transforms the optimisers search space of the optimiser to allow for a
+common step size between the parameters, which is generally is not
+required, but improves convergence in this problem. As a demonstration
+of the parameterisation capabilities of \texttt{PyBOP},
+\autoref{fig:convergence-min-max} (left) shows the rate of convergence
+for each of the distance-minimising cost functions, while
+\autoref{fig:convergence-min-max} (right) shows analogous results for
+maximising a likelihood. The optimisation is performed with SciPy
+Minimize using the gradient-based L-BFGS-B method.
 
@@
 perspective, which we will present below using \texttt{PyBOP}'s sampler
-methods. The optimisation equation presented in equation
+methods. The optimisation equation presented in
 \autoref{eqn:parameterisation} does not represent the Bayesian parameter
@@
 
+\bibitem[\citeproctext]{ref-Planella:2022}
+Brosa Planella, F., Ai, W., Boyce, A. M., Ghosh, A., Korotkin, I., Sahu,
+S., Sulzer, V., Timms, R., Tranter, T. G., Zyskin, M., Cooper, S. J.,
+Edge, J. S., Foster, J. M., Marinescu, M., Wu, B., \& Richardson, G.
+(2022). {A Continuum of Physics-Based Lithium-Ion Battery Models
+Reviewed}. \emph{Progress in Energy}, \emph{4}(4), 042003.
+\url{https://doi.org/10.1088/2516-1083/ac7d31}
+
 \bibitem[\citeproctext]{ref-blackjax:2024}
@@
 \emph{arXiv Preprint arXiv:1912.11554}.
-
-\bibitem[\citeproctext]{ref-Planella:2022}
-Planella, F. B., Ai, W., Boyce, A. M., Ghosh, A., Korotkin, I., Sahu,
-S., Sulzer, V., Timms, R., Tranter, T. G., Zyskin, M., Cooper, S. J.,
-Edge, J. S., Foster, J. M., Marinescu, M., Wu, B., \& Richardson, G.
-(2022). {A Continuum of Physics-Based Lithium-Ion Battery Models
-Reviewed}. \emph{Progress in Energy}, \emph{4}(4), 042003.
-\url{https://doi.org/10.1088/2516-1083/ac7d31}
 

### 2024-12-04 pybop/__init__.py
refactor: update init structure

@@ -120,7 +120,7 @@ from .costs._weighted_cost import WeightedCost
 #
 # Experimental
 #
-from .experimental import BaseJaxCost, JaxSumSquaredError, JaxLogNormalLikelihood, JaxGaussianLogLikelihoodKnownSigma
+from .experimental.jax_costs import BaseJaxCost, JaxSumSquaredError, JaxLogNormalLikelihood, JaxGaussianLogLikelihoodKnownSigma
 
 #
 # Optimiser classes

### 2024-12-04 pybop/experimental/__init__.py

@@ -1,1 +1,0 @@
-from .jax_costs import BaseJaxCost, JaxLogNormalLikelihood, JaxSumSquaredError, JaxGaussianLogLikelihoodKnownSigma

### 2024-12-04 pybop/_utils.py
Add naming utility

@@ -2,6 +2,7 @@ from typing import Optional
 
 import numpy as np
 import pybamm
+import regex as re
 
 
 def is_numeric(x):
@@ -9,6 +10,15 @@ def is_numeric(x):
     Check if a variable is numeric.
     """
     return isinstance(x, (int, float, np.number))
+
+
+def add_spaces(string):
+    """
+    Return the class name as a string with spaces before each new capitalised word.
+    """
+    re_outer = re.compile(r"([^A-Z ])([A-Z])")
+    re_inner = re.compile(r"(?<!^)([A-Z])([^A-Z])")
+    return re_outer.sub(r"\1 \2", re_inner.sub(r" \1\2", string))
 
 
 class SymbolReplacer:

### 2024-12-05 pybop/optimisers/scipy_optimisers.py
Store original cost

@@ -92,8 +92,7 @@ class BaseSciPyOptimiser(BaseOptimiser):
             if self._transformation
             else result.x,
             cost=self.cost,
-            final_cost=self.cost(result.x, apply_transform=True)
-            * (1 if self.minimising else -1),
+            final_cost=self.cost(result.x, apply_transform=True),
             n_iterations=nit,
             scipy_result=result,
             optim=self,
@@ -213,9 +212,9 @@ class SciPyMinimize(BaseSciPyOptimiser):
 
         L, dl = self.cost(x, calculate_grad=True, apply_transform=True)
         self.log_update(x=[x], cost=L)
-        scaled_L = L / self._cost0 * (1 if self.minimising else -1)
-        scaled_dl = dl / self._cost0 * (1 if self.minimising else -1)
-        return (scaled_L, scaled_dl)
+        scaled_L = L / self._cost0
+        scaled_dl = dl / self._cost0
+        return (scaled_L, scaled_dl) if self.minimising else (-scaled_L, -scaled_dl)
 
     def _run_optimiser(self):
         """
@@ -238,15 +237,16 @@ class SciPyMinimize(BaseSciPyOptimiser):
             """
             if isinstance(intermediate_result, OptimizeResult):
                 x_best = intermediate_result.x
-                cost_best = intermediate_result.fun * (1 if self.minimising else -1)
+                cost_best = (
+                    intermediate_result.fun
+                    * self._cost0
+                    * (1 if self.minimising else -1)
+                )
             else:
                 x_best = intermediate_result
                 cost_best = self.cost(x_best, apply_transform=True)
 
-            self.log_update(
-                x_best=x_best,
-                cost_best=cost_best * self._cost0,
-            )
+            self.log_update(x_best=x_best, cost_best=cost_best)
 
         callback = (
             base_callback

### 2024-12-05 tests/integration/test_classification.py
Update test_classification

@@ -2,7 +2,6 @@ import numpy as np
 import pytest
 
 import pybop
-from examples.standalone.cost import StandaloneCost
 from pybop._classification import classify_using_Hessian
 
 
@@ -38,7 +37,7 @@ class TestClassification:
     @pytest.fixture
     def model(self, parameters):
         parameter_set = pybop.ParameterSet(
-            json_path="examples/scripts/parameters/initial_ecm_parameters.json"
+            json_path="examples/parameters/initial_ecm_parameters.json"
         )
         parameter_set.import_parameters()
         parameter_set.params.update({"C1 [F]": 1000})
@@ -103,9 +102,11 @@ class TestClassification:
             assert message == "The optimiser has located a maximum."
 
     @pytest.mark.integration
-    def test_classify_using_Hessian_invalid(self):
-        cost = StandaloneCost()
-        optim = pybop.XNES(cost=cost)
+    def test_classify_using_Hessian_invalid(self, model, parameters, dataset):
+        parameters.remove("R0 [Ohm]")
+        problem = pybop.FittingProblem(model, parameters, dataset)
+        cost = pybop.SumSquaredError(problem)
+        optim = pybop.SNES(cost=cost)
         optim.run()
 
         with pytest.raises(

### 2024-12-09 pybop/costs/_weighted_cost.py
Apply maximise and transform in BaseOptimiser

@@ -2,7 +2,7 @@ from typing import Optional, Union
 
 import numpy as np
 
-from pybop import BaseCost, BaseLikelihood, DesignCost
+from pybop import BaseCost
 
 
 class WeightedCost(BaseCost):
@@ -32,9 +32,6 @@ class WeightedCost(BaseCost):
         self.costs = [cost for cost in costs]
         if len(set(type(cost.problem) for cost in self.costs)) > 1:
             raise TypeError("All problems must be of the same class type.")
-        self.minimising = not any(
-            isinstance(cost, (BaseLikelihood, DesignCost)) for cost in self.costs
-        )
 
         # Check if weights are provided
         if weights is not None:

### 2024-12-09 pybop/costs/base_cost.py

@@ -58,6 +58,7 @@ class BaseCost:
         inputs: Union[Inputs, list],
         calculate_grad: bool = False,
         apply_transform: bool = False,
+        minimising: bool = True,
     ) -> Union[float, tuple[float, np.ndarray]]:
         """
         This method calls the forward model via problem.evaluate(inputs),
@@ -73,6 +74,9 @@ class BaseCost:
             cost is computed.
         apply_transform : bool, optional, default=False
             If True, applies a transformation to the inputs before evaluating the model.
+        minimsing : bool, optional, default=True
+            If False, switches the sign of the cost and gradient to perform maximisation
+            instead of minimisation.
 
         Returns
         -------
@@ -110,10 +114,15 @@ class BaseCost:
                     jac = self.transformation.jacobian(inputs)
                     grad = np.matmul(grad, jac)
 
-                return cost, grad
+                return cost * (1 if minimising else -1), grad * (
+                    1 if minimising else -1
+                )
 
             y = self.problem.evaluate(self.problem.parameters.as_dict())
-        return self.compute(y, dy=dy, calculate_grad=calculate_grad)
+
+        return self.compute(y, dy=dy, calculate_grad=calculate_grad) * (
+            1 if minimising else -1
+        )
 
     def compute(self, y: dict, dy: ndarray, calculate_grad: bool = False):
         """

### 2024-12-09 pybop/optimisers/base_optimiser.py

@@ -62,7 +62,7 @@ class BaseOptimiser:
         # First set attributes to default values
         self.parameters = Parameters()
         self.x0 = optimiser_kwargs.get("x0", [])
-        self.log = dict(x=[], x_best=[], cost=[], cost_best=[], x0=[])
+        self.log = dict(x=[], x_best=[], x_search=[], x0=[], cost=[], cost_best=[])
         self.bounds = None
         self.sigma0 = 0.02
         self.verbose = True
@@ -80,9 +80,11 @@ class BaseOptimiser:
             self._transformation = self.cost.transformation
             self.set_allow_infeasible_solutions()
             if isinstance(cost, WeightedCost):
-                self.minimising = cost.minimising
-            if isinstance(cost, (BaseLikelihood, DesignCost)):
-                self.minimising = False
+                self.minimising = not isinstance(
+                    cost.costs[0], (BaseLikelihood, DesignCost)
+                )
+            else:
+                self.minimising = not isinstance(cost, (BaseLikelihood, DesignCost))
 
         else:
             try:
@@ -169,6 +171,38 @@ class BaseOptimiser:
         """
         raise NotImplementedError
 
+    def cost_call(
+        self,
+        x: Union[Inputs, list],
+        calculate_grad: bool = False,
+    ) -> Union[float, tuple[float, np.ndarray]]:
+        """
+        Call the cost function to minimise, applying any given transformation to the
+        input parameters.
+
+        Parameters
+        ----------
+        x : Inputs or list-like
+            The input parameters for which the cost and optionally the gradient
+            will be computed.
+        calculate_grad : bool, optional, default=False
+            If True, both the cost and gradient will be computed. Otherwise, only the
+            cost is computed.
+
+        Returns
+        -------
+        float or tuple
+            - If `calculate_grad` is False, returns the computed cost (float).
+            - If `calculate_grad` is True, returns a tuple containing the cost (float)
+              and the gradient (np.ndarray).
+        """
+        return self.cost(
+            x,
+            calculate_grad=calculate_grad,
+            apply_transform=True,
+            minimising=self.minimising,
+        )
+
     def run(self):
         """
         Run the optimisation and return the optimised parameters and final cost.
@@ -243,6 +277,7 @@ class BaseOptimiser:
 
         if x is not None:
             x = convert_to_list(x)
+            self.log["x_search"].extend(x)
             x = apply_transformation(x)
[…]

### 2024-12-09 pybop/optimisers/base_pints_optimiser.py

@@ -199,13 +199,7 @@ class BasePintsOptimiser(BaseOptimiser):
 
         # Choose method to evaluate
         def fun(x):
-            if self._needs_sensitivities:
-                L, dl = self.cost(x, calculate_grad=True, apply_transform=True)
-            else:
-                L = self.cost(x, apply_transform=True)
-                dl = None
-            sign = -1 if not self.minimising else 1
-            return (sign * L, sign * dl) if dl is not None else sign * L
+            return self.cost_call(x, calculate_grad=self._needs_sensitivities)
 
         # Create evaluator object
         if self._parallel:
@@ -224,7 +218,7 @@ class BasePintsOptimiser(BaseOptimiser):
         fb = fg = np.inf
 
         # Internally we always minimise! Keep a 2nd value to show the user.
-        fg_user = (fb, fg) if self.minimising else (-fb, -fg)
+        fg_user = (fb, fg)
 
         # Keep track of the last significant change
         f_sig = np.inf
@@ -245,7 +239,7 @@ class BasePintsOptimiser(BaseOptimiser):
                 # Update the scores
                 fb = self.optimiser.f_best()
                 fg = self.optimiser.f_guessed()
-                fg_user = (fb, fg) if self.minimising else (-fb, -fg)
+                fg_user = (fb, fg)
 
                 # Check for significant changes against the absolute and relative tolerance
                 f_new = fg if self._use_f_guessed else fb
@@ -264,8 +258,8 @@ class BasePintsOptimiser(BaseOptimiser):
                 self.log_update(
                     x=xs,
                     x_best=self.optimiser.x_best(),
-                    cost=_fs if self.minimising else [-x for x in _fs],
-                    cost_best=fb if self.minimising else -fb,
+                    cost=_fs,
+                    cost_best=fb,
                 )
 
                 # Check stopping criteria:
@@ -331,11 +325,8 @@ class BasePintsOptimiser(BaseOptimiser):
             print("Current score: " + str(fg_user))
             print("Current position:")
 
-            # Show current parameters
-            x_user = self.optimiser.x_guessed()
-            if self._transformation:
-                x_user = self._transformation.to_model(x_user)
-            for p in x_user:
+            # Show current parameters (with any transformation applied)
+            for p in self.optimiser.x_guessed():
                 print(PintsStrFloat(p))
             print("-" * 40)
             raise
@@ -356,16 +347,11 @@ class BasePintsOptimiser(BaseOptimiser):
             x = self.optimiser.x_best()
[…]

### 2024-12-09 tests/unit/test_optimisation.py

@@ -144,9 +144,15 @@ class TestOptimisation:
     @pytest.mark.unit
     def test_optimiser_kwargs(self, cost, optimiser):
         def assert_log_update(optim):
-            optim.log_update(x=[0.7], cost=[0.01])
-            assert optim.log["x"][-1] == 0.7
-            assert optim.log["cost"][-1] == 0.01
+            x_search = 0.7
+            optim.log_update(x=[x_search], cost=[0.01])
+            assert optim.log["x_search"][-1] == x_search
+            assert optim.log["cost"][-1] == 0.01 * (1 if optim.minimising else -1)
+            assert (
+                optim.log["x"][-1] == optim._transformation.to_model(x_search)
+                if optim._transformation
+                else x_search
+            )
 
         def check_max_iterations(optim):
             results = optim.run()
@@ -619,7 +625,8 @@ class TestOptimisation:
     @pytest.mark.unit
     def test_optimisation_results(self, cost):
         # Construct OptimisationResult
-        results = pybop.OptimisationResult(x=[1e-3], cost=cost, n_iterations=1)
+        optim = pybop.Optimisation(cost=cost)
+        results = pybop.OptimisationResult(optim=optim, x=[1e-3], n_iterations=1)
 
         # Asserts
         assert results.x[0] == 1e-3
@@ -630,4 +637,4 @@ class TestOptimisation:
         with pytest.raises(
             ValueError, match="Optimised parameters do not produce a finite cost value"
         ):
-            pybop.OptimisationResult(x=[1e-5], cost=cost, n_iterations=1)
+            pybop.OptimisationResult(optim=optim, x=[1e-5], n_iterations=1)

### 2024-12-09 tests/unit/test_plots.py
Update test_plots.py

@@ -147,7 +147,7 @@ class TestPlots:
         pybop.plot.contour(optim, gradient=True, steps=5)
 
         # Plot voronoi
-        pybop.plot.surface(optim)
+        pybop.plot.surface(optim, normalised_distance_metric=False)
 
         # Plot voronoi w/ bounds
         pybop.plot.surface(optim, bounds=bounds)

### 2024-12-09 joss/paper.preprint.tex
feat: updates figures, contour plot into plotly subplots.

@@
 The Python Battery Optimisation and Parameterisation (\texttt{PyBOP})
-package provides methods for parameterising and optimising battery
-models, offering both deterministic and stochastic approaches with
-example workflows to assist users. \texttt{PyBOP} enables parameter
-identification of various battery models, including the electrochemical
-and equivalent circuit models provided by the popular open-source
-\texttt{PyBaMM} package (\citeproc{ref-Sulzer:2021}{Sulzer et al.,
-2021}).
-
-Similarly, \texttt{PyBOP} can be used for design optimisation under
-user-defined operating conditions across a variety of model structures
-and design goals. \texttt{PyBOP} facilitates model parameterisation
-using a range of methods, with diagnostics for examining the performance
-and convergence of the resulting parameters. Identified parameters can
-be used for prediction, on-line control, and design optimisation,
-supporting improved battery utilisation and development.
+package provides methods for estimating and optimising battery model
+parameters, offering both deterministic and stochastic approaches with
+example workflows to assist users.
+
+\texttt{PyBOP} enables parameter identification from data for various
+battery models, including both the electrochemical and equivalent
+circuit models provided by the popular open-source \texttt{PyBaMM}
+package (\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}). Using the
+same workflow, \texttt{PyBOP} can also be used for design optimisation
+under user-defined operating conditions across a variety of model
+structures and design goals. \texttt{PyBOP} facilitates optimisation
+using a range of methods, with diagnostics for examining optimiser
+performance and convergence of the cost and corresponding parameters.
+Identified parameters can be used for prediction, on-line control, and
+design optimisation to accelerate research and development and improve
+battery utilisation.
 
@@
 (\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
-battery models. \texttt{PyBOP} is intended to serve a broad audience
+battery models. Together, these packages serve a broad audience
 including students, engineers, and researchers in academia and industry,
-enabling the use of battery models for performance prediction where
-previously this was not possible without specialised knowledge of both
-modelling, parameter inference and software development. \texttt{PyBOP}
-emphasises clear and informative diagnostics and workflows to support
-users with varying expertise, and provides access to a wide range of
-different optimisation and sampling algorithms. These methods are
-provided through interfaces to \texttt{PINTS}
-(\citeproc{ref-Clerx:2019}{Clerx et al., 2019}), \texttt{SciPy}
-(\citeproc{ref-SciPy:2020}{Virtanen et al., 2020}), and \texttt{PyBOP}'s
-own algorithms such as adaptive moment estimation with weight decay
-(AdamW), gradient descent, and Cuckoo search.
+enabling the use of advanced models where previously this was not
+possible without specialised knowledge of battery modelling, parameter
+inference and software development. \texttt{PyBOP} emphasises clear and
+informative diagnostics and workflows to support users with varying
+levels of domain expertise, and provides access to a wide range of
+optimisation and sampling algorithms. These methods are provided through
+interfaces to \texttt{PINTS} (\citeproc{ref-Clerx:2019}{Clerx et al.,
+2019}), \texttt{SciPy} (\citeproc{ref-SciPy:2020}{Virtanen et al.,
+2020}), and \texttt{PyBOP}'s own implementations of algorithms such as
+adaptive moment estimation with weight decay (AdamW), gradient descent,
+and Cuckoo search.
 
@@
 experiments, the need for battery domain knowledge, and the
-computational cost of parameter estimation. \texttt{PyBOP} reduces these
-costs by providing fast parameter estimation, standardised workflows,
-and parameter set interoperability (via BPX).
+computational cost of parameter estimation. \texttt{PyBOP} reduces the
+requirements for the latter two categories by providing fast parameter
+estimation, standardised workflows, and parameter set interoperability
+(via BPX).
 
@@
 pack simulation (\citeproc{ref-Tranter2022}{Tranter et al., 2022}) and
-\texttt{pybamm-eis} for numerical impedance spectroscopy, since the
-identified parameters from \texttt{PyBOP} are easily exportable to these
-other packages.
+\texttt{pybamm-eis} for fast numerical computation of the
+electrochemical impedance for any battery model, since the identified
+parameters from \texttt{PyBOP} are easily exportable to these other
+packages.
 
@@
 
-\texttt{PyBOP} has a layered structure designed to compute forward model
-predictions, process the results, and package the necessary information
-for the optimisation and sampling algorithms. The forward model is
-solved using the popular battery modelling package, \texttt{PyBaMM},
-with construction, parameterisation, and discretisation managed by
-\texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This provides a
-robust object construction process with a consistent interface between
-models and optimisers. The statistical methods and optimisation
-algorithms are constructed to interface cleanly with the forward model
-predictions. Furthermore, identifiability metrics are provided with the
-estimated parameters (through Hessian approximation of the cost
-functions in frequentist workflows, and posterior distributions in
-Bayesian workflows).
+\texttt{PyBOP} has a layered structure designed to package up the
+necessary functionality to compute forward predictions, process the
+results, and run the optimisation and sampling algorithms. The forward
+model is solved using the popular battery modelling package,
+\texttt{PyBaMM}, with construction, parameterisation, and discretisation
+managed by \texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This
+provides a robust object construction process with a consistent
+interface between models and optimisers. The statistical methods and
+optimisation algorithms are constructed to interface cleanly with the
+forward model predictions. Furthermore, identifiability metrics are
+provided with the estimated parameters (through Hessian approximation of
+the cost functions in frequentist workflows, and posterior distributions
+in Bayesian workflows).
 
@@
 \texttt{PyBOP} formulates the inference process into four key classes,
-namely the model, the problem, the cost, and the optimiser/sampler, as
-shown in \autoref{fig:classes}. Each of these objects represents a base
-class with child classes constructing specialised functionality for
-inference or optimisation workflows. The model class constructs a
-\texttt{PyBaMM} forward model for a given set of model equations
-provided by \texttt{PyBaMM}, initial conditions, spatial discretisation,
-and numerical solver. By composing \texttt{PyBaMM} directly into
-\texttt{PyBOP}, specialised models can be constructed alongside the
-standard models, which can be modified, and optimally constructed for
-the inference tasks. One such example is spatial rediscretisation, which
-is performed when geometric parameters are optimised. In this situation,
-\texttt{PyBOP} minimally rediscretises the \texttt{PyBaMM} model while
-maintaining the problem, cost, and optimiser objects, providing improved
-performance benefits to users. Alongside construction of the forward
-model, \texttt{PyBOP}'s model class provides methods for obtaining
-sensitivities from the prediction, enabling gradient-based optimisation
-algorithms. This prediction, along with its corresponding sensitivities,
-is provided to the problem class for processing and exception control. A
-standardised data structure is then provided to the cost classes, which
-then computes a distance, design, or likelihood-based metric for
-optimisation. For deterministic optimisation, the optimisers minimise
-the corresponding cost function or the negative log-likelihood if a
-likelihood class is provided. Bayesian inference is provided by Monte
-Carlo sampling classes, which accept the LogPosterior class and sample
-from it using \texttt{PINTS} based Monte Carlo algorithms at the time of
-submission. In the typical workflow, the classes in
-\autoref{fig:classes} are constructed in sequence.
+namely, the model, the problem, the cost/likelihood, and the
+optimiser/sampler, as shown in \autoref{fig:classes}. Each of these
+objects represents a base class with child classes constructing
+specialised functionality for different workflows. The model class
+constructs a \texttt{PyBaMM} forward model for a given set of model
+equations, initial conditions, spatial discretisation, and numerical
+solver. By composing \texttt{PyBaMM} directly into \texttt{PyBOP},
+specialised models can be constructed alongside the standard models,
+which can be modified and purposefully constructed for different
+inference tasks. One such example is spatial re-discretisation, which is
+required when one or more geometric parameters are being optimised. In
+this situation, \texttt{PyBOP} rebuilds the \texttt{PyBaMM} model a
+minimum number of times while maintaining the problem, cost, and
+optimiser objects, providing improved performance benefits to users.
+Alongside construction of the forward model, \texttt{PyBOP}'s model
+class provides methods for obtaining sensitivities from the prediction,
+enabling gradient-based optimisation algorithms. A forward prediction,
+along with its corresponding sensitivities, is provided to the problem
+class for processing and exception control. A standardised data
+structure is then provided to the cost classes, which then computes a
+distance, design, or likelihood-based metric for optimisation. For
+deterministic optimisation, the optimisers minimise the cost function or
+the negative log-likelihood if a likelihood class is provided. Bayesian
+inference is provided by sampler classes, which accept the LogPosterior
+class and sample from it using \texttt{PINTS} based Monte Carlo
+algorithms at the time of submission. In the typical workflow, the
+classes in \autoref{fig:classes} are constructed in sequence.
 
@@
 specialised inference and optimisation processes. One such instance is
-parameter inference from EIS forward model predicitions. PyBOP
-discretises the EIS forward model inso sparse mass matrix form with
-accompanying auto-differentiated Jacobian. These objects are then
-translated into the frequency domain with a linear solution used to
-compute the battery model impedance. In this situation, the forward
-models are constructed within the spatial rediscretisation workflow,
-allowing for geometric parameter inference from EIS forward model
-predictions. Furthermore, \texttt{PyBOP} builds on the \texttt{JAX}
+parameter inference from electrochemical impedance spectroscopy (EIS)
+simulations. PyBOP discretises the EIS forward model into a sparse mass
+matrix form with accompanying auto-differentiated Jacobian. These
+objects are then translated into the frequency domain with a linear
+solution used to compute the battery model impedance. In this situation,
+the forward models are constructed within the spatial re-discretisation
+workflow, allowing for geometric parameter inference from EIS
+simulations. Furthermore, \texttt{PyBOP} builds on the \texttt{JAX}
 (\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
@@
 \caption{The core \texttt{PyBOP} architecture, showcasing the base class
-interfaces. Each class provide direct mapping to a classical step in the
-optimisation workflow. \label{fig:classes}}
+interfaces. Each class provides a direct mapping to a classical step in
+the optimisation workflow. \label{fig:classes}}
 \end{figure}
@@
 presented in \autoref{tab:optimisers}. It should be noted that
-\texttt{SciPy} minimize has gradient and non-gradient methods. From now
-on, the point-based parameterisation and design optimisation tasks will
-simply be referred to as optimisation tasks. This simplification can be
-justified by examining \autoref{eqn:parameterisation} and
-\autoref{eqn:design} and confirming that deterministic parameterisation
-can be viewed as an optimisation task to minimise a distance-based cost
-function.
+\texttt{SciPy} minimize comprises a range of gradient and non-gradient
+methods. From now on, the point-based parameterisation and design
+optimisation tasks will simply be referred to as optimisation tasks.
+This simplification can be justified by examining
+\autoref{eqn:parameterisation} and \autoref{eqn:design} and confirming
+that deterministic parameterisation can be viewed as an optimisation
+task to minimise a distance-based cost function.
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4722}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2963}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2315}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4359}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2308}}@{}}
 \caption{The currently supported optimisation algorithms classified by
@@
 SciPy minimize & SciPy differential evolution & \\
-& & \\
 \end{longtable}
 
-As discussed above, \texttt{PyBOP} provides both point-based
-optimisation estimates \autoref{tab:subclasses} in addition to Monte
-Carlo sampling routines in order to recover distributions of likely
-parameter values. The Monte Carlco methods construct a posterior
-distribution on the inference parameters, which can be used for
-uncertainty and practical identifiability. The individual sampler
-classes are currently composed within \texttt{PyBOP} from the
-\texttt{PINTS} library, with a base sampling class implemented for
-interoperability and direct integration with the \texttt{PyBOP} model,
-problem, and likelihood classes. The currently supported samplers are
-listed in \autoref{tab:samplers}.
+As discussed above, in addition to point-based optimisation estimates
+\autoref{tab:subclasses}, \texttt{PyBOP} also provides Monte Carlo
+sampling routines in order to recover distributions of likely parameter
+values. The Monte Carlo methods construct a posterior distribution on
+the inference parameters, which can be used to assess uncertainty and
+practical identifiability. The individual sampler classes are currently
+composed within \texttt{PyBOP} from the \texttt{PINTS} library, with a
+base sampler class implemented for interoperability and direct
+integration with the \texttt{PyBOP} model, problem, and likelihood
+classes. The currently supported samplers are listed in
+\autoref{tab:samplers}.
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1810}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2155}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1379}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2069}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2586}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1624}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2393}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1368}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2051}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2564}}@{}}
 \caption{Sampling methods supported by \texttt{PyBOP}, classified
@@
 \end{equation} \begin{equation}
-\mathbf{y}(t) = h(t,\mathbf{x}),
+\mathbf{y}(t) = h(t, \mathbf{x}, \mathbf{\theta}),
 \label{output}
@@
 \citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
-et al., 2023}).
+et al., 2023}; \citeproc{ref-Lu:2021}{Lu et al., 2021}).
 
@@
 \begin{equation}
-\min_{\mathbf{\theta}} ~ \mathcal{L}_{(\mathbf{y}_i)}(\mathbf{\theta}) ~~~
+\min_{\mathbf{\theta}} ~ \mathcal{L}_{(\hat{\mathbf{y}_i})}(\mathbf{\theta}) ~~~
 \textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
@@
 where \(\mathcal{L} : \mathbf{\theta} \mapsto [0,\infty)\) is a cost
-function that quantifies the agreement between the model,
-\(\mathbf{y}(t_i)\) and a sequence of measured observations
-\(\hat{\mathbf{y}_i}\) measured at times \(t_i\). Within the
+function that quantifies the agreement between the model output
+\(\mathbf{y}(t)\) and a sequence of observations
+\((\hat{\mathbf{y}_i})\) measured at times \(t_i\). Within the
 \texttt{PyBOP} framework, the \texttt{FittingProblem} class packages the
 model output along with the measured observations, both of which are
-then passed to the \texttt{FittingCost} classes for the computation of
-the specific cost function. For gradient-based optimisers, the Jacobian
-of the cost function with respect to the unknown parameters,
+then passed to the cost classes for the computation of the specific cost
+function. For gradient-based optimisers, the Jacobian of the cost
+function with respect to the unknown parameters,
 \((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
@@
 As mentioned above, \texttt{PyBOP} also provides inference and
-optimisation capabilities through numerical impedance spectroscopy. This
-is based on the methods presented in the \texttt{pybamm-eis} package and
-allows fast impedance computation by inversion of the sparse PyBaMM mass
-matrix after scaling and subtraction of the auto-differentiated Jacobian
-matrix. More information about this procedure is available in
-(\citeproc{ref-pybamm-eis}{Dhoot et al., n.d.}). The
-\autoref{fig:impedance-landscape} below shows the numerical impedance
-prediction available in \texttt{PyBOP} alongside the cost landscape
-constructed for the inference task. At the time of publication,
-gradient-based optimisation and sampling methods are not available when
-using an impedance workflow.
+optimisation capabilities through numerical computation of the
+electrochemical impedance. This is based on the methods presented in the
+\texttt{pybamm-eis} package and allows fast impedance computation by
+inversion of a sparse mass matrix after scaling and subtraction of the
+auto-differentiated Jacobian matrix. More information about this
+procedure is available in (\citeproc{ref-pybamm-eis}{Dhoot et al.,
+n.d.}). The \autoref{fig:impedance-landscape} below shows the numerical
+impedance prediction available in \texttt{PyBOP} alongside the cost
+landscape constructed for the inference task. At the time of
+publication, gradient-based optimisation and sampling methods are not
+available when using an impedance workflow.
 
@@
 identification in the time-domain; however, in general these two
-forward-model prediction methods can be combined for improved system
-excitation. As gradient information is available for this problem, the
-choice of distance-based cost function and optimiser is not constrained.
-Due to the different magnitudes of the two parameters, we apply the
+simulation methods can be combined for improved system excitation. As
+gradient information is available for this problem, the choice of
+distance-based cost function and optimiser is not constrained. Due to
+the difference in magnitude between the two parameters, we apply the
 logarithmic parameter transformation offered by \texttt{PyBOP}. This
-transforms the optimisers search space of the optimiser to allow for a
-common step size between the parameters, which is generally is not
-required, but improves convergence in this problem. As a demonstration
-of the parameterisation capabilities of \texttt{PyBOP},
+transforms the search space of the optimiser to allow for a common step
+size between the parameters, which is generally is not required, but
+improves convergence in this problem. As a demonstration of the
+parameterisation capabilities of \texttt{PyBOP},
 \autoref{fig:convergence-min-max} (left) shows the rate of convergence
@@
 Next, the performance of the various optimisation algorithms is
-presented by category: gradient-based in
+presented by category: gradient-based methods in
 \autoref{fig:optimiser-inference} (left), evolutionary strategies in
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_total.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_subplot.png}
 \caption{Cost landscape contour plot with corresponding optimisation
-traces. The top row represents the gradient-based optimisers, the middle
-row is the evolution-based, and the bottom row is the (meta)heuristics.
-The order from left to right corresponds to the entries in
-\autoref{tab:optimisers}. \label{fig:optimiser-inference}}
+traces. The three rows show the gradient-based optimisers (top),
+evolution strategies (middle), and (meta)heuristics (bottom).
+\label{fig:optimiser-inference}}
 \end{figure}
@@
 This parameterisation task can also be approached from a Bayesian
-perspective, which we will present below using \texttt{PyBOP}'s sampler
+perspective, which we will present below using \texttt{PyBOP}`s sampler
 methods. The optimisation equation presented in
 \autoref{eqn:parameterisation} does not represent the Bayesian parameter
-identification task, and as such we introduce the Bayes theorem as,
+identification task, and as such we introduce Bayes' theorem as,
 
@@
 
-where, \(P(\theta|D)\) is the posterior and represents the probability
+where, \(P(\theta|D)\) is the posterior, representing the probability
 density function of the parameter. \(P(D|\theta)\) is the likelihood
-function and assesses the parameter values alongside a noise model.
+function, which assesses the parameter values alongside a noise model.
 \(P(\theta)\) encapsulates the prior knowledge about the parameters, and
-finally \(P(D)\) is the model evidence and acts as a normalising
-constant so that the final posterior is a correctly scaled density
-function. Our goal in parameter inference is to identify the parameter
-values with the highest probability, which can be represented as a
-point-based metric or as the posterior distribution, which provides
+finally \(P(D)\) is the model evidence, which acts as a normalising
+constant. Our goal in parameter inference is to identify the parameter
+values with the highest probability, which can be represented either by
+a point-based metric or a posterior. The posterior distribution provides
 additional information about the uncertainty of the identified
-parameters. Monte Carlo sampling methods are available to obtain this
-posterior distribution. These methods sample from the posterior using a
-variety of methods, including gradient-based methods such as No-U-Turn
+parameters. Monte Carlo methods are used to sample from the posterior.
+The selection of Monte Carlo methods available in \texttt{PyBOP}
+includes gradient-based methods such as No-U-Turn
 (\citeproc{ref-NUTS:2011}{Hoffman \& Gelman, 2011}) and Hamiltonian
@@
 (\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}).
-\texttt{PyBOP} offers a sampling class that provides an interface to
-these samplers, which are supported by the Probabilistic Inference of
+\texttt{PyBOP} offers a sampler class that provides the interface to
+these samplers, which are provided by the Probabilistic Inference of
 Noise Time-Series (\texttt{PINTS}) package. \autoref{fig:posteriors}
@@
 viewed in a similar way to the parameterisation workflows described
-above, but with the aim of maximising a distance metric rather than
-minimising it. In the case of design optimisation to maximise
-gravimetric energy density, \texttt{PyBOP} minimises the negative of the
-cost function, where the cost metric is no longer a distance between two
-time series vectors, but the integrated energy of the vector normalised
-by with the corresponding cell mass. This is typically quantified for
-operating conditions such as a 1C discharge, at a given temperature.
-
-Design optimisation can be written in the form of a constrained
-optimisation problem as: \begin{equation}
+above, but with the aim of maximising a cost function rather than
+minimising it. \texttt{PyBOP} performs maximisation by minimising the
+negative of the cost function. In design problems, the cost metric is no
+longer a distance between two time series, but a metric evaluated on a
+model prediction. For example, to maximise the gravimetric energy (or
+power) density, the cost is the integral of the discharge energy (or
+power) normalised by the cell mass. Such metrics are typically
+quantified for operating conditions such as a 1C discharge, at a given
+temperature.
+
+In general, design optimisation can be written in the form of a
+constrained optimisation problem as: \begin{equation}
 \min_{\mathbf{\theta} \in \Omega} ~ \mathcal{L}(\mathbf{\theta}) ~~~
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/design.png}
-\caption{The gravimetric landscape alongside the corresponding initial
-and optimised voltage profiles for a 1C discharge.
+\caption{The gravimetric energy density landscape alongside the
+corresponding initial and optimised voltage profiles for a 1C discharge.
 \label{fig:design_gravimetric}}
@@
 \autoref{fig:design_gravimetric} (left) shows the optimiser's search
-over the gravimetric energy density parameter space. The predicted
-improvement in the discharge profile between the initial and optimised
-parameter values (right) for their respective applied 1C current.
+over the parameter space and (right) the predicted improvement in the
+discharge profile between the initial and optimised parameter values,
+simulated at their respective 1C rates.
 
@@
 
+\bibitem[\citeproctext]{ref-Lu:2021}
+Lu, D., Scott Trimboli, M., Fan, G., Zhang, R., \& Plett, G. L. (2021).
+Implementation of a physics-based model for half-cell open-circuit
+potential and full-cell open-circuit voltage estimates: {P}art {II}.
+Processing full-cell data. \emph{Journal of The Electrochemical
+Society}, \emph{168}(7), 070533.
+\url{https://doi.org/10.1149/1945-7111/ac11a5}
+
 \bibitem[\citeproctext]{ref-metropolis:1953}

### 2024-12-09 joss/design_plots.py

@@ -87,7 +87,13 @@ print(f"Optimised gravimetric energy density: {cost(result.x):.2f} Wh.kg-1")
 
 if create_plot["gravimetric"]:
     # Plot the cost landscape with optimisation path
-    gravimetric_fig = pybop.plot.contour(optim, steps=65, title=None)
+    gravimetric_fig = pybop.plot.contour(
+        optim,
+        steps=65,
+        title=None,
+        xaxis=dict(titlefont_size=14, tickfont_size=14),
+        yaxis=dict(titlefont_size=14, tickfont_size=14),
+    )
     gravimetric_fig.write_image("joss/figures/design_gravimetric.png")
 
 if create_plot["prediction"]:
@@ -99,6 +105,8 @@ if create_plot["prediction"]:
         width=576,
         height=576,
         legend=dict(orientation="h", yanchor="bottom", y=1.02, xanchor="right", x=1),
+        xaxis=dict(titlefont_size=14, tickfont_size=14),
+        yaxis=dict(titlefont_size=14, tickfont_size=14),
         show=False,
     )
     prediction_fig = figs[0]

### 2024-12-09 joss/image_combine.sh

@@ -7,11 +7,3 @@ convert +append figures/convergence_minimising.png figures/convergence_maximisin
 convert +append figures/simulation.png figures/landscape.png figures/joss/sim-landscape.png
 
 convert +append figures/impedance_spectrum.png figures/impedance_contour.png figures/joss/impedance.png
-
-convert +append figures/contour_gradient_0.png figures/contour_gradient_1.png figures/contour_gradient_2.png figures/contour_gradient_3.png figures/joss/contour_gradient.png
-
-convert +append figures/contour_evolution_0.png figures/contour_evolution_1.png figures/contour_evolution_2.png figures/contour_evolution_3.png figures/joss/contour_evolution.png
-
-convert +append figures/contour_heuristic_0.png figures/contour_heuristic_1.png figures/contour_heuristic_2.png figures/joss/contour_heuristic.png
-
-convert -append figures/joss/contour_gradient.png figures/joss/contour_evolution.png figures/joss/contour_heuristic.png figures/joss/contour_total.png

### 2024-12-09 joss/param_plots.py

@@ -10,19 +10,21 @@ import pybop
 from pybop.plot import PlotlyManager
 
 go = PlotlyManager().go
+px = PlotlyManager().px
+make_subplots = PlotlyManager().make_subplots
 np.random.seed(8)
 
 # Choose which plots to show and save
 create_plot = {}
 create_plot["simulation"] = False
-create_plot["landscape"] = True
+create_plot["landscape"] = False
 create_plot["minimising"] = False
 create_plot["maximising"] = False
-create_plot["gradient"] = False
-create_plot["evolution"] = False
-create_plot["heuristic"] = False
+create_plot["gradient"] = True
+create_plot["evolution"] = True
+create_plot["heuristic"] = True
 create_plot["posteriors"] = False
-create_plot["eis"] = True
+create_plot["eis"] = False
 
 
 # Parameter set and model definition
@@ -197,7 +199,7 @@ if create_plot["minimising"]:
         convergence_plot_dict = pybop.plot.StandardPlot(
             x=iteration_numbers,
             y=cost_log,
-            trace_names=type(cost).__name__,
+            trace_names=cost.name,
             trace_options={"line": {"width": 4, "dash": "dash"}},
         )
         convergence_traces.extend(convergence_plot_dict.traces)
@@ -212,8 +214,16 @@ if create_plot["minimising"]:
                 orientation="h", yanchor="bottom", y=1.02, xanchor="right", x=1
             ),
             plot_bgcolor="white",
-            width=576,
-            height=576,
+            yaxis=dict(
+                gridcolor=px.colors.qualitative.Pastel2[7],
+                zerolinecolor=px.colors.qualitative.Pastel2[7],
+                zerolinewidth=1,
+                titlefont_size=14,
+                tickfont_size=14,
+            ),
+            xaxis=dict(titlefont_size=14, tickfont_size=14),
+            width=600,
+            height=600,
         ),
     )
     convergence_fig.show()
@@ -258,13 +268,9 @@ if create_plot["maximising"]:
         convergence_plot_dict = pybop.plot.StandardPlot(
             x=iteration_numbers,
             y=cost_log,
-            trace_names=type(cost).__name__
+            trace_names=cost.name
             + " "
-            + (
-                type(cost.likelihood).__name__
-                if isinstance(cost, pybop.LogPosterior)
-                else ""
-            ),
+            + (cost.likelihood.name if isinstance(cost, pybop.LogPosterior) else ""),
[…]

### 2024-12-09 pybop/optimisers/base_optimiser.py
feat: align object name method, adds plotly express to PlotlyManager,

@@ -261,7 +261,7 @@ class BaseOptimiser:
         if x0 is not None:
             self.log["x0"].extend(x0)
 
-    def name(self):
+    def __name__(self):
         """
         Returns the name of the optimiser, to be overwritten by child classes.
 
@@ -301,6 +301,10 @@ class BaseOptimiser:
     @property
     def needs_sensitivities(self):
         return self._needs_sensitivities
+
+    @property
+    def name(self):
+        return self.__name__
 
 
 class OptimisationResult:

### 2024-12-09 pybop/optimisers/base_pints_optimiser.py

@@ -161,7 +161,7 @@ class BasePintsOptimiser(BaseOptimiser):
                     self.bounds["lower"], self.bounds["upper"]
                 )
 
-    def name(self):
+    def __name__(self):
         """Returns the name of the PINTS optimisation strategy."""
         return self.optimiser.name()
 

### 2024-12-09 pybop/optimisers/scipy_optimisers.py

@@ -283,9 +283,9 @@ class SciPyMinimize(BaseSciPyOptimiser):
             **self._options,
         )
 
-    def name(self):
+    def __name__(self):
         """Provides the name of the optimisation strategy."""
-        return "SciPyMinimize"
+        return "SciPy Minimize"
 
 
 class SciPyDifferentialEvolution(BaseSciPyOptimiser):
@@ -448,6 +448,6 @@ class SciPyDifferentialEvolution(BaseSciPyOptimiser):
             **self._options,
         )
 
-    def name(self):
+    def __name__(self):
         """Provides the name of the optimisation strategy."""
-        return "SciPyDifferentialEvolution"
+        return "SciPy Differential Evolution"

### 2024-12-09 pybop/plot/plotly_manager.py

@@ -17,6 +17,8 @@ class PlotlyManager:
     ----------
     go : module
         The Plotly graph_objects module for creating figures.
+    px : module
+        The Plotly express plots module.
     pio : module
         The Plotly input/output module for configuring the renderer.
     make_subplots : function
@@ -33,6 +35,7 @@ class PlotlyManager:
         """
         self.go = None
         self.pio = None
+        self.px = None
         self.make_subplots = None
         self.ensure_plotly_installed()
         self.check_renderer_settings()
@@ -43,11 +46,13 @@ class PlotlyManager:
         Check if Plotly is installed and import necessary modules; prompt for installation if missing.
         """
         try:
+            import plotly.express as px
             import plotly.graph_objs as go
             import plotly.io as pio
             from plotly.subplots import make_subplots
 
             self.go = go
+            self.px = px
             self.pio = pio
             self.make_subplots = make_subplots
         except ImportError:
@@ -86,11 +91,13 @@ class PlotlyManager:
         """
         Import Plotly modules and set the default renderer after installation.
         """
+        import plotly.express as px
         import plotly.graph_objs as go
         import plotly.io as pio
         from plotly.subplots import make_subplots
 
         self.go = go
+        self.px = px
         self.pio = pio
         self.make_subplots = make_subplots
         if pio.renderers.default == "":

### 2024-12-17 pybop/optimisers/_random_search.py
randomsearch modified

@@ -31,7 +31,7 @@ class RandomSearchImpl(PopulationBasedOptimiser):
         super().__init__(x0, sigma0, boundaries=boundaries)
 
         # Optimisation parameters
-        self._population_size = self._population_size
+        self._n = self._population_size
         self._iterations = 0
 
         # Set boundaries
@@ -41,7 +41,7 @@ class RandomSearchImpl(PopulationBasedOptimiser):
         self._candidates = np.random.uniform(
             low=self._boundaries.lower(),
             high=self._boundaries.upper(),
-            size=(self._population_size, self._dim),
+            size=(self._n, self._dim),
         ) if self._boundaries else np.random.normal(
             self._x0, self._sigma0, size=(self._n, self._dim)
         )
@@ -65,7 +65,7 @@ class RandomSearchImpl(PopulationBasedOptimiser):
             self._candidates = np.random.uniform(
                 low=self._boundaries.lower(),
                 high=self._boundaries.upper(),
-                size=(self._population_size, self._dim),
+                size=(self._n, self._dim),
             )
         else:
             self._candidates = np.random.normal(
@@ -80,12 +80,12 @@ class RandomSearchImpl(PopulationBasedOptimiser):
         accordingly.
         """
         if not self._ready_for_tell:
-            raise RuntimeError("Optimiser not ready for tell()")
+            raise RuntimeError("ask() must be called before tell().")
 
         self._iterations += 1
 
         # Evaluate solutions and update the best
-        for i in range(self._population_size):
+        for i in range(self._n):
             f_new = replies[i]
             if f_new < self._f_best:
                 self._f_best = f_new

### 2024-12-17 pybop/optimisers/_random_search.py
style: pre-commit fixes

@@ -38,12 +38,14 @@ class RandomSearchImpl(PopulationBasedOptimiser):
         self._boundaries = boundaries
 
         # Initialise population
-        self._candidates = np.random.uniform(
-            low=self._boundaries.lower(),
-            high=self._boundaries.upper(),
-            size=(self._n, self._dim),
-        ) if self._boundaries else np.random.normal(
-            self._x0, self._sigma0, size=(self._n, self._dim)
+        self._candidates = (
+            np.random.uniform(
+                low=self._boundaries.lower(),
+                high=self._boundaries.upper(),
+                size=(self._n, self._dim),
+            )
+            if self._boundaries
+            else np.random.normal(self._x0, self._sigma0, size=(self._n, self._dim))
         )
 
         # Initialise best solution
@@ -69,7 +71,7 @@ class RandomSearchImpl(PopulationBasedOptimiser):
             )
         else:
             self._candidates = np.random.normal(
-            self._x0, self._sigma0, size=(self._n, self._dim)
+                self._x0, self._sigma0, size=(self._n, self._dim)
             )
         return self.clip_candidates(self._candidates)
 

### 2024-12-17 pybop/optimisers/_random_search.py
boundary logic updated

@@ -34,9 +34,6 @@ class RandomSearchImpl(PopulationBasedOptimiser):
         self._n = self._population_size
         self._iterations = 0
 
-        # Set boundaries
-        self._boundaries = boundaries
-
         # Initialise population
         self._candidates = (
             np.random.uniform(

### 2024-12-17 tests/unit/test_optimisation.py
unit tests updated

@@ -265,7 +265,6 @@ class TestOptimisation:
             pybop.AdamW,
             pybop.CuckooSearch,
             pybop.GradientDescent,
-            pybop.RandomSearch,
         ]:
             optim = optimiser(cost)
             with pytest.raises(

### 2024-12-18 tests/unit/test_optimisation.py
unit tests updated

@@ -265,6 +265,7 @@ class TestOptimisation:
             pybop.AdamW,
             pybop.CuckooSearch,
             pybop.GradientDescent,
+            pybop.RandomSearch,
         ]:
             optim = optimiser(cost)
             with pytest.raises(
@@ -312,6 +313,11 @@ class TestOptimisation:
                 assert optim.optimiser.n_hyper_parameters() == 5
                 assert optim.optimiser.x_guessed() == optim.optimiser._x0
 
+            if optimiser is pybop.RandomSearch:
+                assert optimiser._dim == len(optim.optimiser._x0)
+                assert optimiser.x_best().shape == optim.optimiser._x0.shape
+                assert optimiser.f_best() == np.inf
+                assert optimiser._iterations == 0
         else:
             x0 = cost.parameters.initial_value()
             assert optim.x0 == x0

### 2024-12-18 tests/unit/test_optimisation.py
unit test changed

@@ -265,7 +265,7 @@ class TestOptimisation:
             pybop.AdamW,
             pybop.CuckooSearch,
             pybop.GradientDescent,
-            pybop.RandomSearch,
+            #pybop.RandomSearch,
         ]:
             optim = optimiser(cost)
             with pytest.raises(

### 2024-12-18 tests/unit/test_optimisation.py
style: pre-commit fixes

@@ -265,7 +265,7 @@ class TestOptimisation:
             pybop.AdamW,
             pybop.CuckooSearch,
             pybop.GradientDescent,
-            #pybop.RandomSearch,
+            # pybop.RandomSearch,
         ]:
             optim = optimiser(cost)
             with pytest.raises(

### 2024-12-18 joss/paper.preprint.tex
update and vectorise figures.

@@
   %
-  \ensuremath\mathparagraph]{Brady Planden%
+  ]{Brady Planden%
     \,\orcidlink{0000-0002-1082-9125}\,%
@@
   %
-  ]{David A. Howey%
+  \ensuremath\mathparagraph]{David A. Howey%
     \,\orcidlink{0000-0002-0620-3955}\,%
@@
 parameters, offering both deterministic and stochastic approaches with
-example workflows to assist users.
-
-\texttt{PyBOP} enables parameter identification from data for various
-battery models, including both the electrochemical and equivalent
-circuit models provided by the popular open-source \texttt{PyBaMM}
-package (\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}). Using the
-same workflow, \texttt{PyBOP} can also be used for design optimisation
-under user-defined operating conditions across a variety of model
-structures and design goals. \texttt{PyBOP} facilitates optimisation
-using a range of methods, with diagnostics for examining optimiser
-performance and convergence of the cost and corresponding parameters.
-Identified parameters can be used for prediction, on-line control, and
-design optimisation to accelerate research and development and improve
-battery utilisation.
+example workflows to assist users. \texttt{PyBOP} enables parameter
+identification from data for various battery models, including the
+electrochemical and equivalent circuit models provided by the popular
+open-source \texttt{PyBaMM} package (\citeproc{ref-Sulzer:2021}{Sulzer
+et al., 2021}). Using the same approaches, \texttt{PyBOP} can also be
+used for design optimisation under user-defined operating conditions
+across a variety of model structures and design goals. \texttt{PyBOP}
+facilitates optimisation with a range of methods, with diagnostics for
+examining optimiser performance and convergence of the cost and
+corresponding parameters. Identified parameters can be used for
+prediction, on-line estimation and control, and design optimisation,
+accelerating battery research and development.
 
@@
 
-\texttt{PyBOP} is a Python package that provides a user-friendly,
+\texttt{PyBOP} is a Python package providing a user-friendly,
 object-oriented interface for optimising battery model parameters.
@@
 (\citeproc{ref-Sulzer:2021}{Sulzer et al., 2021}) to formulate and solve
-battery models. Together, these packages serve a broad audience
-including students, engineers, and researchers in academia and industry,
-enabling the use of advanced models where previously this was not
-possible without specialised knowledge of battery modelling, parameter
-inference and software development. \texttt{PyBOP} emphasises clear and
+battery models. Together, these tools serve a broad audience including
+students, engineers, and researchers in academia and industry, enabling
+the use of advanced models where previously this was not possible
+without specialised knowledge of battery modelling, parameter inference,
+and software development. \texttt{PyBOP} emphasises clear and
 informative diagnostics and workflows to support users with varying
 levels of domain expertise, and provides access to a wide range of
-optimisation and sampling algorithms. These methods are provided through
+optimisation and sampling algorithms. These are enabled through
 interfaces to \texttt{PINTS} (\citeproc{ref-Clerx:2019}{Clerx et al.,
@@
 2020}), and \texttt{PyBOP}'s own implementations of algorithms such as
-adaptive moment estimation with weight decay (AdamW), gradient descent,
-and Cuckoo search.
+adaptive moment estimation with weight decay (AdamW)
+(\citeproc{ref-Loshchilov:2017}{Loshchilov \& Hutter, 2017}), gradient
+descent (\citeproc{ref-Cauchy:1847}{Cauchy \& others, 1847}), and cuckoo
+search (\citeproc{ref-Yang:2009}{Yang \& Suash Deb, 2009}).
 
 \texttt{PyBOP} supports the battery parameter exchange (BPX) standard
-(\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing battery
-parameter sets. These parameter sets are typically costly to obtain due
-to the specialised equipment and time required for characterisation
-experiments, the need for battery domain knowledge, and the
-computational cost of parameter estimation. \texttt{PyBOP} reduces the
-requirements for the latter two categories by providing fast parameter
-estimation, standardised workflows, and parameter set interoperability
-(via BPX).
-
-This package complements other lithium-ion battery modelling packages
+(\citeproc{ref-BPX:2023}{Korotkin et al., 2023}) for sharing parameter
+sets. These are typically costly to obtain due to the specialised
+equipment and time required for characterisation experiments, the need
+for domain knowledge, and the computational cost of estimation.
+\texttt{PyBOP} reduces the requirements for the latter two by providing
+fast parameter estimation methods, standardised workflows, and parameter
+set interoperability (via BPX).
+
+\texttt{PyBOP} complements other lithium-ion battery modelling packages
 built around \texttt{PyBaMM}, such as \texttt{liionpack} for battery
@@
 \texttt{pybamm-eis} for fast numerical computation of the
-electrochemical impedance for any battery model, since the identified
-parameters from \texttt{PyBOP} are easily exportable to these other
-packages.
+electrochemical impedance of any battery model. Identified
+\texttt{PyBOP} parameters are easily exportable to other packages.
 
@@
 
-\texttt{PyBOP} has a layered structure designed to package up the
-necessary functionality to compute forward predictions, process the
-results, and run the optimisation and sampling algorithms. The forward
-model is solved using the popular battery modelling package,
-\texttt{PyBaMM}, with construction, parameterisation, and discretisation
-managed by \texttt{PyBOP}'s model interface to \texttt{PyBaMM}. This
-provides a robust object construction process with a consistent
-interface between models and optimisers. The statistical methods and
-optimisation algorithms are constructed to interface cleanly with the
-forward model predictions. Furthermore, identifiability metrics are
-provided with the estimated parameters (through Hessian approximation of
-the cost functions in frequentist workflows, and posterior distributions
-in Bayesian workflows).
+\texttt{PyBOP} has a layered structure enabling the necessary
+functionality to compute forward predictions, process results, and run
+optimisation and sampling algorithms. The forward model is solved using
+the battery modelling software \texttt{PyBaMM}, with construction,
+parameterisation, and discretisation managed by \texttt{PyBOP}'s model
+interface to \texttt{PyBaMM}. This provides a robust object construction
+process with a consistent interface between forward models and
+optimisers. Furthermore, identifiability metrics are provided along with
+the estimated parameters (through Hessian approximation of the cost
+functions around the optimum point in frequentist workflows, and
+posterior distributions in Bayesian workflows).
 
@@
 \centering
-\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP-high-level.pdf}
-\caption{PyBOP's interface to supporting funding agencies, alongside a
-visualisation of the general workflow for parameterisation and
-optimisation. \label{fig:high-level}}
+\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP_components.drawio.png}
+\caption{The core \texttt{PyBOP} architecture with base class
+interfaces. Each class provides a direct mapping to a step in the
+optimisation workflow. \label{fig:classes}}
 \end{figure}
 
-\texttt{PyBOP} formulates the inference process into four key classes,
-namely, the model, the problem, the cost/likelihood, and the
-optimiser/sampler, as shown in \autoref{fig:classes}. Each of these
-objects represents a base class with child classes constructing
-specialised functionality for different workflows. The model class
-constructs a \texttt{PyBaMM} forward model for a given set of model
-equations, initial conditions, spatial discretisation, and numerical
-solver. By composing \texttt{PyBaMM} directly into \texttt{PyBOP},
-specialised models can be constructed alongside the standard models,
-which can be modified and purposefully constructed for different
-inference tasks. One such example is spatial re-discretisation, which is
-required when one or more geometric parameters are being optimised. In
-this situation, \texttt{PyBOP} rebuilds the \texttt{PyBaMM} model a
-minimum number of times while maintaining the problem, cost, and
-optimiser objects, providing improved performance benefits to users.
-Alongside construction of the forward model, \texttt{PyBOP}'s model
-class provides methods for obtaining sensitivities from the prediction,
-enabling gradient-based optimisation algorithms. A forward prediction,
-along with its corresponding sensitivities, is provided to the problem
-class for processing and exception control. A standardised data
-structure is then provided to the cost classes, which then computes a
-distance, design, or likelihood-based metric for optimisation. For
-deterministic optimisation, the optimisers minimise the cost function or
-the negative log-likelihood if a likelihood class is provided. Bayesian
-inference is provided by sampler classes, which accept the LogPosterior
-class and sample from it using \texttt{PINTS} based Monte Carlo
-algorithms at the time of submission. In the typical workflow, the
-classes in \autoref{fig:classes} are constructed in sequence.
+\texttt{PyBOP} formulates the inference process into four key classes:
+model, problem, cost (or likelihood), and optimiser (or sampler), as
+shown in \autoref{fig:classes}. Each of these objects represents a base
+class with child classes constructing specialised functionality for
+different workflows. The model class constructs a \texttt{PyBaMM}
+forward model with a specified set of equations, initial conditions,
+spatial discretisation, and numerical solver. By composing
+\texttt{PyBaMM} directly into \texttt{PyBOP}, specialised models can be
+constructed alongside the standard models that can also be modified for
+different inference tasks. One such example is spatial
+re-discretisation, which is required when one or more geometric
+parameters are being optimised. In this situation, \texttt{PyBOP}
+rebuilds the \texttt{PyBaMM} model only when necessary, reducing the
+total number of rebuilds, providing improved performance. Alongside
+construction of the forward model, \texttt{PyBOP}'s model class provides
+methods for obtaining sensitivities from the prediction, enabling
+gradient-based optimisation. A forward prediction, along with its
+corresponding sensitivities, is provided to the problem class for
+processing and exception control. A standardised data structure is then
+provided to the cost classes, which then computes a distance, design, or
+likelihood-based metric for optimisation. For point-based optimisation,
+the optimisers minimise the cost function or the negative log-likelihood
+if a likelihood class is provided. Bayesian inference is provided by
+sampler classes, which accept the \texttt{LogPosterior} class and sample
+from it using \texttt{PINTS}-based Monte Carlo algorithms at the time of
+submission. In the typical workflow, the classes in
+\autoref{fig:classes} are constructed in sequence, from left to right in
+the figure.
 
 In addition to the core architecture, \texttt{PyBOP} provides several
-specialised inference and optimisation processes. One such instance is
+specialised inference and optimisation features. One example is
 parameter inference from electrochemical impedance spectroscopy (EIS)
-simulations. PyBOP discretises the EIS forward model into a sparse mass
-matrix form with accompanying auto-differentiated Jacobian. These
-objects are then translated into the frequency domain with a linear
-solution used to compute the battery model impedance. In this situation,
-the forward models are constructed within the spatial re-discretisation
-workflow, allowing for geometric parameter inference from EIS
-simulations. Furthermore, \texttt{PyBOP} builds on the \texttt{JAX}
-(\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical solvers
-provided by \texttt{PyBaMM} by providing \texttt{JAX}-based cost
+simulations, where PyBOP discretises and linearises the EIS forward
+model into a sparse mass matrix form with accompanying
+auto-differentiated Jacobian. This is then translated into the frequency
+domain, giving a direct solution to compute the input-output impedance.
+In this situation, the forward models are constructed within the spatial
+re-discretisation workflow, allowing for geometric parameter inference
+from EIS simulations and data.
+
+A second specialised feature is that \texttt{PyBOP} builds on the
+\texttt{JAX} (\citeproc{ref-jax:2018}{Bradbury et al., 2018}) numerical
+solvers used by \texttt{PyBaMM} by providing \texttt{JAX}-based cost
 functions for automatic forward model differentiation with respect to
-the parameters. This functionality provides a performance improvement,
-and allows users to use many other JAX-based inference packages in order
-to optimise their cost function, such as \texttt{Numpyro}
+the parameters. This functionality provides a performance improvement
+and allows users to harness many other JAX-based inference packages to
+optimise cost functions, such as \texttt{Numpyro}
 (\citeproc{ref-numpyro:2019}{Phan et al., 2019}), \texttt{BlackJAX}
@@
 (\citeproc{ref-optax:2020}{DeepMind et al., 2020}).
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth,height=\textheight]{figures/PyBOP_components.drawio.png}
-\caption{The core \texttt{PyBOP} architecture, showcasing the base class
-interfaces. Each class provides a direct mapping to a classical step in
-the optimisation workflow. \label{fig:classes}}
-\end{figure}
 
@@
 \endlastfoot
-Single particle model (SPM) & Fitting problem & Sum squared error \\
-SPM with electrolyte (SPMe) & Design problem & Root mean squared
+Single-particle model (SPM) & Fitting problem & Sum-squared error \\
+SPM with electrolyte (SPMe) & Design problem & Root-mean-squared
 error \\
 Doyle-Fuller-Newman (DFN) & Observer & Minkowski \\
-Many particle model (MPM) & & Sum of power \\
+Many-particle model (MPM) & & Sum-of-power \\
 Multi-species multi-reaction (MSMR) & & Gaussian log likelihood \\
@@
 & & Gravimetric energy density \\
-& & Unscented Kalman filter \\
 \end{longtable}
 
-Similarly, the current algorithms available for optimisation tasks are
-presented in \autoref{tab:optimisers}. It should be noted that
-\texttt{SciPy} minimize comprises a range of gradient and non-gradient
-methods. From now on, the point-based parameterisation and design
-optimisation tasks will simply be referred to as optimisation tasks.
-This simplification can be justified by examining
-\autoref{eqn:parameterisation} and \autoref{eqn:design} and confirming
-that deterministic parameterisation can be viewed as an optimisation
-task to minimise a distance-based cost function.
+Similarly, the current algorithms available for optimisation are
+presented in \autoref{tab:optimisers}. It should be noted that SciPy
+minimize includes several gradient and non-gradient methods. From here
+on, the point-based parameterisation and design-optimisation tasks will
+simply be referred to as optimisation tasks. This simplification can be
+justified by comparing \autoref{eqn:parameterisation} and
+\autoref{eqn:design}; deterministic parameterisation is just an
+optimisation task to minimise a distance-based cost between model output
+and measured values.
 
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2308}}@{}}
-\caption{The currently supported optimisation algorithms classified by
-candidate solution type, including gradient information.
+\caption{Currently supported optimisers classified by candidate solution
+type, including gradient information.
 \label{tab:optimisers}}\tabularnewline
@@
 
-As discussed above, in addition to point-based optimisation estimates
-\autoref{tab:subclasses}, \texttt{PyBOP} also provides Monte Carlo
-sampling routines in order to recover distributions of likely parameter
-values. The Monte Carlo methods construct a posterior distribution on
-the inference parameters, which can be used to assess uncertainty and
-practical identifiability. The individual sampler classes are currently
-composed within \texttt{PyBOP} from the \texttt{PINTS} library, with a
-base sampler class implemented for interoperability and direct
-integration with the \texttt{PyBOP} model, problem, and likelihood
-classes. The currently supported samplers are listed in
-\autoref{tab:samplers}.
+In addition to deterministic optimisers \autoref{tab:subclasses},
+\texttt{PyBOP} also provides Monte Carlo sampling routines to estimate
+distributions of parameters within a Bayesian framework. These methods
+construct a posterior parameter distribution that can be used to assess
+uncertainty and practical identifiability. The individual sampler
+classes are currently composed within \texttt{PyBOP} from the
+\texttt{PINTS} library, with a base sampler class implemented for
+interoperability and direct integration with \texttt{PyBOP}'s model,
+problem, and likelihood classes. The currently supported samplers are
+listed in \autoref{tab:samplers}.
 
@@
 \endlastfoot
-Monomial Gamma & Delayed Rejection Adaptive & Rank---Shrinking &
-Differential Evolution & Metropolis Random Walk \\
-No-U-Turn & Haario Bardenet & Doubling & & Emcee Hammer \\
-Hamiltonian & Haario & Stepout & & Metropolis Adjusted Langevin \\
+Monomial gamma & Delayed rejection adaptive & Rank shrinking &
+Differential evolution & Metropolis random walk \\
+No-U-turn & Haario Bardenet & Doubling & & Emcee hammer \\
+Hamiltonian & Haario & Stepout & & Metropolis adjusted Langevin \\
 Relativistic & Rao Blackwell & & & \\
@@
 In general, battery models (after spatial discretisation) can be written
-in the form of a differential-algebraic system of equations:
+in the form of a differential-algebraic system of equations,
 \begin{equation}
@@
 \end{equation} \begin{equation}
-0 = g(t, \mathbf{x}, \mathbf{\theta})
+0 = g(t, \mathbf{x}, \mathbf{\theta}),
 \label{algebraic}
@@
 \citeproc{ref-Fuller:1994}{Fuller et al., 1994}) based on porous
-electrode theory and its reduced-order variants including the single
+electrode theory, and its reduced-order variants including the single
 particle model (SPM) (\citeproc{ref-Planella:2022}{Brosa Planella et
-al., 2022}), and the multi-species, multi-reaction (MSMR) model
+al., 2022}) and the multi-species multi-reaction (MSMR) model
 (\citeproc{ref-Verbrugge:2017}{Verbrugge et al., 2017}). Simplified
-models that retain acceptable predictive capabilities at a lower
-computational cost are widely used, for example in battery management
-systems, while physics-based models are required to understand the
-impact of physical parameters on battery performance. This separation of
-complexity traditionally results in multiple parameterisations for a
-single battery type, depending on the model structure.
+models that retain acceptable predictive accuracy at lower computational
+cost are widely used, for example in battery management systems, while
+physics-based models are required to understand the impact of physical
+parameters on performance. This separation of complexity traditionally
+results in multiple parameterisations for a single battery type,
+depending on the model structure.
 
@@
 The parameterisation of battery models is challenging due to the large
-number of parameters that need to be identified compared to the
-measurable outputs (\citeproc{ref-Andersson:2022}{Andersson et al.,
+number of parameters that need to be identified compared to the number
+of measurable outputs (\citeproc{ref-Andersson:2022}{Andersson et al.,
 2022}; \citeproc{ref-Miguel:2021}{Miguel et al., 2021};
 \citeproc{ref-Wang:2022}{Wang et al., 2022}). A complete
-parameterisation often requires a stepwise identification of smaller
-sets of parameters from a variety of excitations and different data sets
+parameterisation often requires stepwise identification of smaller sets
+of parameters from a variety of excitations and different data sets
 (\citeproc{ref-Chen:2020}{Chen et al., 2020};
 \citeproc{ref-Chu:2019}{Chu et al., 2019}; \citeproc{ref-Kirk:2022}{Kirk
-et al., 2023}; \citeproc{ref-Lu:2021}{Lu et al., 2021}).
-
-A generic data fitting optimisation problem may be formulated as:
+et al., 2023}; \citeproc{ref-Lu:2021}{Lu et al., 2021}). Furthermore,
+parameter identifiability can be poor for a given set of excitations and
+data sets, requiring improved experimental design in addition to
+uncertainty capable identification methods
+(\citeproc{ref-Aitio:2020}{Aitio et al., 2020}).
+
+A generic data-fitting optimisation problem may be formulated as:
 \begin{equation}
@@
 function. For gradient-based optimisers, the Jacobian of the cost
-function with respect to the unknown parameters,
-\((\frac{\partial \mathcal{L}}{\partial \theta})\) is computed for step
-size and directional information.
-
-Next, we demonstrate the fitting of synthetic data where the system
-parameters are known. In this example problem, we use \texttt{PyBaMM}'s
-implementation of the single particle model (SPM) with an added contact
-resistance submodel. We assume that the battery model is already
-parameterised except for two dynamic parameters, namely the lithium
-diffusivity of the negative electrode active material particle (denoted
-``negative particle diffusivity'') and the contact resistance. We
-generate synthetic data from a one-hour discharge from 100\% state of
-charge, to 0\% (denoted as 1C rate), followed by 30 minutes of
-relaxation. This data is then corrupted with zero mean Gaussian noise of
-amplitude 2mV, shown by the dots in
+function with respect to unknown parameters,
+\(\partial \mathcal{L} / \partial \theta\), is computed for step-size
+and directional information.
+
+Next, we demonstrate the fitting of synthetic data where the model
+parameters are known. Throughout this section, as an example, we use
+\texttt{PyBaMM}'s implementation of the single particle model with an
+added contact resistance submodel. We assume that the model is already
+fully parameterised apart from two dynamic parameters, namely, the
+lithium diffusivity of the negative electrode active material particles
+(denoted ``negative particle diffusivity'') and the contact resistance
+with corresponding true values of {[}3.3e-14 \(\text{m}^2/\text{s}\), 10
+mOhm{]}. We generate synthetic data with a one-hour discharge from 100\%
+to 0\% state of charge, denoted as 1C rate, followed by 30 minutes of
+relaxation. This dataset is then corrupted with zero-mean Gaussian noise
+of amplitude 2 mV, with the resulting signal shown by the blue dots in
 \autoref{fig:inference-time-landscape} (left). The initial states are
-assumed known, although such an assumption is not generally necessary.
-The \texttt{PyBOP} repository contains multiple
+assumed known, although this assumption is not generally necessary. The
+\texttt{PyBOP} repository contains several other
 \href{https://github.com/pybop-team/PyBOP/tree/develop/examples/notebooks}{example
-notebooks} that follow a similar inference process for readers that
-would like further information. The underlying cost landscape to be
-explored by the optimiser is shown in
-\autoref{fig:inference-time-landscape} (right) with the initial position
-denoted alongside the known system parameters for this synthetic
-inference task. In general, the true parameters are not known a priori.
+notebooks} that follow a similar inference process. The underlying cost
+landscape to be explored by the optimiser is shown in
+\autoref{fig:inference-time-landscape} (right), with the initial
+position denoted alongside the known true system parameters for this
+synthetic inference task. In general, the true parameters are not known.
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
-\caption{The cost landscape for the time-series parameterisation problem
-with a root mean squared error cost function.
+\caption{The fitted synthetic dataset (left) and cost landscape (right)
+for an example time-series battery model parameterisation using a
+root-mean-squared error cost function.
 \label{fig:inference-time-landscape}}
@@
 
-As mentioned above, \texttt{PyBOP} also provides inference and
-optimisation capabilities through numerical computation of the
-electrochemical impedance. This is based on the methods presented in the
-\texttt{pybamm-eis} package and allows fast impedance computation by
-inversion of a sparse mass matrix after scaling and subtraction of the
-auto-differentiated Jacobian matrix. More information about this
-procedure is available in (\citeproc{ref-pybamm-eis}{Dhoot et al.,
-n.d.}). The \autoref{fig:impedance-landscape} below shows the numerical
-impedance prediction available in \texttt{PyBOP} alongside the cost
-landscape constructed for the inference task. At the time of
-publication, gradient-based optimisation and sampling methods are not
-available when using an impedance workflow.
+Using the same model and parameters, we now showcase \texttt{PyBOP}'s
+capability to fit electrochemical impedance data using methods within
+\texttt{pybamm-eis} that enable fast impedance computation of battery
+models (\citeproc{ref-pybamm-eis}{Dhoot et al., 2024}).
+\autoref{fig:impedance-landscape} shows the numerical impedance
+prediction available in \texttt{PyBOP} alongside the cost landscape
+constructed for the inference task. At the time of publication,
+gradient-based optimisation and sampling methods are not available when
+using an impedance workflow.
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/impedance.png}
-\caption{The cost landscape for the frequency domain impedance
-parameterisation problem with a root mean squared error cost function at
-5\% SOC. \label{fig:impedance-landscape}}
+\caption{The data and model fit (left) and cost landscape (right) for a
+frequency-domain impedance parameterisation with a root-mean-squared
+error cost function, at 5\% SOC. \label{fig:impedance-landscape}}
 \end{figure}
 
-To avoid over complicating this example, we will continue with
-identification in the time-domain; however, in general these two
-simulation methods can be combined for improved system excitation. As
-gradient information is available for this problem, the choice of
-distance-based cost function and optimiser is not constrained. Due to
-the difference in magnitude between the two parameters, we apply the
-logarithmic parameter transformation offered by \texttt{PyBOP}. This
-transforms the search space of the optimiser to allow for a common step
-size between the parameters, which is generally is not required, but
-improves convergence in this problem. As a demonstration of the
+To avoid confusion in the remainder of this section, we will continue
+with identification of the example model in the time-domain. In general,
+however, time- and frequency-domain models and data may be combined for
+improved parameterisation. As gradient information is available for our
+time-domain example, the choice of distance-based cost function and
+optimiser is not constrained. Due to the difference in magnitude between
+the two parameters, we apply the logarithmic parameter transformation
+offered by \texttt{PyBOP}. This transforms the search space of the
+optimiser to allow for a common step size between the parameters,
+improving convergence in this particular case. As a demonstration of the
 parameterisation capabilities of \texttt{PyBOP},
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/converge.png}
-\caption{Convergence in the likelihood functions obtained using various
-likelihood functions and the L-BFGS-B algorithm.
+\caption{Optimiser convergence using various cost (left) and likelihood
+(right) functions and the L-BFGS-B algorithm.
 \label{fig:convergence-min-max}}
@@
 
-Next, the performance of the various optimisation algorithms is
-presented by category: gradient-based methods in
-\autoref{fig:optimiser-inference} (left), evolutionary strategies in
-\autoref{fig:optimiser-inference} (middle) and (meta)heuristics in
-\autoref{fig:optimiser-inference} (right) for a mean squared error cost
-function. Note that the performance of the optimiser depends on the cost
-environment, prior information and corresponding hyperparameters for
-each specific problem.
+We now show example optimiser convergence of various algorithms against
+the same model and two parameters, across several categories:
+gradient-based methods in \autoref{fig:optimiser-inference1} (left),
+evolutionary strategies in \autoref{fig:optimiser-inference1} (middle)
+and (meta)heuristics in \autoref{fig:optimiser-inference1} (right) using
+a mean-squared-error cost function. We also show cost functions and
+solution iterations in \autoref{fig:optimiser-inference2}, with the
+three rows showing the gradient-based optimisers (top), evolution
+strategies (middle), and (meta)heuristics (bottom). Note that the
+performance of the optimiser depends on the cost environment, initial
+guess or prior, and corresponding hyperparameters for each specific
+problem.
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/optimisers_parameters.png}
-\caption{Convergence in the parameter values obtained for the various
-optimisation algorithms provided by \texttt{PyBOP}.
-\label{fig:optimiser-inference}}
+\caption{Convergence in parameter values for several optimisation
+algorithms provided by \texttt{PyBOP}. \label{fig:optimiser-inference1}}
 \end{figure}
@@
 \caption{Cost landscape contour plot with corresponding optimisation
-traces. The three rows show the gradient-based optimisers (top),
-evolution strategies (middle), and (meta)heuristics (bottom).
-\label{fig:optimiser-inference}}
+traces, for several optimisers. \label{fig:optimiser-inference2}}
 \end{figure}
 
-This parameterisation task can also be approached from a Bayesian
-perspective, which we will present below using \texttt{PyBOP}`s sampler
-methods. The optimisation equation presented in
-\autoref{eqn:parameterisation} does not represent the Bayesian parameter
-identification task, and as such we introduce Bayes' theorem as,
+This example parameterisation task can also be approached from a
+Bayesian perspective, solved using \texttt{PyBOP}`s sampler methods.
+First, we introduce Bayes' rule,
 
 \begin{equation}
-P(\theta|D) = \frac{P(D|\theta)P(\theta)}{P(D)}
+P(\theta|D) = \frac{P(D|\theta)P(\theta)}{P(D)},
 \label{eqn:bayes_theorem}
@@
 
-where, \(P(\theta|D)\) is the posterior, representing the probability
-density function of the parameter. \(P(D|\theta)\) is the likelihood
-function, which assesses the parameter values alongside a noise model.
-\(P(\theta)\) encapsulates the prior knowledge about the parameters, and
-finally \(P(D)\) is the model evidence, which acts as a normalising
-constant. Our goal in parameter inference is to identify the parameter
-values with the highest probability, which can be represented either by
-a point-based metric or a posterior. The posterior distribution provides
-additional information about the uncertainty of the identified
-parameters. Monte Carlo methods are used to sample from the posterior.
-The selection of Monte Carlo methods available in \texttt{PyBOP}
-includes gradient-based methods such as No-U-Turn
+where \(P(\theta|D)\) is the posterior parameter distribution,
+\(P(D|\theta)\) is the likelihood function, \(P(\theta)\) is the prior
+parameter distribution, and \(P(D)\) is the model evidence, or marginal
+likelihood, which acts as a normalising constant. In the case of maximum
+likelihood estimation or maximum a posteriori estimation, one wishes to
+maximise \(P(D|\theta)\) or \(P(\theta|D)\), respectively, and this may
+be formulated as an optimisation problem as per
+\autoref{eqn:parameterisation}.
+
+To estimate the full posterior parameter distribution, however, one must
+use sampling or other inference methods to reconstruct the function
+\(P(\theta|D)\). The posterior distribution provides information about
+the uncertainty of the identified parameters, e.g., by calculating the
+variance or other moments. Monte Carlo methods are used here to sample
+from the posterior. The selection of Monte Carlo methods available in
+\texttt{PyBOP} includes gradient-based methods such as no-u-turn
 (\citeproc{ref-NUTS:2011}{Hoffman \& Gelman, 2011}) and Hamiltonian
 (\citeproc{ref-Hamiltonian:2011}{Brooks et al., 2011}), as well as
-heuristic methods such as Differential Evolution
-(\citeproc{ref-DiffEvolution:2006}{Braak, 2006}), and finally
-conventional methods based on random sampling with rejection criteria
+heuristic methods such as differential evolution
+(\citeproc{ref-DiffEvolution:2006}{Braak, 2006}), and also conventional
+methods based on random sampling with rejection criteria
 (\citeproc{ref-metropolis:1953}{Metropolis et al., 1953}).
 \texttt{PyBOP} offers a sampler class that provides the interface to
-these samplers, which are provided by the Probabilistic Inference of
-Noise Time-Series (\texttt{PINTS}) package. \autoref{fig:posteriors}
-below shows the sampled posterior for the synthetic workflow described
-above, using an adaptive covariance-based sampler, Haario Bardenet
-(\citeproc{ref-Haario:2001}{Haario et al., 2001}).
+samplers, the latter being provided by the probabilistic inference on
+noisy time-series (\texttt{PINTS}) package. \autoref{fig:posteriors}
+shows the sampled posteriors for the synthetic model described
+previously, using an adaptive covariance-based sampler called Haario
+Bardenet (\citeproc{ref-Haario:2001}{Haario et al., 2001}).
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/posteriors.png}
-\caption{Posterior distributions for model parameters alongside
-identified noise on the observations. Shaded area denotes confidence
-bounds for each parameter. \label{fig:posteriors}}
+\caption{Posterior distributions of model parameters alongside
+identified noise on the observations. Shaded area denotes the 95th
+percentile credible interval for each parameter. \label{fig:posteriors}}
 \end{figure}
@@
 
-Design optimisation is supported within \texttt{PyBOP} to guide future
-battery design development by identifying parameter sensitivities that
-can unlock improvements in battery performance. This problem can be
-viewed in a similar way to the parameterisation workflows described
-above, but with the aim of maximising a cost function rather than
-minimising it. \texttt{PyBOP} performs maximisation by minimising the
-negative of the cost function. In design problems, the cost metric is no
-longer a distance between two time series, but a metric evaluated on a
-model prediction. For example, to maximise the gravimetric energy (or
-power) density, the cost is the integral of the discharge energy (or
-power) normalised by the cell mass. Such metrics are typically
-quantified for operating conditions such as a 1C discharge, at a given
-temperature.
-
-In general, design optimisation can be written in the form of a
-constrained optimisation problem as: \begin{equation}
+Design optimisation is supported in \texttt{PyBOP} to guide device
+design development by identifying parameter sensitivities that can
+unlock improvements in performance. This problem can be viewed in a
+similar way to the parameterisation workflows described previously, but
+with the aim of maximising a design-objective cost function rather than
+minimising a distance cost function. \texttt{PyBOP} performs
+maximisation by minimising the negative of the cost function. In design
+problems, the cost metric is no longer a distance between two time
+series, but a metric evaluated on a model prediction. For example, to
+maximise the gravimetric energy (or power) density, the cost is the
+integral of the discharge energy (or power) normalised by the cell mass.
+Such metrics are typically quantified for operating conditions such as a
+1C discharge, at a given temperature.
+
+In general, design optimisation can be written as a constrained
+optimisation problem, \begin{equation}
 \min_{\mathbf{\theta} \in \Omega} ~ \mathcal{L}(\mathbf{\theta}) ~~~
-\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions})}
+\textrm{subject to equations (\ref{dynamics})\textrm{-}(\ref{initial_conditions}),}
 \label{eqn:design}
@@
 
-As an example, we consider the problem of maximising the gravimetric
-energy density subject to constraints on two of the geometric electrode
-parameters (\citeproc{ref-Couto:2023}{Couto et al., 2023}). For this
-example, we use the\texttt{PyBaMM} implementation of the single particle
-model with electrolyte (SPMe) to investigate the effect of the positive
+As an example, we consider the challenge of maximising the gravimetric
+energy density, subject to constraints on two of the geometric electrode
+parameters (\citeproc{ref-Couto:2023}{Couto et al., 2023}). In this case
+we use the \texttt{PyBaMM} implementation of the single particle model
+with electrolyte (SPMe) to investigate the impact of the positive
 electrode thickness and the active material volume fraction on the
-target cost. Since the active material volume fraction is related to the
-electrode porosity, the porosity is defined with a driven constraint
-from the volume fraction. In this problem, we estimate the 1C rate from
-the theoretical capacity for each iteration of the design. For this
-example, we employ the Particle Swarm Optimisation (PSO) algorithm.
+energy density. Since the active material volume fraction and electrode
+porosity sum to unity, the porosity for each optimisation iteration is
+defined in relation to the volume fraction. In this problem, we estimate
+the 1C rate from the theoretical capacity for each iteration of the
+design and use the particle swarm optimisation algorithm.
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{figures/joss/design.png}
-\caption{The gravimetric energy density landscape alongside the
-corresponding initial and optimised voltage profiles for a 1C discharge.
-\label{fig:design_gravimetric}}
+\caption{Gravimetric energy density cost landscape alongside the initial
+and optimised voltage profiles, for a fixed 1C discharge set from the
+initial design. \label{fig:design_gravimetric}}
 \end{figure}
 
-\autoref{fig:design_gravimetric} (left) shows the optimiser's search
+\autoref{fig:design_gravimetric} (left) shows the Nelder-Mead search
 over the parameter space and (right) the predicted improvement in the
@@
 \href{https://github.com/pybop-team/PyBOP?tab=readme-ov-file\#contributors-}{contributors}
-to this package. This work was supported by the Faraday Institution
-Multiscale Modelling (MSM) project (ref. FIRG059), UKRI's Horizon Europe
+to \texttt{PyBOP}. This work was supported by the Faraday Institution
+Multiscale Modelling project (ref. FIRG059), UKRI's Horizon Europe
 Guarantee (ref. 10038031), and EU IntelLiGent project (ref. 101069765).
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-Aitio:2020}
+Aitio, A., Marquis, S. G., Ascencio, P., \& Howey, D. (2020). Bayesian
+parameter estimation applied to the li-ion battery single particle model
+with electrolyte dynamicsa⁎this work was carried out with funding
+received from the faraday institution (faraday.ac.uk; EP/S003053/1, ref.
+FIRG003). Scott marquis was supported by the EPSRC centre for doctoral
+training in industrially focused mathematical modelling (EP/L015803/1)
+in collaboration with siemens corporate technology.
+\emph{IFAC-PapersOnLine}, \emph{53}(2), 12497--12504.
+https://doi.org/\url{https://doi.org/10.1016/j.ifacol.2020.12.1770}
+
 \bibitem[\citeproctext]{ref-Andersson:2022}
@@
 
+\bibitem[\citeproctext]{ref-Cauchy:1847}
+Cauchy, A., \& others. (1847). M{é}thode g{é}n{é}rale pour la
+r{é}solution des systemes d'{é}quations simultan{é}es. \emph{Comp. Rend.
+Sci. Paris}, \emph{25}(1847), 536--538.
+
 \bibitem[\citeproctext]{ref-Chen:2020}
@@
 \bibitem[\citeproctext]{ref-pybamm-eis}
-Dhoot, R., Timms, R., \& Please, C. (n.d.). \emph{PyBaMM EIS: Efficient
+Dhoot, R., Timms, R., \& Please, C. (2024). \emph{PyBaMM EIS: Efficient
 linear algebra methods to determine li-ion battery behaviour} (Version
@@
 Faraday Institution. \url{https://github.com/FaradayInstitution/BPX}
+
+\bibitem[\citeproctext]{ref-Loshchilov:2017}
+Loshchilov, I., \& Hutter, F. (2017). \emph{Decoupled {Weight} {Decay}
+{Regularization}}. arXiv.
+\url{https://doi.org/10.48550/ARXIV.1711.05101}
 
@@
 
+\bibitem[\citeproctext]{ref-Yang:2009}
+Yang, X.-S., \& Suash Deb. (2009). Cuckoo {Search} via levy flights.
+\emph{2009 {World} {Congress} on {Nature} \& {Biologically} {Inspired}
+{Computing} ({NaBIC})}, 210--214.
+\url{https://doi.org/10.1109/NABIC.2009.5393690}
+
 \end{CSLReferences}

### 2024-12-18 joss/design_plots.py

@@ -1,5 +1,7 @@
 # A script to generate design optimisation plots for the JOSS paper.
 
+import time
+
 import numpy as np
 import pybamm
 from pybamm import Parameter
@@ -38,7 +40,6 @@ parameter_set.update(
     },
     check_already_exists=False,
 )
-initial_capacity = parameter_set["Nominal cell capacity [A.h]"]
 model = pybop.lithium_ion.SPMe(
     parameter_set=parameter_set, solver=pybamm.IDAKLUSolver(atol=1e-6, rtol=1e-6)
 )
@@ -73,6 +74,23 @@ problem = pybop.DesignProblem(
     initial_state={"Initial SoC": 1.0},
     update_capacity=True,
 )
+parameter_set.update(
+    {
+        "Positive electrode thickness [m]": parameters[
+            "Positive electrode thickness [m]"
+        ].initial_value
+    }
+)
+parameter_set.update(
+    {
+        "Positive electrode active material volume fraction": parameters[
+            "Positive electrode active material volume fraction"
+        ].initial_value
+    }
+)
+initial_capacity = problem.model.approximate_capacity(parameter_set)
+problem.model.parameter_set.update({"Nominal cell capacity [A.h]": initial_capacity})
+
 cost = pybop.GravimetricEnergyDensity(problem)
 optim = pybop.NelderMead(
     cost,
@@ -94,12 +112,14 @@ if create_plot["gravimetric"]:
     # Plot the cost landscape with optimisation path
     gravimetric_fig = pybop.plot.contour(
         optim,
-        steps=65,
+        steps=60,
         title=None,
         xaxis=dict(titlefont_size=14, tickfont_size=14),
         yaxis=dict(titlefont_size=14, tickfont_size=14),
     )
-    gravimetric_fig.write_image("joss/figures/design_gravimetric.png")
+    gravimetric_fig.write_image("joss/figures/design_gravimetric.pdf")
+    time.sleep(3)
+    gravimetric_fig.write_image("joss/figures/design_gravimetric.pdf")
 
 problem.update_capacity = False
 problem.model.parameter_set.update({"Nominal cell capacity [A.h]": initial_capacity})
@@ -120,4 +140,4 @@ if create_plot["prediction"]:
     prediction_fig.data[0].name = f"Initial result: {cost(result.x0):.2f} Wh.kg-1"
     prediction_fig.data[1].name = f"Optimised result: {cost(result.x):.2f} Wh.kg-1"
[…]

### 2024-12-18 joss/image_combine.sh

@@ -1,9 +1,9 @@
-convert +append figures/design_gravimetric.png figures/design_prediction.png figures/joss/design.png
+pdfjam --nup 2x1 figures/design_prediction.pdf figures/design_gravimetric.pdf --landscape --outfile figures/joss/design.pdf --papersize '{450px,900px}'
 
-convert +append figures/gradient_parameters.png figures/evolution_parameters.png figures/heuristic_parameters.png figures/joss/optimisers_parameters.png
+pdfjam --nup 3x1 figures/gradient_parameters.pdf figures/evolution_parameters.pdf figures/heuristic_parameters.pdf --landscape --outfile figures/joss/optimisers_parameters.pdf --papersize '{768px,1296px}'
 
-convert +append figures/convergence_minimising.png figures/convergence_maximising.png figures/joss/converge.png
+pdfjam --nup 2x1 figures/convergence_minimising.pdf figures/convergence_maximising.pdf --landscape --outfile figures/joss/converge.pdf --papersize '{450px,900px}'
 
-convert +append figures/simulation.png figures/landscape.png figures/joss/sim-landscape.png
+pdfjam --nup 2x1 figures/simulation.pdf figures/landscape.pdf --landscape --outfile figures/joss/sim-landscape.pdf --papersize '{450px,900px}'
 
-convert +append figures/impedance_spectrum.png figures/impedance_contour.png figures/joss/impedance.png
+pdfjam --nup 2x1 figures/impedance_spectrum.pdf figures/impedance_contour.pdf --landscape --outfile figures/joss/impedance.pdf --papersize '{450px,900px}'

### 2024-12-18 joss/param_plots.py

@@ -1,5 +1,7 @@
 # A script to generate parameterisation plots for the JOSS paper.
 
+import time
+
 import matplotlib.pyplot as plt
 import numpy as np
 import plotly
@@ -16,16 +18,15 @@ np.random.seed(8)
 
 # Choose which plots to show and save
 create_plot = {}
-create_plot["simulation"] = False
-create_plot["landscape"] = False
-create_plot["minimising"] = False
-create_plot["maximising"] = False
+create_plot["simulation"] = True
+create_plot["landscape"] = True
+create_plot["minimising"] = True
+create_plot["maximising"] = True
 create_plot["gradient"] = True
 create_plot["evolution"] = True
 create_plot["heuristic"] = True
-create_plot["posteriors"] = False
-create_plot["eis"] = False
-
+create_plot["posteriors"] = True
+create_plot["eis"] = True
 
 # Parameter set and model definition
 parameter_set = pybop.ParameterSet.pybamm("Chen2020")
@@ -94,7 +95,9 @@ if create_plot["simulation"]:
         legend_traceorder="reversed",
     )
     simulation_fig.show()
-    simulation_fig.write_image("joss/figures/simulation.png")
+    simulation_fig.write_image("joss/figures/simulation.pdf")
+    time.sleep(3)
+    simulation_fig.write_image("joss/figures/simulation.pdf")
 
 # Form dataset
 dataset = pybop.Dataset(
@@ -153,7 +156,7 @@ if create_plot["landscape"]:
         )
     )
     landscape_fig.show()
-    landscape_fig.write_image("joss/figures/landscape.png")
+    landscape_fig.write_image("joss/figures/landscape.pdf")
 
 
 # Categorise the costs
@@ -227,7 +230,7 @@ if create_plot["minimising"]:
         ),
     )
     convergence_fig.show()
-    convergence_fig.write_image("joss/figures/convergence_minimising.png")
+    convergence_fig.write_image("joss/figures/convergence_minimising.pdf")
 
 
 if create_plot["maximising"]:
@@ -303,7 +306,7 @@ if create_plot["maximising"]:
         ),
     )
     convergence_fig.show()
-    convergence_fig.write_image("joss/figures/convergence_maximising.png")
+    convergence_fig.write_image("joss/figures/convergence_maximising.pdf")
 
 
 # Categorise the optimisers
@@ -342,7 +345,7 @@ subplot_contour_fig = make_subplots(
     shared_xaxes=True,
     x_title="Negative particle diffusivity [m2.s-1]",
[…]

### 2024-12-18 pybop/optimisers/pints_optimisers.py
fix: RandomSearch with multistart

@@ -703,6 +703,7 @@ class RandomSearch(BasePintsOptimiser):
         max_iterations: int = None,
         min_iterations: int = 2,
         max_unchanged_iterations: int = 15,
+        multistart: int = 1,
         parallel: bool = False,
         **optimiser_kwargs,
     ):
@@ -712,6 +713,7 @@ class RandomSearch(BasePintsOptimiser):
             max_iterations,
             min_iterations,
             max_unchanged_iterations,
+            multistart,
             parallel,
             **optimiser_kwargs,
         )

### 2024-12-18 pybop/_classification.py
Update input to OptimisationResult

@@ -1,25 +1,26 @@
 import numpy as np
 
+from pybop import OptimisationResult
 
-def classify_using_Hessian(optim, x=None, epsilon=1e-2):
+
+def classify_using_Hessian(result: OptimisationResult, epsilon=1e-2):
     """
     A simple check for parameter correlations based on numerical approximation
     of the Hessian matrix at the optimal point using central finite differences.
 
     Parameters
     ---------
-    x : array-like, optional
-        The parameters values assumed to be the optimal point (default: optim.result.x).
     epsilon : float, optional
+    result : OptimisationResult
+        The PyBOP optimisation results.
         A small positive value used to check proximity to the parameter bounds and as the
         perturbation distance used in the finite difference calculations (default: 1e-2).
     """
-
-    if x is None:
-        x = optim.result.x
-        final_cost = optim.result.final_cost
-    else:
-        final_cost = optim.cost(x)
+    x = result.x
+    final_cost = result.final_cost
+    cost = result.cost
+    parameters = result.cost.parameters
+    minimising = result.optim.minimising
 
     n = len(x)
     if n != 2:
@@ -29,7 +30,7 @@ def classify_using_Hessian(optim, x=None, epsilon=1e-2):
         )
 
     # Get a list of parameter names for use in the output message
-    names = list(optim.cost.parameters.keys())
+    names = list(parameters.keys())
 
     # Evaluate the cost for a grid of surrounding points
     dx = x * epsilon
@@ -39,17 +40,18 @@ def classify_using_Hessian(optim, x=None, epsilon=1e-2):
             if i == j == 1:
                 costs[1, 1, 0] = final_cost
                 costs[1, 1, 1] = final_cost
-            costs[i, j, 0] = optim.cost(x + np.multiply([i - 1, j - 1], dx))
-            costs[i, j, 1] = optim.cost(x + np.multiply([i - 1, j - 1], 2 * dx))
+            else:
+                costs[i, j, 0] = cost(x + np.multiply([i - 1, j - 1], dx))
+                costs[i, j, 1] = cost(x + np.multiply([i - 1, j - 1], 2 * dx))
 
     # Classify the result
-    if (optim.minimising and np.any(costs < final_cost)) or (
-        not optim.minimising and np.any(costs > final_cost)
+    if (minimising and np.any(costs < final_cost)) or (
+        not minimising and np.any(costs > final_cost)
     ):
         message = "The optimiser has not converged to a stationary point."
 
[…]

### 2024-12-18 tests/integration/test_classification.py

@@ -75,38 +75,39 @@ class TestClassification:
     @pytest.mark.integration
     def test_classify_using_Hessian(self, problem):
         cost = pybop.RootMeanSquaredError(problem)
-        optim = pybop.XNES(cost=cost)
-
-        x = np.asarray(cost.parameters.true_value())
+        x = cost.parameters.true_value()
         bounds = cost.parameters.get_bounds()
-        optim_x = np.clip(x, bounds["lower"], bounds["upper"])
+        x0 = np.clip(x, bounds["lower"], bounds["upper"])
+        optim = pybop.Optimisation(cost=cost)
+        results = pybop.OptimisationResult(x=x0, optim=optim, cost=cost)
 
         if np.all(x == np.asarray([0.05, 0.05])):
-            message = classify_using_Hessian(optim, optim_x)
+            message = classify_using_Hessian(results)
             assert message == "The optimiser has located a minimum."
         elif np.all(x == np.asarray([0.1, 0.05])):
-            message = classify_using_Hessian(optim, optim_x)
+            message = classify_using_Hessian(results)
             assert message == (
                 "The optimiser has not converged to a stationary point."
                 " The result is near the upper bound of R0 [Ohm]."
             )
         elif np.all(x == np.asarray([0.05, 0.01])):
-            message = classify_using_Hessian(optim, optim_x)
+            message = classify_using_Hessian(results)
             assert message == (
                 "The optimiser has not converged to a stationary point."
                 " The result is near the lower bound of R1 [Ohm]."
             )
         else:
-            raise Exception("Please add a check for these values.")
+            raise Exception(f"Please add a check for these values: {x}")
 
         if np.all(x == np.asarray([0.05, 0.05])):
             cost = pybop.GaussianLogLikelihoodKnownSigma(problem, sigma0=0.002)
-            optim = pybop.XNES(cost=cost)
+            optim = pybop.Optimisation(cost=cost)
+            results = pybop.OptimisationResult(x=x, optim=optim, cost=cost)
 
-            message = classify_using_Hessian(optim, x)
+            message = classify_using_Hessian(results)
             assert message == "The optimiser has located a maximum."
 
-        # message = classify_using_Hessian(optim, x)
+        # message = classify_using_Hessian(results)
         # assert message == "The optimiser has located a saddle point."
 
     @pytest.mark.integration
@@ -163,10 +164,11 @@ class TestClassification:
         ):
[…]

### 2024-12-18 pybop/_classification.py
Improve insensitivity tests

@@ -1,39 +1,45 @@
+from typing import Optional
+
 import numpy as np
 
 from pybop import OptimisationResult
 
 
-def classify_using_Hessian(result: OptimisationResult, epsilon=1e-2):
+def classify_using_Hessian(
+    result: OptimisationResult, dx=None, cost_tolerance: Optional[float] = 1e-5
+):
     """
     A simple check for parameter correlations based on numerical approximation
     of the Hessian matrix at the optimal point using central finite differences.
 
     Parameters
     ---------
-    epsilon : float, optional
     result : OptimisationResult
         The PyBOP optimisation results.
-        A small positive value used to check proximity to the parameter bounds and as the
-        perturbation distance used in the finite difference calculations (default: 1e-2).
+    dx : array-like, optional
+        An array of small positive values used to check proximity to the parameter
+        bounds and as the perturbation distance in the finite difference calculations.
+    cost_tolerance : float, optional
+        A small positive tolerance used for cost value comparisons (default: 1e-5).
     """
     x = result.x
+    dx = np.asarray(dx) if dx is not None else np.maximum(x, 1e-40) * 1e-2
     final_cost = result.final_cost
     cost = result.cost
     parameters = result.cost.parameters
     minimising = result.optim.minimising
 
     n = len(x)
-    if n != 2:
+    if n != 2 or len(dx) != n:
         raise ValueError(
-            "The function classify_using_Hessian currently only works"
-            " in the case of 2 parameters."
+            "The function classify_using_Hessian currently only works in the case "
+            "of 2 parameters, and dx must have the same length as x."
         )
 
     # Get a list of parameter names for use in the output message
     names = list(parameters.keys())
 
     # Evaluate the cost for a grid of surrounding points
-    dx = x * epsilon
     costs = np.zeros((3, 3, 2))
     for i in np.arange(0, 3):
         for j in np.arange(0, 3):
@@ -45,7 +51,10 @@ def classify_using_Hessian(result: OptimisationResult, epsilon=1e-2):
                 costs[i, j, 1] = cost(x + np.multiply([i - 1, j - 1], 2 * dx))
 
     # Classify the result
-    if (minimising and np.any(costs < final_cost)) or (
+    if not np.all([np.isfinite(cost) for cost in costs]):
+        message = "Classification cannot proceed due to infinite cost value(s)."
+
+    elif (minimising and np.any(costs < final_cost)) or (
[…]

### 2024-12-18 tests/integration/test_classification.py

@@ -112,34 +112,22 @@ class TestClassification:
 
     @pytest.mark.integration
     def test_insensitive_classify_using_Hessian(self, parameter_set):
-        param_R0 = pybop.Parameter(
-            "R0 [Ohm]",
+        param_R0_a = pybop.Parameter(
+            "R0_a [Ohm]",
             bounds=[0, 0.002],
             true_value=0.001,
         )
-        param_R0_mod = pybop.Parameter(
-            "R0 modification [Ohm]",
-            bounds=[-1e-4, 1e-4],
-            true_value=0,
-        )
-        param_R1_mod = pybop.Parameter(
-            "R1 modification [Ohm]",
+        param_R0_b = pybop.Parameter(
+            "R0_b [Ohm]",
             bounds=[-1e-4, 1e-4],
             true_value=0,
         )
         parameter_set.params.update(
-            {
-                "R0 modification [Ohm]": 0,
-                "R1 modification [Ohm]": 0,
-            },
+            {"R0_a [Ohm]": 0.001, "R0_b [Ohm]": 0},
             check_already_exists=False,
         )
-        R0, R1 = parameter_set["R0 [Ohm]"], parameter_set["R1 [Ohm]"]
         parameter_set.params.update(
-            {
-                "R0 [Ohm]": R0 + Parameter("R0 modification [Ohm]"),
-                "R1 [Ohm]": R1 + Parameter("R1 modification [Ohm]"),
-            }
+            {"R0 [Ohm]": Parameter("R0_a [Ohm]") + Parameter("R0_b [Ohm]")},
         )
         model = pybop.empirical.Thevenin(parameter_set=parameter_set)
 
@@ -155,35 +143,34 @@ class TestClassification:
             }
         )
 
-        for i, parameters in enumerate(
-            [
-                pybop.Parameters(param_R0_mod, param_R1_mod),
-                pybop.Parameters(param_R1_mod, param_R0),
-                pybop.Parameters(param_R0, param_R1_mod),
-            ]
-        ):
+        for parameters in [
+            pybop.Parameters(param_R0_b, param_R0_a),
+            pybop.Parameters(param_R0_a, param_R0_b),
+        ]:
             problem = pybop.FittingProblem(model, parameters, dataset)
-            cost = pybop.SumofPower(problem, p=3)
+            cost = pybop.SumofPower(problem, p=1)
             x = cost.parameters.true_value()
             optim = pybop.Optimisation(cost=cost)
[…]

### 2024-12-18 pybop/_classification.py
Improve correlation checks

@@ -112,20 +112,25 @@ def classify_using_Hessian(
             message = "The optimiser has located a maximum."
         elif np.all(np.abs(eigenvalues) > cost_tolerance):
             message = "The optimiser has located a saddle point."
+        elif np.all(np.abs(eigenvalues) < cost_tolerance):
+            message = f"The cost variation is smaller than the cost tolerance: {cost_tolerance}."
         else:
-            # At least one eigenvalue is too small to classify with certainty
-            if np.all(np.abs(eigenvalues) < cost_tolerance):
-                message = f"The cost variation is smaller than the cost tolerance: {cost_tolerance}."
-            elif np.allclose(eigenvectors[0], np.array([1, 0])):
-                message = (
-                    f"The cost is insensitive to a change of {dx[0]:.2g} in {names[0]}."
-                )
+            # One eigenvalue is too small to classify with certainty
+            message = "The cost variation is too small to classify with certainty."
+
+        # Check for parameter correlations
+        if np.any(np.abs(eigenvalues) > cost_tolerance):
+            if np.allclose(eigenvectors[0], np.array([1, 0])):
+                message += f" The cost is insensitive to a change of {dx[0]:.2g} in {names[0]}."
             elif np.allclose(eigenvectors[0], np.array([0, 1])):
-                message = (
-                    f"The cost is insensitive to a change of {dx[1]:.2g} in {names[1]}."
-                )
+                message += f" The cost is insensitive to a change of {dx[1]:.2g} in {names[1]}."
             else:
-                message = "There may be a correlation between these parameters."
+                diagonal_costs = [
+                    cost(x - np.multiply(eigenvectors[:, 0], dx)),
+                    cost(x + np.multiply(eigenvectors[:, 0], dx)),
+                ]
+                if np.allclose(final_cost, diagonal_costs, atol=cost_tolerance, rtol=0):
+                    message += " There may be a correlation between these parameters."
 
     print(message)
     return message

### 2024-12-18 tests/integration/test_classification.py

@@ -154,20 +154,21 @@ class TestClassification:
             results = pybop.OptimisationResult(x=x, optim=optim, cost=cost)
 
             message = classify_using_Hessian(results)
-            assert (
-                message == "The cost is insensitive to a change of 1e-42 in R0_b [Ohm]."
+            assert message == (
+                "The cost variation is too small to classify with certainty."
+                " The cost is insensitive to a change of 1e-42 in R0_b [Ohm]."
             )
 
-        results = pybop.OptimisationResult(x=[0.0008, 0.0002], optim=optim, cost=cost)
-
-        message = classify_using_Hessian(
-            results,
-            dx=[0.0001, 0.0001], cost_tolerance=1e-2
+        message = classify_using_Hessian(results, dx=[0.0001, 0.0001])
+        assert message == (
+            "The optimiser has located a minimum."
+            " There may be a correlation between these parameters."
         )
-        assert message == "There may be a correlation between these parameters."
 
         message = classify_using_Hessian(results, cost_tolerance=1e-2)
-        assert message == "The cost variation is smaller than the cost tolerance: 0.01."
+        assert message == (
+            "The cost variation is smaller than the cost tolerance: 0.01."
+        )
 
         message = classify_using_Hessian(results, dx=[1, 1])
         assert message == "Classification cannot proceed due to infinite cost value(s)."

### 2024-12-18 tests/unit/test_optimisation.py
unit test updated

@@ -276,7 +276,7 @@ class TestOptimisation:
             pybop.AdamW,
             pybop.CuckooSearch,
             pybop.GradientDescent,
-            # pybop.RandomSearch,
+            pybop.RandomSearch,
         ]:
             optim = optimiser(cost)
             with pytest.raises(
@@ -324,11 +324,11 @@ class TestOptimisation:
                 assert optim.optimiser.n_hyper_parameters() == 5
                 assert optim.optimiser.x_guessed() == optim.optimiser._x0
 
-            if optimiser is pybop.RandomSearch:
-                assert optimiser._dim == len(optim.optimiser._x0)
-                assert optimiser.x_best().shape == optim.optimiser._x0.shape
-                assert optimiser.f_best() == np.inf
-                assert optimiser._iterations == 0
+            # if optimiser is pybop.RandomSearch:
+            #     assert optimiser._dim == len(optim.optimiser._x0)
+            #     assert optimiser.x_best().shape == optim.optimiser._x0.shape
+            #     assert optimiser.f_best() == np.inf
+            #     assert optimiser._iterations == 0
         else:
             x0 = cost.parameters.initial_value()
             assert optim.x0 == x0
@@ -343,11 +343,11 @@ class TestOptimisation:
         optim.run()
         assert optim.optimiser._boundaries is None
 
-    # @pytest.mark.unit
-    # def test_randomsearch(self, cost):
-    #     optim = pybop.RandomSearch(cost=cost, max_iterations=1)
-    #     results = optim.run()
-    #     assert results.final_cost is not None
+    @pytest.mark.unit
+    def test_randomsearch(self, cost):
+        optim = pybop.RandomSearch(cost=cost, bounds=None, max_iterations=1)
+        optim.run()
+        assert optim.optimiser._boundaries is None
 
     @pytest.mark.unit
     def test_scipy_minimize_with_jac(self, cost):

### 2024-12-18 tests/unit/test_optimisation.py
unit test modified

@@ -324,11 +324,6 @@ class TestOptimisation:
                 assert optim.optimiser.n_hyper_parameters() == 5
                 assert optim.optimiser.x_guessed() == optim.optimiser._x0
 
-            # if optimiser is pybop.RandomSearch:
-            #     assert optimiser._dim == len(optim.optimiser._x0)
-            #     assert optimiser.x_best().shape == optim.optimiser._x0.shape
-            #     assert optimiser.f_best() == np.inf
-            #     assert optimiser._iterations == 0
         else:
             x0 = cost.parameters.initial_value()
             assert optim.x0 == x0

### 2024-12-18 pybop/costs/base_cost.py
Add documentation
Co-authored-by: Brady Planden <55357039+BradyPlanden@users.noreply.github.com>

@@ -32,7 +32,7 @@ class BaseCost:
     _de : float
         The gradient of the cost function to use if an error occurs during
         evaluation. Defaults to 1.0.
-    minimsing : bool, optional, default=True
+    minimising : bool, optional, default=True
         If False, switches the sign of the cost and gradient to perform maximisation
         instead of minimisation.
     """
@@ -104,7 +104,13 @@ class BaseCost:
         else:
             model_inputs = inputs
 
-        # Check whether we are maximising or minimising
+        # Check whether we are maximising or minimising via:
+        #| `minimising` | `self.minimising` | `for_optimiser` |
+        #|--------------|-------------------|-----------------|
+        #| `True`       | `True`            | `True`          |
+        #| `True`       | `True`            | `False`         |
+        #| `False`      | `False`           | `True`          |
+        #| `True`       | `False`           | `False`         |
         minimising = self.minimising or not for_optimiser
 
         # Validate inputs, update parameters

### 2024-12-18 pybop/costs/base_cost.py
style: pre-commit fixes

@@ -105,12 +105,12 @@ class BaseCost:
             model_inputs = inputs
 
         # Check whether we are maximising or minimising via:
-        #| `minimising` | `self.minimising` | `for_optimiser` |
-        #|--------------|-------------------|-----------------|
-        #| `True`       | `True`            | `True`          |
-        #| `True`       | `True`            | `False`         |
-        #| `False`      | `False`           | `True`          |
-        #| `True`       | `False`           | `False`         |
+        # | `minimising` | `self.minimising` | `for_optimiser` |
+        # |--------------|-------------------|-----------------|
+        # | `True`       | `True`            | `True`          |
+        # | `True`       | `True`            | `False`         |
+        # | `False`      | `False`           | `True`          |
+        # | `True`       | `False`           | `False`         |
         minimising = self.minimising or not for_optimiser
 
         # Validate inputs, update parameters

### 2024-12-18 pybop/costs/_weighted_cost.py
Set minimising for each cost and make a property of BaseOptimiser only (#588)
* Apply maximise and transform in BaseOptimiser

* Rename iterable cost to internal_cost

* Update CHANGELOG.md

* Move minimising onto the cost classes

* Remove no longer needed fg_user

* Update CHANGELOG.md

* Apply minimising to BaseLikelihood

* Update test_weighted_fitting_cost

* Fix order

* Add documentation

Co-authored-by: Brady Planden <55357039+BradyPlanden@users.noreply.github.com>

* style: pre-commit fixes

---------

Co-authored-by: Brady Planden <55357039+BradyPlanden@users.noreply.github.com>
Co-a

@@ -2,7 +2,7 @@ from typing import Optional, Union
 
 import numpy as np
 
-from pybop import BaseCost, BaseLikelihood, DesignCost
+from pybop import BaseCost
 
 
 class WeightedCost(BaseCost):
@@ -32,9 +32,6 @@ class WeightedCost(BaseCost):
         self.costs = [cost for cost in costs]
         if len(set(type(cost.problem) for cost in self.costs)) > 1:
             raise TypeError("All problems must be of the same class type.")
-        self.minimising = not any(
-            isinstance(cost, (BaseLikelihood, DesignCost)) for cost in self.costs
-        )
 
         # Check if weights are provided
         if weights is not None:
@@ -61,6 +58,14 @@ class WeightedCost(BaseCost):
 
         for cost in self.costs:
             self.join_parameters(cost.parameters)
+
+        # Apply the minimising property from each cost
+        for i, cost in enumerate(self.costs):
+            self.weights[i] = self.weights[i] * (1 if cost.minimising else -1)
+        if all(not cost.minimising for cost in self.costs):
+            # If all costs are maximising, convert the weighted cost to maximising
+            self.weights = -self.weights
+            self.minimising = False
 
         # Weighted costs do not use this functionality
         self._has_separable_problem = False

### 2024-12-18 pybop/optimisers/base_optimiser.py

@@ -4,15 +4,7 @@ from typing import Optional, Union
 import numpy as np
 from scipy.optimize import OptimizeResult
 
-from pybop import (
-    BaseCost,
-    BaseLikelihood,
-    DesignCost,
-    Inputs,
-    Parameter,
-    Parameters,
-    WeightedCost,
-)
+from pybop import BaseCost, Inputs, Parameter, Parameters
 
 
 class BaseOptimiser:
@@ -43,9 +35,6 @@ class BaseOptimiser:
         Not all methods will use this information.
     verbose : bool, optional
         If True, the optimisation progress is printed (default: False).
-    minimising : bool, optional
-        If True, the target is to minimise the cost, else target is to maximise by minimising
-        the negative cost (default: True).
     physical_viability : bool, optional
         If True, the feasibility of the optimised parameters is checked (default: False).
     allow_infeasible_solutions : bool, optional
@@ -62,13 +51,13 @@ class BaseOptimiser:
         # First set attributes to default values
         self.parameters = Parameters()
         self.x0 = optimiser_kwargs.get("x0", [])
-        self.log = dict(x=[], x_best=[], cost=[], cost_best=[], x0=[])
+        self.log = dict(x=[], x_best=[], x_search=[], x0=[], cost=[], cost_best=[])
         self.bounds = None
         self.sigma0 = 0.02
         self.verbose = True
-        self.minimising = True
         self._transformation = None
         self._needs_sensitivities = False
+        self._minimising = True
         self.physical_viability = False
         self.allow_infeasible_solutions = False
         self.default_max_iterations = 1000
@@ -79,10 +68,7 @@ class BaseOptimiser:
             self.parameters = self.cost.parameters
             self._transformation = self.cost.transformation
             self.set_allow_infeasible_solutions()
-            if isinstance(cost, WeightedCost):
-                self.minimising = cost.minimising
-            if isinstance(cost, (BaseLikelihood, DesignCost)):
-                self.minimising = False
+            self._minimising = self.cost.minimising
 
         else:
             try:
@@ -98,7 +84,6 @@ class BaseOptimiser:
                     self.parameters.add(
                         Parameter(name=f"Parameter {i}", initial_value=value)
                     )
-                self.minimising = True
 
             except Exception as e:
                 raise Exception(
@@ -147,7 +132,6 @@ class BaseOptimiser:
 
         # Set other options
         self.verbose = self.unset_options.pop("verbose", self.verbose)
[…]

### 2024-12-18 pybop/optimisers/base_pints_optimiser.py

@@ -198,13 +198,7 @@ class BasePintsOptimiser(BaseOptimiser):
 
         # Choose method to evaluate
         def fun(x):
-            if self._needs_sensitivities:
-                L, dl = self.cost(x, calculate_grad=True, apply_transform=True)
-            else:
-                L = self.cost(x, apply_transform=True)
-                dl = None
-            sign = -1 if not self.minimising else 1
-            return (sign * L, sign * dl) if dl is not None else sign * L
+            return self.cost_call(x, calculate_grad=self._needs_sensitivities)
 
         # Create evaluator object
         if self._parallel:
@@ -221,9 +215,6 @@ class BasePintsOptimiser(BaseOptimiser):
 
         # Keep track of current best and best-guess scores.
         fb = fg = np.inf
-
-        # Internally we always minimise! Keep a 2nd value to show the user.
-        fg_user = (fb, fg) if self.minimising else (-fb, -fg)
 
         # Keep track of the last significant change
         f_sig = np.inf
@@ -244,7 +235,6 @@ class BasePintsOptimiser(BaseOptimiser):
                 # Update the scores
                 fb = self.optimiser.f_best()
                 fg = self.optimiser.f_guessed()
-                fg_user = (fb, fg) if self.minimising else (-fb, -fg)
 
                 # Check for significant changes against the absolute and relative tolerance
                 f_new = fg if self._use_f_guessed else fb
@@ -263,8 +253,8 @@ class BasePintsOptimiser(BaseOptimiser):
                 self.log_update(
                     x=xs,
                     x_best=self.optimiser.x_best(),
-                    cost=_fs if self.minimising else [-x for x in _fs],
-                    cost_best=fb if self.minimising else -fb,
+                    cost=_fs,
+                    cost_best=fb,
                 )
 
                 # Check stopping criteria:
@@ -327,14 +317,11 @@ class BasePintsOptimiser(BaseOptimiser):
             # Show last result and exit
             print("\n" + "-" * 40)
             print("Unexpected termination.")
-            print("Current score: " + str(fg_user))
+            print("Current score: " + str((fb, fg)))
             print("Current position:")
 
-            # Show current parameters
-            x_user = self.optimiser.x_guessed()
-            if self._transformation:
-                x_user = self._transformation.to_model(x_user)
-            for p in x_user:
+            # Show current parameters (with any transformation applied)
+            for p in self.optimiser.x_guessed():
                 print(PintsStrFloat(p))
[…]

### 2024-12-18 pybop/experimental/multivariate_priors.py
draft of multivariate priors including multivariate normal

@@ -1,0 +1,171 @@
+import numpy as np
+import scipy.integrate as integrate
+import scipy.stats as stats
+from scipy.linalg import sqrtm
+
+from pybop import BasePrior
+
+
+def insert_x_at_dim(other, x, dim):
+    """
+    Inserts each element of x at dim in other.
+
+    Parameters
+    ----------
+    other : numpy.ndarray
+        The fixed other variables in all dimensions but dim.
+    x : numpy.ndarray
+        The point(s) at which to evaluate in dim.
+    dim : int
+        The dimension at which to insert x.
+    """
+    x = np.atleast_1d(x)
+    return np.array(
+        [np.concatenate(other[: dim - 1], [entry], other[dim - 1 :]) for entry in x]
+    )
+
+
+class BaseMultivariatePrior(BasePrior):
+    """
+    A base class for defining multivariate prior distributions.
+
+    This class extends ``pybop.BasePrior`` with methods that reduce the
+    output of a multivariate distribution to its individual dimensions.
+
+    Note that, unlinke in ``pybop.BasePrior``, distribution attributes
+    are stored in a dictionary, as multivariate distributiosn in SciPy
+    do not follow the loc/scale convention.
+
+    Attributes
+    ----------
+    prior : scipy.stats.rv_continuous
+        The underlying continuous random variable distribution.
+    properties : dict
+        A dictionary with distribution keyword argument names as string
+        keys and their values as float values.
+    """
+
+    def pdf(self, x):
+        return self.prior.pdf(x, **self.properties)
+
+    def pdf_marginal(self, x, dim):
+        """
+        Integrates the probability density function (PDF) at x down to
+        one variable.
+
+        Parameters
+        ----------
+        x : numpy.ndarray
+            The point(s) at which to evaluate the pdf.
+        dim : int
+            The dimension to which to reduce the pdf to.
+
+        Returns
+        -------
+        float
+            The marginal probability density function value at x in dim.
+        """
+        unnormalized_pdf = integrate.nquad(
+            lambda y: self.pdf(insert_x_at_dim(y, x, dim)),
+            [
+                [-np.inf, np.inf] * (dim - 1)
+                + [-np.inf, np.inf] * (self.prior.dim - dim)
+            ],
+        )[0]
+        return unnormalized_pdf / np.sum(unnormalized_pdf)
+
+    def logpdf(self, x):
+        return self.prior.logpdf(x, **self.properties)
+
+    def logpdf_marginal(self, x, dim):
+        """
[…]

### 2024-12-18 pybop/experimental/multivariate_parameters.py
draft of Parameters extension for multivariate priors

@@ -1,0 +1,96 @@
+import numpy as np
+
+from pybop import MultivariatePrior, Parameters
+
+
+class MultivariateParameters(Parameters):
+    """
+    Represents a correlated set of uncertain parameters within the PyBOP
+    framework.
+
+    This class encapsulates the definition of each of its parameters,
+    including their names, bounds, and margins to ensure the parameter
+    stays within feasible limits during optimisation or sampling.
+
+    Parameters
+    ----------
+    parameter_list : pybop.Parameter or Dict
+    prior : pybop.MultivariatePrior
+        The joint multivariate prior.
+    """
+
+    def __init__(self, *args, prior=None):
+        self.prior = prior
+        super().__init__(*args)
+        for param in self.param.values():
+            # Ensure that no individual priors are mixed with the joint
+            # one. They may have been used for setting boundaries.
+            param.prior = None
+
+    def get_margins(self) -> list:
+        """
+        Collects the margins of all parameters.
+
+        Returns
+        -------
+            array-like
+                A list of the margin attributes of each parameter.
+        """
+        return [param.margin for param in self.param.values()]
+
+    def rvs(
+        self, n_samples: int = 1, random_state=None, apply_transform: bool = False
+    ) -> np.ndarray:
+        """
+        Draw random samples from the joint parameters prior distribution.
+
+        The samples are constrained to be within each parameter's bounds,
+        excluding a pre-defined margin at the boundaries.
+
+        Parameters
+        ----------
+        n_samples : int
+            The number of samples to draw (default: 1).
+        random_state : int, optional
+            The random state seed for reproducibility (default: None).
+        apply_transform : bool
+            If True, the transformation is applied to the output
+            (default: False).
+
+        Returns
+        -------
+            array-like
+                A matrix (i.e., a 2D array) of samples drawn from the
+                joint prior distribution inside parameter boundaries.
+        """
+        samples = self.prior.rvs(n_samples, random_state=random_state)
+
+        # Constrain samples to be within bounds.
+        bounds = self.get_bounds(apply_transform=False)
+        margins = self.get_margins()
+        for i in len(samples):
+            offset = margins[i] * (bounds["upper"][i] - bounds["lower"][i])
+            samples[i] = np.clip(
+                samples[i], bounds["lower"][i] + offset, bounds["upper"][i] + offset
+            )
+
+        transformations = self.get_transformations()
+        if apply_transform:
+            for i in len(samples):
+                if transformations[i]:
[…]

### 2024-12-18 pybop/_classification.py
Update in line with OptimisationResult

@@ -26,8 +26,8 @@ def classify_using_Hessian(
     dx = np.asarray(dx) if dx is not None else np.maximum(x, 1e-40) * 1e-2
     final_cost = result.final_cost
     cost = result.cost
-    parameters = result.cost.parameters
-    minimising = result.optim.minimising
+    parameters = cost.parameters
+    minimising = result.minimising
 
     n = len(x)
     if n != 2 or len(dx) != n:

### 2024-12-18 tests/integration/test_classification.py

@@ -79,7 +79,7 @@ class TestClassification:
         bounds = cost.parameters.get_bounds()
         x0 = np.clip(x, bounds["lower"], bounds["upper"])
         optim = pybop.Optimisation(cost=cost)
-        results = pybop.OptimisationResult(x=x0, optim=optim, cost=cost)
+        results = pybop.OptimisationResult(x=x0, optim=optim)
 
         if np.all(x == np.asarray([0.05, 0.05])):
             message = classify_using_Hessian(results)
@@ -102,7 +102,7 @@ class TestClassification:
         if np.all(x == np.asarray([0.05, 0.05])):
             cost = pybop.GaussianLogLikelihoodKnownSigma(problem, sigma0=0.002)
             optim = pybop.Optimisation(cost=cost)
-            results = pybop.OptimisationResult(x=x, optim=optim, cost=cost)
+            results = pybop.OptimisationResult(x=x, optim=optim)
 
             message = classify_using_Hessian(results)
             assert message == "The optimiser has located a maximum."
@@ -151,7 +151,7 @@ class TestClassification:
             cost = pybop.SumofPower(problem, p=1)
             x = cost.parameters.true_value()
             optim = pybop.Optimisation(cost=cost)
-            results = pybop.OptimisationResult(x=x, optim=optim, cost=cost)
+            results = pybop.OptimisationResult(x=x, optim=optim)
 
             message = classify_using_Hessian(results)
             assert message == (
@@ -180,7 +180,7 @@ class TestClassification:
         cost = pybop.SumSquaredError(problem)
         x = cost.parameters.true_value()
         optim = pybop.Optimisation(cost=cost)
-        results = pybop.OptimisationResult(x=x, optim=optim, cost=cost)
+        results = pybop.OptimisationResult(x=x, optim=optim)
 
         with pytest.raises(
             ValueError,

### 2024-12-18 examples/scripts/comparison_examples/random_search.py
suggested changes incorporated

@@ -70,4 +70,4 @@ pybop.plot.convergence(optim)
 pybop.plot.parameters(optim)
 
 # Plot the cost landscape with optimisation path
-pybop.plot.contour(optim, steps=15)
+pybop.plot.contour(optim, steps=10)

### 2024-12-18 pybop/optimisers/_random_search.py

@@ -25,10 +25,10 @@ class RandomSearchImpl(PopulationBasedOptimiser):
     """
 
     def __init__(self, x0, sigma0=0.05, boundaries=None):
-        # Problem dimensionality
-        self._dim = len(x0)  # Initialize _dim first
-
         super().__init__(x0, sigma0, boundaries=boundaries)
+
+        # Problem dimensionality
+        self._dim = len(x0)
 
         # Optimisation parameters
         self._n = self._population_size

### 2024-12-18 pybop/optimisers/_random_search.py
suggested changes updated

@@ -7,21 +7,21 @@ class RandomSearchImpl(PopulationBasedOptimiser):
     Random Search (RS) optimisation algorithm.
     This algorithm explores the parameter space by randomly sampling points.
 
-    The algorithm is simple:
+    The algorithm does the following:
     1. Initialise a population of solutions.
-    2. At each iteration, generate new random solutions within boundaries.
-    3. Evaluate the quality/fitness of the solutions.
-    4. Update the best solution found so far.
+    2. At each iteration, generate `n` number of random positions within boundaries.
+    3. Evaluate the quality/fitness of the positions.
+    4. Replace the best position with improved position if found.
 
     Parameters:
-    - population_size (optional): Number of solutions to evaluate per iteration.
+        population_size (optional): Number of solutions to evaluate per iteration.
 
     References:
-    - The Random Search algorithm implemented in this work is based on principles outlined
+    The Random Search algorithm implemented in this work is based on principles outlined
     in "Introduction to Stochastic Search and Optimization: Estimation, Simulation, and
     Control" by Spall, J. C. (2003).
-    The implementation leverages the pints library framework, which provides tools for
-    population-based optimization methods.
+
+    The implementation inherits from the PINTS PopulationOptimiser.
     """
 
     def __init__(self, x0, sigma0=0.05, boundaries=None):
@@ -32,18 +32,6 @@ class RandomSearchImpl(PopulationBasedOptimiser):
 
         # Optimisation parameters
         self._n = self._population_size
-        self._iterations = 0
-
-        # Initialise population
-        self._candidates = (
-            np.random.uniform(
-                low=self._boundaries.lower(),
-                high=self._boundaries.upper(),
-                size=(self._n, self._dim),
-            )
-            if self._boundaries
-            else np.random.normal(self._x0, self._sigma0, size=(self._n, self._dim))
-        )
 
         # Initialise best solution
         self._x_best = np.copy(x0)
@@ -53,35 +41,32 @@ class RandomSearchImpl(PopulationBasedOptimiser):
 
     def ask(self):
         """
-        Returns a list of next points in the parameter-space
-        to evaluate from the optimiser.
+        Returns a list of positions to evaluate in the optimiser-space.
         """
         self._ready_for_tell = True
         self._running = True
 
-        # Generate random solutions within the boundaries
+        # Generate random solutions
         if self._boundaries:
             self._candidates = np.random.uniform(
                 low=self._boundaries.lower(),
                 high=self._boundaries.upper(),
[…]

### 2024-12-18 pybop/optimisers/pints_optimisers.py

@@ -662,6 +662,9 @@ class RandomSearch(BasePintsOptimiser):
     Random Search is a simple optimisation algorithm that samples parameter sets randomly
     within the given boundaries and identifies the best solution based on fitness.
 
+    This optimiser has been implemented for benchmarking and comparisons, convergence will be
+    better with one of other optimisers in the majority of cases.
+
     Parameters
     ----------
     cost : callable
@@ -672,6 +675,9 @@ class RandomSearch(BasePintsOptimiser):
         Minimum number of iterations before termination.
     max_unchanged_iterations : int, optional (default=15)
         Maximum number of iterations without improvement before termination.
+    multistart : int, optional (default=1)
+        Number of optimiser restarts from randomly sample position. These positions
+        are sampled from the priors.
     parallel : bool, optional (default=False)
         Whether to run the optimisation in parallel.
     **optimiser_kwargs : optional

### 2024-12-18 pybop/costs/fitting_costs.py
Tidy up the cost classes to avoid repetition

@@ -1,4 +1,4 @@
-from typing import Union
+from typing import Optional, Union
 
 import numpy as np
 
@@ -6,16 +6,16 @@ from pybop.costs.base_cost import BaseCost
 from pybop.observers.observer import Observer
 
 
-class RootMeanSquaredError(BaseCost):
+class FittingCost(BaseCost):
     """
-    Root mean square error cost function.
+    Overwrites and extends `BaseCost` class for fitting-type cost functions.
 
-    Computes the root mean square error between model predictions and the target
-    data, providing a measure of the differences between predicted values and
-    observed values.
+    This class is intended to be subclassed to create specific cost functions
+    for evaluating model predictions against a set of data. The cost function
+    quantifies the goodness-of-fit between the model predictions and the
+    observed data, with a lower cost value indicating a better fit.
 
     Inherits all parameters and attributes from ``BaseCost``.
-
     """
 
     def __init__(self, problem):
@@ -24,8 +24,7 @@ class RootMeanSquaredError(BaseCost):
     def compute(
         self,
         y: dict,
-        dy: np.ndarray = None,
-        calculate_grad: bool = False,
+        dy: Optional[np.ndarray] = None,
     ) -> Union[float, tuple[float, np.ndarray]]:
         """
         Computes the cost function for the given predictions.
@@ -36,26 +35,76 @@ class RootMeanSquaredError(BaseCost):
             The dictionary of predictions with keys designating the signals for fitting.
         dy : np.ndarray, optional
             The corresponding gradient with respect to the parameters for each signal.
-        calculate_grad : bool, optional
-            A bool condition designating whether to calculate the gradient.
 
         Returns
         -------
-        float
-            The root mean square error.
+        tuple or float
+            If dy is not None, returns a tuple containing the cost (float) and the
+            gradient (np.ndarray), otherwise returns only the computed cost (float).
         """
-        # Verify we have dy if calculate_grad is True
-        self.verify_args(dy, calculate_grad)
-
         # Early return if the prediction is not verified
         if not self.verify_prediction(y):
-            return (np.inf, self.grad_fail) if calculate_grad else np.inf
+            return (np.inf, self.grad_fail) if dy is not None else np.inf
 
-        # Calculate residuals and error
+        # Compute the residual
         r = np.asarray([y[signal] - self._target[signal] for signal in self.signal])
+
+        return self._error_measure(r=r, dy=dy)
+
[…]

### 2024-12-18 tests/unit/test_cost.py

@@ -172,14 +172,6 @@ class TestCosts:
         with pytest.raises(TypeError, match="Inputs must be a dictionary or numeric."):
             cost(["StringInputShouldNotWork"])
 
-        # Test ValueError for none dy w/ calculate_grad == True
-        if not isinstance(cost, pybop.ObserverCost):
-            with pytest.raises(
-                ValueError,
-                match="Forward model sensitivities need to be provided alongside `calculate_grad=True` for `cost.compute`.",
-            ):
-                cost.compute([1.1], dy=None, calculate_grad=True)
-
     @pytest.mark.unit
     def test_minkowski(self, problem):
         # Incorrect order

### 2024-12-18 pybop/costs/fitting_costs.py
Update tests

@@ -66,7 +66,7 @@ class FittingCost(BaseCost):
         dy : np.ndarray, optional
             The corresponding gradient with respect to the parameters for each signal.
         """
-        return NotImplementedError
+        raise NotImplementedError
 
 
 class RootMeanSquaredError(FittingCost):

### 2024-12-18 tests/unit/test_cost.py

@@ -112,11 +112,13 @@ class TestCosts:
             )
 
     @pytest.mark.unit
-    def test_base(self, problem):
-        base_cost = pybop.BaseCost(problem)
-        assert base_cost.problem == problem
-        with pytest.raises(NotImplementedError):
-            base_cost([0.5])
+    def test_base(self, model, parameters, dataset):
+        problem = pybop.FittingProblem(model, parameters, dataset)
+        for cost_class in [pybop.BaseCost, pybop.FittingCost, pybop.DesignCost]:
+            base_cost = cost_class(problem)
+            assert base_cost.problem == problem
+            with pytest.raises(NotImplementedError):
+                base_cost([0.5])
 
     @pytest.mark.unit
     def test_costs(self, cost):
@@ -229,7 +231,6 @@ class TestCosts:
     @pytest.mark.parametrize(
         "cost_class, expected_name",
         [
-            (pybop.DesignCost, "Design Cost"),
             (pybop.GravimetricEnergyDensity, "Gravimetric Energy Density"),
             (pybop.VolumetricEnergyDensity, "Volumetric Energy Density"),
             (pybop.GravimetricPowerDensity, "Gravimetric Power Density"),
@@ -242,33 +243,25 @@ class TestCosts:
         cost = cost_class(design_problem)
         assert cost.name == expected_name
 
-        if cost_class in [pybop.DesignCost]:
-            with pytest.raises(NotImplementedError):
-                cost([0.5])
-        else:
-            # Test type of returned value
-            assert np.isscalar(cost([0.5]))
-            assert cost([0.4]) >= 0  # Should be a viable design
-            assert (
-                cost([0.8]) == -np.inf
-            )  # Should exceed active material + porosity < 1
-            assert cost([1.4]) == -np.inf  # Definitely not viable
-            assert cost([-0.1]) == -np.inf  # Should not be a viable design
+        # Test type of returned value
+        assert np.isscalar(cost([0.5]))
+        assert cost([0.4]) >= 0  # Should be a viable design
+        assert cost([0.8]) == -np.inf  # Should exceed active material + porosity < 1
+        assert cost([1.4]) == -np.inf  # Definitely not viable
+        assert cost([-0.1]) == -np.inf  # Should not be a viable design
 
-            # Test infeasible locations
-            cost.problem.model.allow_infeasible_solutions = False
-            assert cost([1.1]) == -np.inf
+        # Test infeasible locations
+        cost.problem.model.allow_infeasible_solutions = False
[…]

### 2024-12-18 tests/unit/test_optimisation.py
unit tests for RandomSearch added

@@ -339,10 +339,35 @@ class TestOptimisation:
         assert optim.optimiser._boundaries is None
 
     @pytest.mark.unit
-    def test_randomsearch(self, cost):
-        optim = pybop.RandomSearch(cost=cost, bounds=None, max_iterations=1)
-        optim.run()
-        assert optim.optimiser._boundaries is None
+    def test_randomsearch_bounds(self, two_param_cost):
+        # Initialize RandomSearch with boundaries
+        bounds = {"upper": [0.62,0.57], "lower": [0.58,0.53]}
+
+        optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=bounds, max_iterations=1)
+
+        # Define candidates outside boundaries
+        candidates = np.array([[0.57, 0.52], [0.63, 0.58]])
+
+        # Use clip_candidates to clip to boundaries
+        clipped_candidates = optimiser.optimiser.clip_candidates(candidates)
+
+        # Expected clipped candidates
+        expected_clipped = np.array([[0.58, 0.53], [0.62, 0.57]])
+
+        # Assert that the clipped candidates match the expected values
+        assert np.allclose(clipped_candidates, expected_clipped)
+
+        # Initialize optimiser without boundaries
+        optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=None, max_iterations=1)
+
+        # Define candidates outside typical boundaries
+        candidates = np.array([[0.57, 0.52], [0.63, 0.58]])
+
+        # Use clip_candidates, which should return the input unmodified
+        clipped_candidates = optimiser.optimiser.clip_candidates(candidates)
+
+        # Assert that the clipped candidates are unchanged
+        assert np.allclose(clipped_candidates, candidates)
 
     @pytest.mark.unit
     def test_scipy_minimize_with_jac(self, cost):

### 2024-12-18 tests/unit/test_optimisation.py
style: pre-commit fixes

@@ -341,9 +341,11 @@ class TestOptimisation:
     @pytest.mark.unit
     def test_randomsearch_bounds(self, two_param_cost):
         # Initialize RandomSearch with boundaries
-        bounds = {"upper": [0.62,0.57], "lower": [0.58,0.53]}
+        bounds = {"upper": [0.62, 0.57], "lower": [0.58, 0.53]}
 
-        optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=bounds, max_iterations=1)
+        optimiser = pybop.RandomSearch(
+            cost=two_param_cost, bounds=bounds, max_iterations=1
+        )
 
         # Define candidates outside boundaries
         candidates = np.array([[0.57, 0.52], [0.63, 0.58]])
@@ -358,7 +360,9 @@ class TestOptimisation:
         assert np.allclose(clipped_candidates, expected_clipped)
 
         # Initialize optimiser without boundaries
-        optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=None, max_iterations=1)
+        optimiser = pybop.RandomSearch(
+            cost=two_param_cost, bounds=None, max_iterations=1
+        )
 
         # Define candidates outside typical boundaries
         candidates = np.array([[0.57, 0.52], [0.63, 0.58]])

### 2024-12-18 tests/unit/test_optimisation.py
unit tests modified

@@ -340,33 +340,18 @@ class TestOptimisation:
 
     @pytest.mark.unit
     def test_randomsearch_bounds(self, two_param_cost):
-        # Initialize RandomSearch with boundaries
+        # Test clip_candidates with bound
         bounds = {"upper": [0.62,0.57], "lower": [0.58,0.53]}
-
         optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=bounds, max_iterations=1)
-
-        # Define candidates outside boundaries
         candidates = np.array([[0.57, 0.52], [0.63, 0.58]])
-
-        # Use clip_candidates to clip to boundaries
         clipped_candidates = optimiser.optimiser.clip_candidates(candidates)
-
-        # Expected clipped candidates
         expected_clipped = np.array([[0.58, 0.53], [0.62, 0.57]])
-
-        # Assert that the clipped candidates match the expected values
         assert np.allclose(clipped_candidates, expected_clipped)
 
-        # Initialize optimiser without boundaries
+        # Test clip_candidates without bound
         optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=None, max_iterations=1)
-
-        # Define candidates outside typical boundaries
         candidates = np.array([[0.57, 0.52], [0.63, 0.58]])
-
-        # Use clip_candidates, which should return the input unmodified
         clipped_candidates = optimiser.optimiser.clip_candidates(candidates)
-
-        # Assert that the clipped candidates are unchanged
         assert np.allclose(clipped_candidates, candidates)
 
     @pytest.mark.unit

### 2024-12-18 tests/unit/test_optimisation.py
style: pre-commit fixes

@@ -341,15 +341,19 @@ class TestOptimisation:
     @pytest.mark.unit
     def test_randomsearch_bounds(self, two_param_cost):
         # Test clip_candidates with bound
-        bounds = {"upper": [0.62,0.57], "lower": [0.58,0.53]}
-        optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=bounds, max_iterations=1)
+        bounds = {"upper": [0.62, 0.57], "lower": [0.58, 0.53]}
+        optimiser = pybop.RandomSearch(
+            cost=two_param_cost, bounds=bounds, max_iterations=1
+        )
         candidates = np.array([[0.57, 0.52], [0.63, 0.58]])
         clipped_candidates = optimiser.optimiser.clip_candidates(candidates)
         expected_clipped = np.array([[0.58, 0.53], [0.62, 0.57]])
         assert np.allclose(clipped_candidates, expected_clipped)
 
         # Test clip_candidates without bound
-        optimiser = pybop.RandomSearch(cost=two_param_cost, bounds=None, max_iterations=1)
+        optimiser = pybop.RandomSearch(
+            cost=two_param_cost, bounds=None, max_iterations=1
+        )
         candidates = np.array([[0.57, 0.52], [0.63, 0.58]])
         clipped_candidates = optimiser.optimiser.clip_candidates(candidates)
         assert np.allclose(clipped_candidates, candidates)

### 2024-12-20 papers/joss/paper.preprint.tex
Updates figures, /joss directory -> /papers

--- 2024-12-18 joss/paper.preprint.tex
@@
 ]{article}
+\usepackage[a4paper, total={6in, 8in}]{geometry}
 \usepackage{amsmath,amssymb}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\date{04 December 2024}
+\date{19 December 2024}
 
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.png}
-\caption{The fitted synthetic dataset (left) and cost landscape (right)
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/sim-landscape.pdf}
+\caption{The synthetic fitting dataset (left) and cost landscape (right)
 for an example time-series battery model parameterisation using a
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/impedance.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/impedance.pdf}
 \caption{The data and model fit (left) and cost landscape (right) for a
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/converge.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/converge.pdf}
 \caption{Optimiser convergence using various cost (left) and likelihood
@@
 
-\begin{figure}
+\begin{figure}[htb]
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/optimisers_parameters.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/optimisers_parameters.pdf}
 \caption{Convergence in parameter values for several optimisation
@@
 
-\begin{figure}
+\begin{figure}[htb]
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_subplot.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/contour_subplot.pdf}
 \caption{Cost landscape contour plot with corresponding optimisation
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/posteriors.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/posteriors.pdf}
 \caption{Posterior distributions of model parameters alongside
@@
 \centering
-\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/design.png}
+\includegraphics[width=1\textwidth,height=\textheight]{figures/joss/design.pdf}
 \caption{Gravimetric energy density cost landscape alongside the initial

### 2024-12-20 papers/joss/design_plots.py

@@ -98,7 +98,7 @@ optim = pybop.NelderMead(
     parallel=True,
     allow_infeasible_solutions=False,
     max_iterations=250,
-    max_unchanged_iterations=35,
+    max_unchanged_iterations=45,
     sigma0=0.1,
 )
 
@@ -112,10 +112,21 @@ if create_plot["gravimetric"]:
     # Plot the cost landscape with optimisation path
     gravimetric_fig = pybop.plot.contour(
         optim,
-        steps=60,
+        steps=45,
+        show=False,
+        yaxis=dict(titlefont_size=15, tickfont_size=15),
+        xaxis=dict(titlefont_size=15, tickfont_size=15),
+        legend=dict(
+            orientation="h",
+            yanchor="bottom",
+            y=1.02,
+            xanchor="right",
+            x=1,
+            font=dict(size=15),
+        ),
+        coloraxis_colorbar=dict(tickfont=dict(size=18)),
+        margin=dict(t=50),
         title=None,
-        xaxis=dict(titlefont_size=14, tickfont_size=14),
-        yaxis=dict(titlefont_size=14, tickfont_size=14),
     )
     gravimetric_fig.write_image("joss/figures/design_gravimetric.pdf")
     time.sleep(3)
@@ -129,15 +140,39 @@ if create_plot["prediction"]:
         problem,
         problem_inputs=result.x,
         title=None,
-        width=576,
-        height=576,
-        legend=dict(orientation="h", yanchor="bottom", y=1.02, xanchor="right", x=1),
-        xaxis=dict(titlefont_size=14, tickfont_size=14),
-        yaxis=dict(titlefont_size=14, tickfont_size=14),
+        width=585,
+        height=585,
+        legend=dict(
+            orientation="h",
+            yanchor="bottom",
+            y=1.02,
+            xanchor="right",
+            x=1,
+            font=dict(size=14),
+        ),
+        yaxis=dict(
+            showline=True,
+            linewidth=2,
+            linecolor="black",
+            mirror=True,
+            titlefont_size=15,
+            tickfont_size=15,
+        ),
+        xaxis=dict(
+            showline=True,
+            linewidth=2,
+            linecolor="black",
+            mirror=True,
+            titlefont_size=15,
+            tickfont_size=15,
+        ),
+        margin=dict(r=50, t=50),
         show=False,
     )
+
[…]

### 2024-12-20 papers/joss/image_combine.sh

@@ -1,6 +1,6 @@
 pdfjam --nup 2x1 figures/design_prediction.pdf figures/design_gravimetric.pdf --landscape --outfile figures/joss/design.pdf --papersize '{450px,900px}'
 
-pdfjam --nup 3x1 figures/gradient_parameters.pdf figures/evolution_parameters.pdf figures/heuristic_parameters.pdf --landscape --outfile figures/joss/optimisers_parameters.pdf --papersize '{768px,1296px}'
+pdfjam --nup 3x1 figures/gradient_parameters.pdf figures/evolution_parameters.pdf figures/heuristic_parameters.pdf --landscape --outfile figures/joss/optimisers_parameters.pdf --papersize '{920px,1440px}'
 
 pdfjam --nup 2x1 figures/convergence_minimising.pdf figures/convergence_maximising.pdf --landscape --outfile figures/joss/converge.pdf --papersize '{450px,900px}'
 

### 2024-12-20 papers/joss/param_plots.py

@@ -19,14 +19,14 @@ np.random.seed(8)
 # Choose which plots to show and save
 create_plot = {}
 create_plot["simulation"] = True
-create_plot["landscape"] = True
-create_plot["minimising"] = True
-create_plot["maximising"] = True
+create_plot["landscape"] = False
+create_plot["minimising"] = False
+create_plot["maximising"] = False
 create_plot["gradient"] = True
 create_plot["evolution"] = True
 create_plot["heuristic"] = True
-create_plot["posteriors"] = True
-create_plot["eis"] = True
+create_plot["posteriors"] = False
+create_plot["eis"] = False
 
 # Parameter set and model definition
 parameter_set = pybop.ParameterSet.pybamm("Chen2020")
@@ -80,19 +80,43 @@ if create_plot["simulation"]:
         x=solution["Time [s]"].data,
         y=[corrupt_values, solution["Battery open-circuit voltage [V]"].data, values],
         trace_names=[
-            "Voltage w/ Gaussian noise [V]",
-            "Open-circuit voltage [V]",
-            "Voltage [V]",
+            "Voltage w/ noise",
+            "Open-circuit voltage",
+            "Voltage",
         ],
     )
     simulation_plot_dict.traces[0].mode = "markers"
     simulation_fig = simulation_plot_dict(show=False)
     simulation_fig.update_layout(
+        width=595,
+        height=595,
         xaxis_title="Time / s",
         yaxis_title="Voltage / V",
-        width=576,
-        height=576,
-        legend_traceorder="reversed",
+        yaxis=dict(
+            showline=True,
+            linewidth=2,
+            linecolor="black",
+            mirror=True,
+            titlefont_size=15,
+            tickfont_size=15,
+        ),
+        xaxis=dict(
+            showline=True,
+            linewidth=2,
+            linecolor="black",
+            mirror=True,
+            titlefont_size=15,
+            tickfont_size=15,
+        ),
+        legend=dict(
+            orientation="h",
+            yanchor="bottom",
+            y=1.02,
+            xanchor="right",
+            x=1,
+            font=dict(size=15),
+        ),
+        margin=dict(r=50, t=50),
     )
     simulation_fig.show()
     simulation_fig.write_image("joss/figures/simulation.pdf")
@@ -118,9 +142,21 @@ if create_plot["landscape"]:
[…]

### 2024-12-20 pybop/plot/nyquist.py

@@ -80,11 +80,7 @@ def nyquist(problem, problem_inputs: Inputs = None, show=True, **layout_kwargs):
                 scaleratio=1,
             ),
             legend=dict(
-                x=0.02,
-                y=0.98,
-                bgcolor="rgba(255, 255, 255, 0.5)",
-                bordercolor="black",
-                borderwidth=1,
+                orientation="h", yanchor="bottom", y=1.02, xanchor="right", x=1
             ),
             width=600,
             height=600,
@@ -99,8 +95,8 @@ def nyquist(problem, problem_inputs: Inputs = None, show=True, **layout_kwargs):
 
         plot_dict.traces[0].update(
             mode="lines+markers",
-            line=dict(color="blue", width=2),
-            marker=dict(size=8, color="blue", symbol="circle"),
+            line=dict(color="#00CC96", width=2),
+            marker=dict(size=8, color="#00CC96", symbol="circle"),
         )
 
         target_trace = plot_dict.create_trace(
@@ -108,26 +104,12 @@ def nyquist(problem, problem_inputs: Inputs = None, show=True, **layout_kwargs):
             y=-target_output[i].imag,
             name="Reference",
             mode="markers",
-            marker=dict(size=8, color="red", symbol="circle-open"),
+            marker=dict(size=8, color="#636EFA", symbol="circle-open"),
             showlegend=True,
         )
         plot_dict.traces.append(target_trace)
 
         fig = plot_dict(show=False)
-
-        # Add minor gridlines
-        fig.update_xaxes(
-            showgrid=True,
-            gridwidth=1,
-            gridcolor="lightgray",
-            zerolinecolor="lightgray",
-        )
-        fig.update_yaxes(
-            showgrid=True,
-            gridwidth=1,
-            gridcolor="lightgray",
-            zerolinecolor="lightgray",
-        )
 
         # Overwrite with user-kwargs
         fig.update_layout(**layout_kwargs)

### 2024-12-20 examples/scripts/comparison_examples/maximum_likelihood.py
Adds coverage, updates evaluation, and improves integration between BaseCost & BaseJaxCost

@@ -8,7 +8,7 @@ parameter_set = pybop.ParameterSet.pybamm("Chen2020")
 parameter_set.update(
     {
         "Negative electrode active material volume fraction": 0.63,
-        "Positive electrode active material volume fraction": 0.62,
+        "Positive electrode active material volume fraction": 0.51,
     }
 )
 solver = pybamm.IDAKLUSolver()

### 2024-12-20 pybop/__init__.py

@@ -123,7 +123,11 @@ from .costs._weighted_cost import WeightedCost
 # Experimental
 #
 from .experimental.jax_costs import BaseJaxCost, JaxSumSquaredError, JaxLogNormalLikelihood, JaxGaussianLogLikelihoodKnownSigma
-from .experimental._evaluation import SequentialJaxEvaluator, SciPyEvaluator
+
+#
+# Evaluation
+#
+from ._evaluation import SequentialJaxEvaluator, SciPyEvaluator
 
 #
 # Optimiser classes

### 2024-12-20 pybop/_evaluation.py

@@ -1,15 +1,14 @@
+import jax.numpy as jnp
 import numpy as np
 from pints import Evaluator as PintsEvaluator
 
 
 class SequentialJaxEvaluator(PintsEvaluator):
     """
-    Evaluates a JAX function (or callable object) for a list of input values,
-    and returns a list or NumPy array of the calculated function evaluations.
-
-    This class extends PintsEvaluator to provide an interface for evaluating
-    functions using JAX, with optional additional arguments passed to the
-    function.
+    Sequential evaluates a function (or callable object)
+    for either a single or multiple positions. This class is based
+    off the PintsSequentialEvaluator class, with additions for
+    PyBOP's JAX cost classes.
 
     Parameters
     ----------
@@ -20,9 +19,6 @@ class SequentialJaxEvaluator(PintsEvaluator):
         A sequence containing extra arguments to be passed to the function.
         If specified, the function will be called as `function(x, *args)`.
     """
-
-    def __init__(self, function, args=None):
-        super().__init__(function, args)
 
     def _evaluate(self, positions):
         scores = [self._function(x, *self._args) for x in positions]
@@ -36,12 +32,8 @@ class SequentialJaxEvaluator(PintsEvaluator):
 
 class SciPyEvaluator(PintsEvaluator):
     """
-    Evaluates a function (or callable object) with SciPy for a list of input values,
-    and returns a list or NumPy array of the calculated function evaluations.
-
-    This class extends PintsEvaluator to provide an interface for evaluating
-    functions using JAX, with optional additional arguments passed to the
-    function.
+    Evaluates a function (or callable object) for the SciPy optimisers
+    for either a single or multiple positions.
 
     Parameters
     ----------
@@ -53,14 +45,13 @@ class SciPyEvaluator(PintsEvaluator):
         If specified, the function will be called as `function(x, *args)`.
     """
 
-    def __init__(self, function, args=None):
-        super().__init__(function, args)
-
     def _evaluate(self, positions):
         scores = [self._function(x, *self._args) for x in [positions]]
 
         # If gradient provided, convert jnp to np and return
-        if isinstance(scores[0], tuple):
-            return [(score[0].item(), score[1]) for score in scores]
+        if not isinstance(scores[0], tuple):
+            return np.asarray(scores)[0]
 
-        return np.asarray(scores)[0]
+        if isinstance(scores[0][0], jnp.ndarray):
+            return [(score[0].item(), score[1]) for score in scores][0]
+        return [(score[0], score[1]) for score in scores][0]

### 2024-12-20 pybop/experimental/jax_costs.py

@@ -44,15 +44,17 @@ class BaseJaxCost(BaseCost):
         float
             The Sum of Squared Error.
         """
-        inputs = self.parameters.verify(inputs)
+        self.has_transform = self.transformation is not None and apply_transform
+        model_inputs = self.parameters.verify(self._apply_transformations(inputs))
+
         if calculate_grad != self.model.calculate_sensitivities:
             self._update_solver_sensitivities(calculate_grad)
 
         if calculate_grad:
-            y, dy = jax.value_and_grad(self.evaluate)(inputs)
+            y, dy = jax.value_and_grad(self.evaluate)(model_inputs)
             return y, np.asarray(list(dy.values()))
         else:
-            return self.evaluate(inputs)
+            return self.evaluate(model_inputs)
 
     def _update_solver_sensitivities(self, calculate_grad: bool) -> None:
         """

### 2024-12-20 pybop/models/base_model.py

@@ -745,33 +745,28 @@ class BaseModel:
 
     def jaxify_solver(self, t_eval, calculate_sensitivities=False):
         """
-        Jaxify the IDAKLU Solver and store a copy for reconstructing
-        the jax'ed solver in the future. This is helpful for reconstructing
-        the solver with `calculate_sensitivities=True` as this solver requires
-        casting sensitivity information on construction.
+        Jaxify the IDAKLU Solver and store a copy for future reconstruction.
+        Handles sensitivity calculations during solver construction.
         """
         self._calculate_sensitivities = calculate_sensitivities
+
+        if not isinstance(self._solver, (pybamm.IDAKLUSolver, pybamm.IDAKLUJax)):
+            raise ValueError("Solver must be pybamm.IDAKLUSolver to jaxify.")
+
+        # Store original solver if not already stored, and create local copy
         if isinstance(self._solver, pybamm.IDAKLUSolver):
             self._IDAKLU_stored = self._solver.copy()
-            self._solver = self._solver.jaxify(
-                model=self._built_model,
-                t_eval=t_eval
-                if calculate_sensitivities is True
-                else [t_eval[0], t_eval[-1]],  # Current bug in PyBaMM v24.11
-                t_interp=t_eval,
-                calculate_sensitivities=calculate_sensitivities,
-            )
-        elif isinstance(self._solver, pybamm.IDAKLUJax):
-            self._solver = self._IDAKLU_stored.jaxify(
-                model=self._built_model,
-                t_eval=t_eval
-                if calculate_sensitivities is True
-                else [t_eval[0], t_eval[-1]],  # Current bug in PyBaMM v24.11
-                t_interp=t_eval,
-                calculate_sensitivities=calculate_sensitivities,
-            )
-        else:
-            raise ValueError("Solver is not pybamm.IDAKLUSolver, cannot jaxify it.")
+        base_solver = self._IDAKLU_stored
+
+        # Handle PyBaMM v24.11 bug: use full t_eval only when calculating sensitivities
+        t_eval_adjusted = t_eval if calculate_sensitivities else [t_eval[0], t_eval[-1]]
+
+        self._solver = base_solver.jaxify(
+            model=self._built_model,
+            t_eval=t_eval_adjusted,
+            t_interp=t_eval,
+            calculate_sensitivities=calculate_sensitivities,
+        )
 
     def check_params(
         self,

### 2024-12-20 pybop/optimisers/base_pints_optimiser.py

@@ -264,7 +264,7 @@ class BasePintsOptimiser(BaseOptimiser):
                     x=xs,
                     x_best=self.optimiser.x_best(),
                     cost=_fs,
-                    cost_best=fb,  # TODO: check shape, [fb]?
+                    cost_best=fb,
                 )
 
                 # Check stopping criteria:

### 2024-12-20 pybop/optimisers/scipy_optimisers.py

@@ -199,6 +199,9 @@ class SciPyMinimize(BaseSciPyOptimiser):
                 # Nest this option within an options dictionary for SciPy minimize
                 self._options["options"]["maxiter"] = self.unset_options.pop(key)
 
+        if self._options["jac"] is True:
+            self._needs_sensitivities = True
+
     def cost_wrapper(self, x):
         """
         Scale the cost function, preserving the sign convention, and eliminate nan values
@@ -214,7 +217,7 @@ class SciPyMinimize(BaseSciPyOptimiser):
                 )  # for fake finite gradient
             return scaled_cost
 
-        L, dl = self.cost_call(x, calculate_grad=True)
+        L, dl = self.evaluator.evaluate(x)
         self.log_update(x=[x], cost=L)
         return (L / self._cost0, dl / self._cost0)
 
@@ -242,7 +245,8 @@ class SciPyMinimize(BaseSciPyOptimiser):
                 cost_best = intermediate_result.fun * self._cost0
             else:
                 x_best = intermediate_result
-                cost_best = self.cost_call(x_best)
+                result = self.evaluator.evaluate(x_best)
+                cost_best = result[0] if self._needs_sensitivities else result
 
             self.log_update(x_best=x_best, cost_best=cost_best)
 
@@ -253,7 +257,8 @@ class SciPyMinimize(BaseSciPyOptimiser):
         )
 
         # Compute the absolute initial cost and resample if required
-        self._cost0 = np.abs(self.cost_call(self.x0))
+        result = self.evaluator.evaluate(self.x0)
+        self._cost0 = np.abs(result[0] if self._needs_sensitivities else result)
         if np.isinf(self._cost0):
             for _i in range(1, self.num_resamples):
                 try:
@@ -265,7 +270,8 @@ class SciPyMinimize(BaseSciPyOptimiser):
                         stacklevel=2,
                     )
                     break
-                self._cost0 = np.abs(self.cost_call(self.x0))
+                result = self.evaluator.evaluate(self.x0)
+                self._cost0 = np.abs(result[0] if self._needs_sensitivities else result)
                 if not np.isinf(self._cost0):
                     break
             if np.isinf(self._cost0):

### 2024-12-20 tests/unit/experimental/test_jax_costs.py

@@ -138,7 +138,7 @@ class TestJaxCosts:
         # Test jaxify with incorrect solver
         model.solver = pybamm.CasadiSolver()
         with pytest.raises(
-            ValueError, match="Solver is not pybamm.IDAKLUSolver, cannot jaxify it."
+            ValueError, match="Solver must be pybamm.IDAKLUSolver to jaxify."
         ):
             model.jaxify_solver(t_eval=np.linspace(0, 1, 100))
 

### 2024-12-20 tests/unit/test_optimisation.py

@@ -338,6 +338,11 @@ class TestOptimisation:
         assert optim.optimiser._boundaries is None
 
     @pytest.mark.unit
+    def test_adamw_impl_bounds(self):
+        with pytest.warns(UserWarning, match="Boundaries ignored by AdamW"):
+            pybop.AdamWImpl(x0=[0.1], boundaries=[0.0, 0.2])
+
+    @pytest.mark.unit
     def test_scipy_minimize_with_jac(self, cost):
         # Check a method that uses gradient information
         optim = pybop.SciPyMinimize(

### 2024-12-20 tests/unit/test_plots.py

@@ -100,6 +100,7 @@ class TestPlots:
         # Test plot of Problem objects
         pybop.plot.quick(fitting_problem, title="Optimised Comparison")
         pybop.plot.quick(design_problem)
+        pybop.plot.quick(jax_fitting_problem)
 
         # Test conversion of values into inputs
         pybop.plot.quick(fitting_problem, problem_inputs=[0.6, 0.6])

### 2024-12-20 examples/scripts/comparison_examples/maximum_likelihood.py
Adds SciPyEvaluator, updates examples, tests.

@@ -11,9 +11,8 @@ parameter_set.update(
         "Positive electrode active material volume fraction": 0.62,
     }
 )
-options = {"max_num_steps": int(1e6), "max_error_test_failures": 60}
-solver = pybamm.IDAKLUSolver(atol=1e-6, rtol=1e-6, options=options)
-model = pybop.lithium_ion.DFN(parameter_set=parameter_set, solver=solver)
+solver = pybamm.IDAKLUSolver()
+model = pybop.lithium_ion.SPMe(parameter_set=parameter_set, solver=solver)
 
 # Fitting parameters
 parameters = pybop.Parameters(
@@ -60,7 +59,7 @@ dataset = pybop.Dataset(
 signal = ["Voltage [V]", "Bulk open-circuit voltage [V]"]
 # Generate problem, cost function, and optimisation class
 problem = pybop.FittingProblem(model, parameters, dataset, signal=signal)
-likelihood = pybop.JaxGaussianLogLikelihoodKnownSigma(problem, sigma0=sigma)
+likelihood = pybop.GaussianLogLikelihoodKnownSigma(problem, sigma0=sigma)
 optim = pybop.XNES(
     likelihood,
     max_unchanged_iterations=20,
@@ -79,9 +78,6 @@ pybop.plot.convergence(optim)
 
 # Plot the parameter traces
 pybop.plot.parameters(optim)
-
-# Plot the cost landscape
-pybop.plot.contour(likelihood, steps=15)
 
 # Plot the cost landscape with optimisation path
 bounds = np.asarray([[0.55, 0.77], [0.48, 0.68]])

### 2024-12-20 examples/scripts/getting_started/jax-solver-example.py

@@ -16,12 +16,13 @@ parameters = pybop.Parameters(
     pybop.Parameter(
         "Negative electrode active material volume fraction",
         initial_value=0.55,
+        prior=pybop.Gaussian(0.6, 0.03),
         bounds=[0.5, 0.8],
     ),
     pybop.Parameter(
         "Positive electrode active material volume fraction",
         initial_value=0.55,
-        bounds=[0.5, 0.8],
+        prior=pybop.Gaussian(0.6, 0.03),
     ),
 )
 

### 2024-12-20 pybop/__init__.py

@@ -123,7 +123,7 @@ from .costs._weighted_cost import WeightedCost
 # Experimental
 #
 from .experimental.jax_costs import BaseJaxCost, JaxSumSquaredError, JaxLogNormalLikelihood, JaxGaussianLogLikelihoodKnownSigma
-from .experimental._evaluation import SequentialJaxEvaluator
+from .experimental._evaluation import SequentialJaxEvaluator, SciPyEvaluator
 
 #
 # Optimiser classes

### 2024-12-20 pybop/experimental/_evaluation.py

@@ -32,3 +32,35 @@ class SequentialJaxEvaluator(PintsEvaluator):
             return [(score[0].item(), score[1]) for score in scores]
 
         return np.asarray(scores)
+
+
+class SciPyEvaluator(PintsEvaluator):
+    """
+    Evaluates a function (or callable object) with SciPy for a list of input values,
+    and returns a list or NumPy array of the calculated function evaluations.
+
+    This class extends PintsEvaluator to provide an interface for evaluating
+    functions using JAX, with optional additional arguments passed to the
+    function.
+
+    Parameters
+    ----------
+    function : callable
+        The function to evaluate. This function should accept an input and
+        optionally additional arguments, returning either a single value or a tuple.
+    args : sequence, optional
+        A sequence containing extra arguments to be passed to the function.
+        If specified, the function will be called as `function(x, *args)`.
+    """
+
+    def __init__(self, function, args=None):
+        super().__init__(function, args)
+
+    def _evaluate(self, positions):
+        scores = [self._function(x, *self._args) for x in [positions]]
+
+        # If gradient provided, convert jnp to np and return
+        if isinstance(scores[0], tuple):
+            return [(score[0].item(), score[1]) for score in scores]
+
+        return np.asarray(scores)[0]

### 2024-12-20 pybop/experimental/jax_costs.py

@@ -25,6 +25,7 @@ class BaseJaxCost(BaseCost):
         inputs: Inputs,
         calculate_grad: bool = False,
         apply_transform: bool = False,
+        for_optimiser: bool = False,
     ) -> Union[np.array, tuple[float, np.ndarray]]:
         """
         Computes the cost function for the given predictions.

### 2024-12-20 pybop/optimisers/scipy_optimisers.py

@@ -5,7 +5,7 @@ from typing import Union
 import numpy as np
 from scipy.optimize import Bounds, OptimizeResult, differential_evolution, minimize
 
-from pybop import BaseOptimiser, OptimisationResult
+from pybop import BaseOptimiser, OptimisationResult, SciPyEvaluator
 
 
 class BaseSciPyOptimiser(BaseOptimiser):
@@ -77,6 +77,14 @@ class BaseSciPyOptimiser(BaseOptimiser):
         result : pybop.Result
             The result of the optimisation including the optimised parameter values and cost.
         """
+
+        # Choose method to evaluate
+        def fun(x):
+            return self.cost_call(x, calculate_grad=self._needs_sensitivities)
+
+        # Create evaluator object
+        self.evaluator = SciPyEvaluator(fun)
+
         # Run with timing
         start_time = time()
         result = self._run_optimiser()
@@ -196,7 +204,7 @@ class SciPyMinimize(BaseSciPyOptimiser):
         Scale the cost function, preserving the sign convention, and eliminate nan values
         """
         if not self._options["jac"]:
-            cost = self.cost_call(x)
+            cost = self.evaluator.evaluate(x)
             self.log_update(x=[x], cost=cost)
             scaled_cost = cost / self._cost0
             if np.isinf(scaled_cost):
@@ -427,7 +435,7 @@ class SciPyDifferentialEvolution(BaseSciPyOptimiser):
             )
 
         def cost_wrapper(x):
-            cost = self.cost_call(x)
+            cost = self.evaluator.evaluate(x)
             self.log_update(x=[x], cost=cost)
             return cost
 

### 2024-12-20 tests/integration/test_spm_parameterisations.py

@@ -125,6 +125,11 @@ class Test_SPM_Parameterisation:
             if isinstance(cost, pybop.GaussianLogLikelihood)
             else sigma0,
         }
+        if (
+            isinstance(cost, pybop.JaxSumSquaredError)
+            and optimiser is pybop.SciPyDifferentialEvolution
+        ):
+            common_args["bounds"] = [[0.375, 0.775], [0.375, 0.775]]
 
         if isinstance(cost, pybop.LogPosterior):
             for i in cost.parameters.keys():
@@ -218,6 +223,12 @@ class Test_SPM_Parameterisation:
             if isinstance(spm_two_signal_cost, pybop.GaussianLogLikelihood)
             else 0.02,
         }
+
+        if (
+            isinstance(spm_two_signal_cost, pybop.JaxSumSquaredError)
+            and multi_optimiser is pybop.SciPyDifferentialEvolution
+        ):
+            common_args["bounds"] = [[0.375, 0.775], [0.375, 0.775]]
 
         # Test each optimiser
         optim = multi_optimiser(**common_args)

### 2024-12-20 pybop/__init__.py
Update function name

@@ -173,7 +173,7 @@ from .observers.observer import Observer
 #
 # Classification classes
 #
-from ._classification import classify_using_Hessian
+from ._classification import classify_using_hessian
 
 #
 # Plotting classes

### 2024-12-20 pybop/_classification.py

@@ -5,7 +5,7 @@ import numpy as np
 from pybop import OptimisationResult
 
 
-def classify_using_Hessian(
+def classify_using_hessian(
     result: OptimisationResult, dx=None, cost_tolerance: Optional[float] = 1e-5
 ):
     """
@@ -32,7 +32,7 @@ def classify_using_Hessian(
     n = len(x)
     if n != 2 or len(dx) != n:
         raise ValueError(
-            "The function classify_using_Hessian currently only works in the case "
+            "The function classify_using_hessian currently only works in the case "
             "of 2 parameters, and dx must have the same length as x."
         )
 
@@ -51,10 +51,7 @@ def classify_using_Hessian(
                 costs[i, j, 1] = cost(x + np.multiply([i - 1, j - 1], 2 * dx))
 
     # Classify the result
-    if not np.all([np.isfinite(cost) for cost in costs]):
-        message = "Classification cannot proceed due to infinite cost value(s)."
-
-    elif (minimising and np.any(costs < final_cost)) or (
+    if (minimising and np.any(costs < final_cost)) or (
         not minimising and np.any(costs > final_cost)
     ):
         message = "The optimiser has not converged to a stationary point."
@@ -68,6 +65,9 @@ def classify_using_Hessian(
 
                 if value < bounds["lower"][i] + dx[i]:
                     message += f" The result is near the lower bound of {names[i]}."
+
+    elif not np.all([np.isfinite(cost) for cost in costs]):
+        message = "Classification cannot proceed due to infinite cost value(s)."
 
     else:
         # Estimate the Hessian using second-order accurate central finite differences

### 2024-12-20 tests/integration/test_classification.py

@@ -70,7 +70,7 @@ class TestClassification:
         return pybop.FittingProblem(model, parameters, dataset)
 
     @pytest.mark.integration
-    def test_classify_using_Hessian(self, problem):
+    def test_classify_using_hessian(self, problem):
         cost = pybop.RootMeanSquaredError(problem)
         x = cost.parameters.true_value()
         bounds = cost.parameters.get_bounds()
@@ -79,16 +79,16 @@ class TestClassification:
         results = pybop.OptimisationResult(x=x0, optim=optim)
 
         if np.all(x == np.asarray([0.05, 0.05])):
-            message = pybop.classify_using_Hessian(results)
+            message = pybop.classify_using_hessian(results)
             assert message == "The optimiser has located a minimum."
         elif np.all(x == np.asarray([0.1, 0.05])):
-            message = pybop.classify_using_Hessian(results)
+            message = pybop.classify_using_hessian(results)
             assert message == (
                 "The optimiser has not converged to a stationary point."
                 " The result is near the upper bound of R0 [Ohm]."
             )
         elif np.all(x == np.asarray([0.05, 0.01])):
-            message = pybop.classify_using_Hessian(results)
+            message = pybop.classify_using_hessian(results)
             assert message == (
                 "The optimiser has not converged to a stationary point."
                 " The result is near the lower bound of R1 [Ohm]."
@@ -101,14 +101,14 @@ class TestClassification:
             optim = pybop.Optimisation(cost=cost)
             results = pybop.OptimisationResult(x=x, optim=optim)
 
-            message = pybop.classify_using_Hessian(results)
+            message = pybop.classify_using_hessian(results)
             assert message == "The optimiser has located a maximum."
 
-        # message = pybop.classify_using_Hessian(results)
+        # message = pybop.classify_using_hessian(results)
         # assert message == "The optimiser has located a saddle point."
 
     @pytest.mark.integration
-    def test_insensitive_classify_using_Hessian(self, parameter_set):
+    def test_insensitive_classify_using_hessian(self, parameter_set):
         param_R0_a = pybop.Parameter(
             "R0_a [Ohm]",
             bounds=[0, 0.002],
@@ -150,22 +150,22 @@ class TestClassification:
             optim = pybop.Optimisation(cost=cost)
             results = pybop.OptimisationResult(x=x, optim=optim)
 
-            message = pybop.classify_using_Hessian(results)
+            message = pybop.classify_using_hessian(results)
             assert message == (
[…]

### 2024-12-20 tests/unit/test_classifier.py

@@ -36,7 +36,7 @@ class TestClassifier:
         return pybop.FittingProblem(model, parameters, dataset)
 
     @pytest.mark.unit
-    def test_classify_using_Hessian_invalid(self, problem):
+    def test_classify_using_hessian_invalid(self, problem):
         cost = pybop.SumSquaredError(problem)
         optim = pybop.Optimisation(cost=cost)
         x = np.asarray([0.001])
@@ -44,7 +44,7 @@ class TestClassifier:
 
         with pytest.raises(
             ValueError,
-            match="The function classify_using_Hessian currently only works"
+            match="The function classify_using_hessian currently only works"
             " in the case of 2 parameters, and dx must have the same length as x.",
         ):
-            pybop.classify_using_Hessian(results)
+            pybop.classify_using_hessian(results)

### 2024-12-20 pybop/_classification.py
Check bounds proximity if infinite cost

@@ -50,24 +50,27 @@ def classify_using_hessian(
                 costs[i, j, 0] = cost(x + np.multiply([i - 1, j - 1], dx))
                 costs[i, j, 1] = cost(x + np.multiply([i - 1, j - 1], 2 * dx))
 
+    def check_proximity_to_bounds(parameters, x, dx, names) -> str:
+        bounds = parameters.get_bounds()
+        if bounds is not None:
+            for i, value in enumerate(x):
+                if value > bounds["upper"][i] - dx[i]:
+                    return f" The result is near the upper bound of {names[i]}."
+
+                if value < bounds["lower"][i] + dx[i]:
+                    return f" The result is near the lower bound of {names[i]}."
+        return ""
+
     # Classify the result
     if (minimising and np.any(costs < final_cost)) or (
         not minimising and np.any(costs > final_cost)
     ):
         message = "The optimiser has not converged to a stationary point."
-
-        # Check proximity to bounds
-        bounds = parameters.get_bounds()
-        if bounds is not None:
-            for i, value in enumerate(x):
-                if value > bounds["upper"][i] - dx[i]:
-                    message += f" The result is near the upper bound of {names[i]}."
-
-                if value < bounds["lower"][i] + dx[i]:
-                    message += f" The result is near the lower bound of {names[i]}."
+        message += check_proximity_to_bounds(parameters, x, dx, names)
 
     elif not np.all([np.isfinite(cost) for cost in costs]):
         message = "Classification cannot proceed due to infinite cost value(s)."
+        message += check_proximity_to_bounds(parameters, x, dx, names)
 
     else:
         # Estimate the Hessian using second-order accurate central finite differences

### 2024-12-20 tests/integration/test_classification.py

@@ -168,4 +168,7 @@ class TestClassification:
         )
 
         message = pybop.classify_using_hessian(results, dx=[1, 1])
-        assert message == "Classification cannot proceed due to infinite cost value(s)."
+        assert message == (
+            "Classification cannot proceed due to infinite cost value(s)."
+            " The result is near the upper bound of R0_a [Ohm]."
+        )

### 2024-12-20 pybop/__init__.py
Add Hessian check for 2-parameter problems (#363)
* Add Hessian check

* Update test_classification.py

* Update to implicitly concatenated strings

* Update test_classification

* Add tests on insensitivity

* Update CHANGELOG.md

* Fix typo

Co-authored-by: Brady Planden <55357039+BradyPlanden@users.noreply.github.com>

* Update input to OptimisationResult

* Improve insensitivity tests

* Improve correlation checks

* Update in line with OptimisationResult

* Make one test a unit test

* Update function name

* Check bounds proximity if infinite cost

---------

Co-authored-by: Brady Plande

@@ -171,6 +171,11 @@ from .observers.unscented_kalman import UnscentedKalmanFilterObserver
 from .observers.observer import Observer
 
 #
+# Classification classes
+#
+from ._classification import classify_using_hessian
+
+#
 # Plotting classes
 #
 from . import plot as plot

### 2024-12-20 tests/unit/test_classifier.py

@@ -1,0 +1,50 @@
+import numpy as np
+import pytest
+
+import pybop
+
+
+class TestClassifier:
+    """
+    A class to test the classification of different optimisation results.
+    """
+
+    @pytest.fixture
+    def problem(self):
+        model = pybop.empirical.Thevenin()
+        experiment = pybop.Experiment(
+            [
+                "Discharge at 0.5C for 2 minutes (4 seconds period)",
+                "Charge at 0.5C for 2 minutes (4 seconds period)",
+            ]
+        )
+        solution = model.predict(experiment=experiment)
+        dataset = pybop.Dataset(
+            {
+                "Time [s]": solution["Time [s]"].data,
+                "Current function [A]": solution["Current [A]"].data,
+                "Voltage [V]": solution["Voltage [V]"].data,
+            }
+        )
+        parameters = pybop.Parameters(
+            pybop.Parameter(
+                "R0 [Ohm]",
+                prior=pybop.Uniform(0.001, 0.1),
+                bounds=[1e-4, 0.1],
+            ),
+        )
+        return pybop.FittingProblem(model, parameters, dataset)
+
+    @pytest.mark.unit
+    def test_classify_using_hessian_invalid(self, problem):
+        cost = pybop.SumSquaredError(problem)
+        optim = pybop.Optimisation(cost=cost)
+        x = np.asarray([0.001])
+        results = pybop.OptimisationResult(x=x, optim=optim)
+
+        with pytest.raises(
+            ValueError,
+            match="The function classify_using_hessian currently only works"
+            " in the case of 2 parameters, and dx must have the same length as x.",
+        ):
+            pybop.classify_using_hessian(results)
